### 2016-12-01 diagrams/context_distribution.tex
chapter 4

\begin{tikzpicture}[thick,scale=0.8, every node/.style={transform shape}]
\begin{axis}[
    ybar stacked,
	bar width=12pt,
	nodes near coords = {%
    \pgfmathprintnumberto[fixed,assume math mode=true]{\pgfplotspointmeta}{\myval}%
    \pgfmathparse{\myval<101?:\myval}\pgfmathresult%
},
    enlargelimits=0.15,
    legend style={at={(0.5,-0.28)},
      anchor=north,legend columns=-1},
    ylabel={Proportion [\%]},
    symbolic x coords={AddressBook, TimelClock, WebChess, Drupal,
    phpBB, phpMyAdmin, Anchor, Kirby, Automad, Monstra, Nibbleblog}, xtick=data,
    x tick label style={rotate=45,anchor=east},
    cycle list name=ColorListBar,
    ]

\addplot+[ybar] plot coordinates {(AddressBook, 28) 
  (TimelClock, 0) (WebChess, 86) (Drupal, 100) (phpBB, 96) (phpMyAdmin, 92)
  (Anchor, 100) (Kirby, 96) (Automad, 93) (Monstra, 83) (Nibbleblog, 91)};
\addplot+[ybar] plot coordinates {(AddressBook, 5) 
  (TimelClock, 0) (WebChess, 0) (Drupal, 0) (phpBB, 0) (phpMyAdmin, 1)
  (Anchor, 0) (Kirby, 4) (Automad, 0) (Monstra, 2) (Nibbleblog, 0)};
\addplot+[ybar] plot coordinates {(AddressBook, 67) 
  (TimelClock, 100) (WebChess, 14) (Drupal, 0) (phpBB, 4) (phpMyAdmin, 7)
  (Anchor, 0) (Kirby, 0) (Automad, 7) (Monstra, 15) (Nibbleblog, 9)};

\legend{Function Definition, HTML Template, Miscellaneous}
\end{axis}
%\caption{Relative distribution of contexts of not reached output candidates}
\end{tikzpicture}

### 2016-12-01 diagrams/coverage.tex
chapter 4

\pgfplotstableread[header=false, col sep=comma]{ 
	86.32, AdressBook
	100.0, SchoolMate
	92.36, TimeClock
	91.28, WebChess
	31.44, phpBB
	19.99, Drupal
	5.000, phpMyAdmin
	90.48, Anchor
	76.3,  Kirby
	46.87, Automad
	79.47, Monstra
	71.37, Nibbleblog
	}\datatablex

	\pgfplotstableread[header=false, col sep=comma]{ 
	84.44, AdressBook
	98.24, SchoolMate
	91.22, TimeClock
	88.72, WebChess
	4.76, phpBB
	18.39, Drupal
	3.79, phpMyAdmin
	89.56, Anchor
	76.30,  Kirby
	40.15, Automad
	61.38, Monstra
	59.43, Nibbleblog
	}\datatable

	\begin{tikzpicture}[thick,scale=0.8, every node/.style={transform shape}]
	  \begin{axis}[   
	  	,legend style={at={(0.25,-0.2)},
		anchor=north,legend columns=-1} 
	    ,xbar
	    ,xlabel={Coverage [\%]}
	    ,bar width=1ex, y=3ex
	    ,enlarge y limits={abs=0.75}
	    ,ytick={data} % (\empty,data,{coordinate list})
	    ,label={Reach}
	    ,yticklabels from table={\datatable}{1}
	  ]

	  \addplot[red!20!black,fill=gray!80!white] table [
	    y expr=-\coordindex, % Use negative coordinate index as y coordinate
	    x index=0 % Use first column as x coordinate
	  ] {\datatable};
	  \addplot[red!20!black,fill=RYB2!100!white] table [
	    y expr=-\coordindex, % Use negative coordinate index as y coordinate
	    x index=0 % Use first column as x coordinate
	  ] {\datatablex};

	  \legend{Output, Reached}
	  \end{axis}
	\end{tikzpicture}

### 2016-12-01 diagrams/include_coverage.tex
chapter 4

        \pgfplotstableread[header=false, col sep=comma]{ 
		96.77, AdressBook
		100, SchoolMate
		92.48, TimeClock
		94.64, WebChess
		65.42, phpBB
		49.03, Drupal
		39.75, phpMyAdmin
		84.38, Anchor
		39.13,  Kirby
		37.5, Automad
		37.5, Monstra
		100, Nibbleblog
		}\reachedincludes

		\begin{tikzpicture}[thick,scale=0.8, every node/.style={transform shape}]
		  \begin{axis}[   
		  	,legend style={at={(0.25,-0.2)},
			anchor=north,legend columns=-1} 
		    ,xbar
		    ,xlabel={Reached include expressions [\%]}
		    ,bar width=1.2ex, y=2ex
		    ,enlarge y limits={abs=0.75}
		    ,ytick={data} % (\empty,data,{coordinate list})
		    ,label={Reach}
		    ,yticklabels from table={\reachedincludes}{1}
		  ]

		  \addplot[red!20!black,fill=RYB10!80!white] table [
		    y expr=-\coordindex, % Use negative coordinate index as y coordinate
		    x index=0 % Use first column as x coordinate
		  ] {\reachedincludes};

		  \end{axis}
		\end{tikzpicture}

### 2016-12-01 diagrams/success_rate.tex
chapter 4

\pgfplotstableread[header=false, col sep=comma]{ 
		95, AdressBook
		100, SchoolMate
		100, TimeClock
		100, WebChess
		74, phpBB
		88, Drupal
		99, phpMyAdmin
		0, Anchor
		0,  Kirby
		100, Automad
		27, Monstra
		90, Nibbleblog
		}\successfulincludes

		\begin{tikzpicture}[thick,scale=0.8, every node/.style={transform shape}]
		  \begin{axis}[   
		  	,legend style={at={(0.25,-0.2)},
			anchor=north,legend columns=-1} 
		    ,xbar
		    ,xlabel={Successfully resolved and reached include expressions [\%]}
		    ,bar width=1.2ex, y=2ex
		    ,enlarge y limits={abs=0.75}
		    ,ytick={data} % (\empty,data,{coordinate list})
		    ,label={Reach}
		    ,yticklabels from table={\successfulincludes}{1}
		  ]

		  \addplot[red!20!black,fill=RYB11!80!white] table [
		    y expr=-\coordindex, % Use negative coordinate index as y coordinate
		    x index=0 % Use first column as x coordinate
		  ] {\successfulincludes};

		  \end{axis}
		\end{tikzpicture}

### 2016-12-01 paper.tex
chapter 4

\documentclass{sig-alternate-05-2015}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{booktabs}

\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{ucs}
\usepackage{framed}

\usepackage[labelfont=bf]{caption}

\hypersetup{
    colorlinks,
    linkcolor={red!65!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\usepackage{listings}

% ARIAL
%\usepackage{helvet}
%\renewcommand{\familydefault}{\sfdefault}

% TIMES
\usepackage{mathptmx}% http://ctan.org/pkg/mathptmx

\usepackage{pgfplots, pgfplotstable}

\usepackage{xcolor}

\definecolor{orange}{RGB}{230, 100, 20}
\definecolor{gruen}{RGB}{255, 190, 69}

\definecolor{RYB1}{HTML}{8F83DB}
\definecolor{RYB2}{HTML}{DDDDDD}
\definecolor{RYB3}{HTML}{FFFFFF}

\definecolor{RYB4}{HTML}{A699FF}
\definecolor{RYB5}{HTML}{FFBA49}
\definecolor{RYB6}{HTML}{FFFFFF}

\definecolor{RYB7}{HTML}{6A62A3}
\definecolor{RYB8}{HTML}{BEB4FF}
\definecolor{RYB9}{HTML}{E6E3FF}

\definecolor{RYB10}{HTML}{FFB030}
\definecolor{RYB11}{HTML}{FFCD7C}

\pgfplotscreateplotcyclelist{ColorListBar}{
{fill=RYB1},
{fill=RYB2},
{fill=RYB3},
}

\pgfplotscreateplotcyclelist{ColorListBar2}{
{fill=RYB4},
{fill=RYB5},
{fill=RYB6},
}

\pgfplotscreateplotcyclelist{ColorListBar3}{
{fill=RYB7},
{fill=RYB8},
{fill=RYB9},
}

\begin{document}
\lstset{
	language=PHP,
	basicstyle=\footnotesize\ttfamily,
    commentstyle = \color{gray},
    extendedchars = \true,
    inputencoding = utf8x,
    keepspaces = true,
    keywordstyle = \bfseries
}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{10.475/123_4}

% ISBN
\isbn{123-4567-24-567/08/06}

%Conference
\conferenceinfo{ISSTA '17}{July 9--13, 2013, Santa Barbara, CA, USA}

\acmPrice{\$15.00}

%
% --- Author Metadata here ---
\conferenceinfo{ISSTA '17}{July 9--13, 2013, Santa Barbara, CA, USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data
% (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Does It Scale? Static Output Approximation of PHP Web Applications Using
Symbolic Execution}

%

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Stefan Mühlbauer\\
       \affaddr{TU Braunschweig, Germany}\\
       %\email{s.muehlbauer@tu-bs.de}
% 2nd. author
\alignauthor
%Christian Kästner\\
%      \affaddr{Carnegie Mellon University, USA}\\
       %\email{ckaestner@cs.cmu.edu}
% 3rd. author
\alignauthor
%Tien N. Nguyen\\
%       \affaddr{University of Texas at Dallas, USA}\\
%      %\email{tien.n.nguyen@utdallas.edu}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
Dynamic web applications have become widely popular and are to a large
proportion based on the scripting language PHP. Output approximation of web
applications enables a range of additional tool support as well as possibilities
for vulnerability detection. Unfortunately, recent approximation approaches have
only been evaluated for smaller systems due to both conceptual limitations and
unsupported language features.

This paper presents an experience report about the scalability of output
approximation using symbolic execution of state-of-the-art PHP web
applications. For a symbolic execution engine extended with support for
object-oriented programming and arrays, we identified language features and
corresponding programming patterns that impede symbolic execution and limit the
scalability of this approach. Our findings include: (1) dynamic features such as
functions and includes are prone to fail for certain programming patterns. (2)
Expressions containing elements from I/O, databases or files can heavily impede
symbolic execution. Our findings provide useful guidelines to design new
tools and also to improve the development process of statically analyzable web
applications.

\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%
%  Use this command to print the description
%
\printccsdesc

% We no longer use \terms command
%\terms{Theory}

\keywords{Static program behavior, Static metrics, Dynamic language features,
Static Analysis, PHP}

\section{Introduction}
Dynamic web applications have become widely popular and various different
implementation techniques have emerged, ranging from server-side scripting
languages to state-of-the-art frameworks for web development. For scritpting
languages such as PHP, source code usually contains script code along with nested
target code since the HTML pages are dynamically generated.

\section{State Of the Art}
\subsection{Symbolic Execution (and Choices?)}
\subsection{Use Cases for Output Approximation}

\begin{figure}

	\centering
	\includegraphics[width=0.5\textwidth]{images-paper/approach}
	\caption{Approach overview; the different analyses in Varis are based upon the
	symbolic output. Step 1 is the \emph{symbolic execution}, step 2 is the
	\emph{variability-aware parsing} and step enfolds subsequent analyses based
	upon the symbolic output.}
	\label{approach}
\end{figure}

\begin{itemize}
  \item Steps: Symbolic Execution $\rightarrow$ Parsing $\rightarrow$
  Analysis, see Figure \ref{approach}
  \item Varis: IDE support \cite{Nguyen:2015:VIS:2819009.2819140} for PHP with
  code slicing \cite{Nguyen:2015:CPS:2786805.2786872}, code navigation
  \cite{Nguyen:2014:BCG:2635868.2635928} and
  HTML validation \cite{Nguyen:2011:AFH:2190078.2190142}.
\end{itemize}

\section{Towards scalable symbolic execution}
\subsection{Can we have a practical tool?}
\begin{itemize}
  \item Are the assumprions just engineering or do not matter for accuracy? 
\end{itemize}
\subsection{Oak: A Symbolic Interpreter}
\begin{itemize}
  \item \textsc{Oak}\footnote{\url{https://github.com/smba/oak}}:
  Re-implementing and extending the symbolic execution engine with OOP and array support
  \item Test-driven implementation strategy 
  \item WordPress!
\end{itemize}
\subsection{Experience Report}
\subsubsection{Problems}
\subsubsection{Concrete/Symbolic Boundary}
\subsubsection{Further experiments}
\begin{itemize}
  \item SPL approach for Oak in order to evaluate different execution modes
  (feature model?)
  \item ``\emph{Things we tried, some improvement but had no significant impact;
  tradeoffs between accuracy and performance etc.}''
\end{itemize}

\section{How can we explain low code coverage?}
As already stated in the previous section, we encountered sevaral problems
during the development of the symbolic interpreter. Therefore this section
describes our methods to evaluate the precision of the output approximation, and
possible explanation theses.

\subsection{How to measure approximation success?} \label{heuristic}
In the absence of ground truth for the approximated output we
are bound to choose a \emph{heuristic approach} in order to measure the accuracy
of the aproximation. We define a string literal to be an \emph{output candidate} if there exists an
execution path reaching the string literal and passing it to an 
output-generating statement (e.g., \texttt{echo} or \texttt{print}).

We approximate output candidates as those string literals containing the
characters \texttt{<}, \texttt{>} or both. We evaluated this heuristic manually
with a sample of 400 string literals from the entire corpus and measured a
precision of 94\% and a recall of 50\%. This means, our classification has a
rate of 6\% false positives and only half of the actual output candidates
respond. These were mostly HTML-containing literals, whereas plain text did not
respond.

\subsection{Experimental Evaluation} 
For the evaluation we chose to selected a corpus of twelve PHP systems with regard to
previous work, but chose to additionally scale system size and recency. The full list of twelve PHP systems is shown in Table \ref{corpus}. Our selection contains
twelve systems:

\begin{itemize}
  \item Four small-scale/old systems are systems that have been subject to
  related studies and approaches in order to have comparable results. These
  projects do not make comprehensive use of PHP language features and output
  approximation on these systems achieved fairly high precision
  \cite{minamide2005static,Nguyen:2014:BCG:2635868.2635928}.

  \item Three large-scale/modern systems that have been subject to related
  studies, especially their use of language features as we aim to study
  scalability of symbolic execution The feature usage of these systems has been
  studied by Hills et al. \cite{Hills:2013:ESP:2483760.2483786}

  \item Five small-scale/modern systems to compare the impact
  of system size to the second part of the corpus. The systems are selected
  from a list of recent content management systems (CMS) \cite{codegeekz}. We only selected five systems 	due to limited support of the Quercus parser used in our implementation.
\end{itemize}

\begin{table*}[t]
\centering 
	\begin{tabular}{p{4cm}rrrl}
	\toprule
	\textbf{System} & \textbf{Version} & \textbf{File Count} & \textbf{SLOC} &
	\textbf{Description}
	\\
	\midrule
	AddressBook & 8.2.5.2 & 239 & 51,907 & Adress Administration \\
	SchoolMate & 1.5.4 & 65 & 8,118 & School Administration \\
	TimeClock & 1.04 & 63 & 20,800 & XXX \\
	WebChess & 1.0.0 & 28 & 5,219 & XXX \\
	\midrule
	Drupal & 7.5.0 & 125 & 52,464 & Content Management System \\
	phpBB & 3.1.9 & 1,398 & 327,371 & Bulletin Board \\
	phpMyAdmin & 4.6.3 & 871 & 303,582 & Database Administration \\
	\midrule
	Anchor & 0.12.1 & 201 & 15,054 & Content Management System \\
	Kirby & XXX & XXX & XXX & Content Management System \\
	Automad & XXX & XXX & XXX & Content Management System \\
	Monstra & XXX & XXX & XXX & Content Management System \\
	Nibbleblog & XXX & XXX & XXX & Content Management System \\
	\bottomrule
	\end{tabular}
	\caption{Corpus of twelve PHP systems. The file count includes files with a .php,
	.inc, .bit or .module extension.}
	\label{corpus}
\end{table*}

For the experimental analysis we symbolically executed per system each script file or entry point (files with a \texttt{.php}, \texttt{.inc}, \texttt{.bit} or \texttt{.module} extension). All following measurements are cumulated per system for each entry point.

\subsection{Metrics for Understanding Limitations} \label{metrics}
The following metrics are used throughout the description of the results.

\subsubsection{Reach Coverage and Output Coverage} \label{coverage_section}
In Section \ref{heuristic} we introduced the definition of output candidates. In
order to describe how much of the expected output was actually reached by the
symbolic interpreter or even was part of the output respectively.
We define the metric \emph{reach coverage} as the ratio of output candidates
that are reached by any execution path during symbolic execution and the total
number of output candidates in the analyzed system. Moreover, the metric \emph{output coverage} describes the ratio of output
candidates that are contained in the output of the symbolic interpreter and the total
number of output candidates of the analyzed system.

\subsubsection{Include Expressions: Coverage and Success} \label{include_coverage_section}
One metric to approach \emph{accuracy} is to measure (1) the ratio of reached
include expressions and the total number of include expressionsm in a system
respectively, and (2) the ratio of successfully resolved include expressions and
reached include expressions in a system. We therefore measure the coverage
of include expressions as well as the resolution success rate. For a most
accurate result, all include expressions are reached and resolved.

\subsubsection{String Literal Context} \label{context_section}
Another approach to understand not-reached output candidates is to understand why their surrounding program code was not accessed/accessible. Of special interest are those output candidates which are located in HTML template files or are surrounded by a function definition.

First, HTML templates are script files containing HTML and additional nested script snippets. Any HTML inside a template is part of the output once the template file is included. Consequently a not-reached output candidate residing in a template file means that the file is not included (dead code) or an include failed at runtime. Second, an output candidate located inside a function definition indicates that either the defined function is never called (dead code), or that the entire function definition is never reached, hence, it is not included or has failed to be included.

Apparently the explanations for both two cases are not mutually exclusive as the surrounding code does not give any reference to why it has not been accessed or could not be accessed. Nevertheless, this classification of not-reached output candidates may give us some insight on the question whether code accessibility is a matter, and if so, where to look for further answers. We will refer to this two cases as \emph{function context} and \emph{template context}.

\subsubsection{Dead Code Features} \label{dead_feature_section}
A refinement of the previous metric is to explicitly measure those cases where (1) not-reached output candidates are located inside a function context and the corresponding function is defined at runtime. Here we can a priori exclude a failed include from the set of possible explanations. We are also interested in those cases where (2) the not-reached output candidate is located in a file that has never been included from any other script file at runtime). Consequently these script files are either never to be included, or every attempt to include these files failed.

We refer to those two cases as dead code features, especially (1) \emph{dead functions} or (2) \emph{dead includes}. The explanations for these two cases are mutually exclusive since for (1) relevant script files need to be included and for (2) relevant script files must not be included.

\subsubsection{Callback Candidates} \label{callback_candidate_section}
Aside from failed or successful includes, this metric approaches the explanation of why functions containing not-reached output candidates are never called. The trivial explanation would be that there are no direct calls of those functions, yet PHP offers a number of ways to call a function indirectly, for instance using built-in functions (\texttt{call\_user\_func} or \texttt{call\_user\_func\_args}), or simply evaluating strings as PHP code using the \texttt{eval} function.

Moreover we need to take into account that one missed function execution can result in missing even more function calls (direct or indirect) and executions. Thus, for functions containing not-reached output candidates we measure whether they can be (1) only, (2) partially or (3) never accessed transitively from functions that have no direct call sites. We refer to those functions (or access points) having no direct call site as callback candidates since the only way to reach them is through indirect mechanisms.

Measuring how many not-reached output candidates can be explained by functions that are callback candidates helps to understand whether, and if so, indirect call mechanisms and their usage may impact our analysis' code coverage.

The callgraph snippets in Figure \ref{callgraph} illustrate the idea of detecting the functions' access points: The green function definitions depict callback candidates, the gray ones depict function definitions containing not-reached output candidates. In Figure \ref{callgraph_a} function \texttt{x()} can only be accessed through callback candidates, as for Figure \ref{callgraph_b} function \texttt{y()} can be partially accessed by callback candidates, whereas function \texttt{z()} is never accessible through callback candidates.

\begin{figure}
	\centering
	\begin{subfigure}[t]{0.18\textwidth}
	\centering
		\includegraphics[scale=0.4]{images-paper/callgraph_a}
		\caption{Completely explained by callback candidates (no call from
		\texttt{main})}
		\label{callgraph_a}
    \end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.25\textwidth}
	\centering
		\includegraphics[scale=0.4]{images-paper/callgraph_b}
		\caption{Partially (left; one call from \texttt{main}) or not at all  (right)
		explained by callback candidates}
		\label{callgraph_b}
    \end{subfigure}
    \caption{Callgraph analysis. The blue nodes represent entry points to
    relevant functions (gray).}
    \label{callgraph}
\end{figure}

\subsection{Results} \label{results}

% TODO Result description 
The coverage results are illustrated in Figure \ref{coverage}. We could replicate high coverage results for the four small-scale systems with a reach coverage and output coverage over 80~\% respectively. For the three modern/large-scale systems we measured poor coverage with reach coverage ranging from 5~\% to 30~\% and output coverage ranging from 4~\% to 20~\%. We measured medium to high coverage for the more recent small-scale systems: reach coverage ranging from 43~\% to 85~\%, output coverage ranging from 40~\% to 85~\%.

% RESULTS
% REACH COVERAGE and
% OUTPUT COVERAGE
\begin{figure}
	\centering	
	\input{diagrams/coverage}	
	\caption{Coverage results for the PHP corpus}
	\label{coverage}
\end{figure}

\subsubsection{Output Candidates in Function Definitions}
Based on these results we investigated why output candidates could not be reached. First, we started classifying not-reached output candidates by their string literal context (see Section \ref{context_section}). As the classification results in Figure \ref{contexts} illustrate, for almost all (except for AddressBook and TimeClock) systems not-reached output candidates had a function context. Note that SchoolMate was excluded from the diagram since our analysis reached all output candidates for that system.

% CONTEXT DISTRIBUTION
% - FUNCTION CONTEXT
% - TEMPLATE CONTEXT
\begin{figure}
	\input{diagrams/context_distribution}
	\caption{Context distribution of never reached output candidates for the PHP
	corpus.}
	\label{contexts}
\end{figure}

\subsubsection{Developing Hypotheses: Inaccessible Code}
Furthermore, this lead us to investigate whether we could explain not-reached output candidates by inaccessible dead code features (see Section \ref{dead_feature_section}), especially either dead functions or dead includes. As Figure \ref{dead_explanations} illustrates, the two explanation hypotheses did explain a portion of not-reached output candidates completely. If we look at Anchor for example, we see that all not-reached output candidates had a function context, and that 30~\% were explained by dead functions and dead includes respectively. The missing 40~\% correspond to functions that were defined and called, but whose function body was not completely covered by the execution, i.e., certain branches were missed.
Nevertheless, we see that inaccessible code is possibly one of the main reasons for missing output candidates.

% DEAD FUNCTIONS and DEAD INCLUDES
\begin{figure}
	\input{diagrams/dead_features}
	\caption{Explanation of never reached output candidates by dead code features.}
	\label{dead_explanations}
\end{figure}

\subsubsection{Why were code sections inaccessible?}  
Eventually we started analyzing (1) why include expression may not be resolved/evaluated correctly, and (2) why a reasonable portion of defined functions is never called. For include expressions we  sampled failed attempts and manually evaluated the cause of failure.

We identified for (1) that all failed attempts were dynamic include expressions, i.e., those which are resolved at runtime and were assembled from fragments of environment-dependent information including database retrievals, user input, or files. Consequently, the resolved include expressions contained symbolic values which apparently did not represent identifiable script files.

For (2) we again sampled dead function cases and identified the lack of direct calls to those functions. Interestingly, also functions despite having no direct call sites were called through indirect call features (see Section \ref{callback_candidate_section}). Again we identified dynamically assembled expressions (function names) to be responsible for failed attempts of calling a function indirectly as the resolved function name eventually contained symbolic values. These expressions, as well as include expressions, contained information dependent on the system environment.

\subsubsection{Quantitative Analysis}
Finally, we measured for include expressions the overall reach coverage and the overall resolution success rate per system. As illustrated in Figure \ref{include_coverage_results1}, except for the small-scale systems and two recent small-scale systems we did not even really reach a great portion of include expressions, although for most systems the resolution success rate was over 80~\% (see Figure \ref{include_coverage_results2}). As described in Section \ref{include_coverage_section} one failed attempt to include a script file possibly results in a cascade of missing more include expressions.

% INCLUDE COVERAGE / SUCCESS RATE
\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.5 \textwidth}    	
        \centering        
        % Include Coverage Plot        
		\input{diagrams/include_coverage}		
        \caption{Proportion of reached include expressions per project.}
        \label{include_coverage_results1}
    \end{subfigure}%
	\hfill
    \begin{subfigure}[t]{0.5\textwidth}    
        \centering
        % Success Rate Plot        
        \input{diagrams/success_rate}        
        \caption{Rate of successful resolutions of reached include expressions
        per project.}
        \label{include_coverage_results2}
    \end{subfigure}
    \label{include_coverage_results}
    \caption{Include coverage and accuracy.}
\end{figure*}

For failed function calls to functions containing not-reached output candidates we measured how many of those functions were only, or partially accessibly through indirect calling mechanisms (see Section \ref{callback_candidate_section}). As Figure \ref{output_candidate_explanation} illustrates, around 80~\% (except for TimeClock) all missed output candidates of each system were only accessible through callback candidates, i.e., functions that are only accessible through indirect calling mechanisms.

% CALLBACK CANDIDATE ACCESSIBILITY ANALYSIS
\begin{figure}
	\input{diagrams/output_candidates}
	\caption{Distribution dead function output candidates that are partially or
	completely explained due to callback candidates.}
	\label{output_candidate_explanation}
\end{figure}

\section{Lessons Learned}
\begin{itemize}
  \item Certain language features are prone to be used in a way that symbolic
  execution is impeded.
  \item Tendency to use functions/includes with I/O, databases or configuration
  files is problematic.
\end{itemize}

\begin{framed}
\textcolor{RYB1}{The use of include expressions whose evaluation at runtime
involves concatenation, especially of fragments from critical sources, impede
symbolic execution due to inaccuracy.}
\end{framed}

\section{Related Work}
\begin{itemize}
  \item Static output approximation using context-free grammars
  by Minamide \cite{minamide2005static} and detection of SQL
  injection vulnerabilities \cite{wassermann2007sound}.
  \item Static analysis of feature usage in PHP systems by Hills et al.
  \cite{Hills:2013:ESP:2483760.2483786} and static resolution of include
  expressions \cite{Hills:2013:ESP:2483760.2483786,hills2014static}
\end{itemize}

\newpage
\bibliographystyle{abbrv}
\bibliography{bibliography}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
\end{document}

### 2016-12-08 paper.tex
08.12.16

@@
 \alignauthor
-%Christian Kästner\\
-%      \affaddr{Carnegie Mellon University, USA}\\
+Christian Kästner\\
+      \affaddr{Carnegie Mellon University, USA}\\
        %\email{ckaestner@cs.cmu.edu}
@@
 \alignauthor
-%Tien N. Nguyen\\
-%       \affaddr{University of Texas at Dallas, USA}\\
+Tien N. Nguyen\\
+       \affaddr{University of Texas at Dallas, USA}\\
 %      %\email{tien.n.nguyen@utdallas.edu}
@@
 corresponding programming patterns that impede symbolic execution and limit the
-scalability of this approach. Our findings include: (1) dynamic features such as
+scalability of this approach. Our findings include: (1) Dynamic features such as
 functions and includes are prone to fail for certain programming patterns. (2)
@@
 \section{Introduction}
-Dynamic web applications have become widely popular and various different
-implementation techniques have emerged, ranging from server-side scripting
-languages to state-of-the-art frameworks for web development. For scritpting
-languages such as PHP, source code usually contains script code along with nested
-target code since the HTML pages are dynamically generated.
+Dynamic web applications have become widely popular and various different implementation techniques have emerged, ranging from server-side scripting languages to state-of-the-art frameworks for web development.
+For all  ways to implement web applications, one common ground is the dynamic nature of output generation: HTML output is generated dynamically during a session, for instance when a form is submitted to a web application.
+Output can be dependent on different classes of inputs.
+
+Symbolic execution is a static analysis technique between program testing and program proving.
+For a program, symbolic, i.e., abstract, yet fix inputs are applied instead of  concrete inputs \cite{Darringer1978,King1976}.
+The underlying concept is to map program input (symbolically) to program output.
+Different classes of inputs may result in variational output, representing  dependencies of output and input.
+Symbolic execution extends concrete execution semantics with semantics to handle symbolic values.
+Hence, concrete execution semantics is contained in symbolic execution semantics.
+
+Symbolic execution has been used and studied for generating test suites, but has more recently also been applied for approximating the output of web applications.
+Knowledge about different output variants enabled useful analyses for the domain of web applications, such as detecting and locating HTML validation errors \cite{Nguyen:2011:AFH:2190078.2190142}.
+As for developers, from an output model, outlines and callgraphs for easier IDE code navigation \cite{Nguyen:2014:BCG:2635868.2635928}, or program slices \cite{Nguyen:2015:CPS:2786805.2786872} can be computed.
+Existing work has been integrated in the Varis plugin for the Eclipse IDE \cite{Nguyen:2015:VIS:2819009.2819140}.
+Despite the various use cases for analyses based on an approximated output model, static output approximation for PHP web applications has so far only been tested for smaller systems that are not developed anymore.
+
+\subsection*{Can we have a practical tool?}
+We ask for the symbolic execution engine whether we can have a practical tool built from it.
+In the context of useful analyses stated earlier, we define a practical symbolic engine to be accurate and precise enough for the analyses' results to be useful.
+An accurate symbolic execution engine allows a single concrete execution to be retraced from a symbolic execution representing variability.
+For a precise symbolic execution engine, each concrete execution represented in a symbolic execution could have been performed by a concrete execution engine.
+
+We see concerns for practicality in intrinsic properties of symbolic execution (abstraction of real input sources), and in the language properties of PHP.
+With no standardization the language has become successful one the one hand, but did not grew coordinately leaving constructs that are partially  functionally redundant, or inconsistent for different versions.
+
+\subsection*{We can have a practical tool, can't we?}
+To investigate this question, we build a symbolic execution engine for PHP that re-implements the specifications of previous symbolic execution semantics \cite{Nguyen:2014:BCG:2635868.2635928} for PHP with additional features, such as object-oriented programming, and extensive array support.
+
+Based on our observations and evaluation results, we learn that for dynamic PHP constructs, expressions that are assembled dynamically, in combination with symbolic values symbolic execution can be impractical.
+For object-orientation and method calls, symbolic execution semantics can be a tightrope walk between state space explosion and imprecision.
+Moreover, the nature of PHP challenges a practical symbolic execution engine since  for it to be accurate, a large number of functions from the standard library needs to be supported.
+
+Our key contributions in this paper include: (1) Ihe tool infrastructure to statically approximate the output for PHP web applications using symbolic execution and (2) observations as well as an anecdotal and empirical evaluation to identify and measure PHP constructs impeding symbolic execution of PHP code.
 
@@
 	\centering	
+	\resizebox{0.4\textwidth}{!}{%
 		\input{diagrams/coverage}	
+	}
 	\caption{Coverage results for the PHP corpus}
@@
 \begin{figure}
+	\resizebox{0.4\textwidth}{!}{%
 		\input{diagrams/context_distribution}
+	}	
 	\caption{Context distribution of never reached output candidates for the PHP
@@
   expressions \cite{Hills:2013:ESP:2483760.2483786,hills2014static}
+  \item bener \cite{King1976} \cite{multiSE}
 \end{itemize}

### 2016-12-19 paper.tex
19.1

@@
-\documentclass{sig-alternate-05-2015}
+\documentclass[preprint]{sig-alternate-05-2015}
 
@@
 
+\usepackage{enumerate}
 \usepackage{subcaption}
@@
 for vulnerability detection. Unfortunately, recent approximation approaches have
-only been evaluated for smaller systems due to both conceptual limitations and
-unsupported language features.
+only been evaluated for smaller systems.
 
@@
 \section{Introduction}
-Dynamic web applications have become widely popular and various different implementation techniques have emerged, ranging from server-side scripting languages to state-of-the-art frameworks for web development.
-For all  ways to implement web applications, one common ground is the dynamic nature of output generation: HTML output is generated dynamically during a session, for instance when a form is submitted to a web application.
-Output can be dependent on different classes of inputs.
+With the world wide web emerging, dynamic web applications have become widely popular.
+Various different implementation techniques have emerged, ranging from web application frameworks for programming (JSP, ASP .NET) and script languages (Ruby On Rails, Django) to languages tailored specifically to the domain of web applications.
+PHP \cite{phpNET} is a programming language focused on server-side application development.
+For 2012, it was used by 78.8 percent of the ten million most popular websites (according to Alexa popularity ranking) \cite{alexaPHP}, ranking 7th on the TIOBE programming community index \cite{tiobePHP}, and was the ranked as the 6th most popular language on GitHub \cite{githubPHP}.
+
+For all ways to implement web applications, one common ground is the nature of output generation: HTML output is generated dynamically during a session, for instance when a form is submitted to a web application.
+Thus, output can be dependent on different classes of inputs.
 
@@
 during the development of the symbolic interpreter. Therefore this section
-describes our methods to evaluate the precision of the output approximation, and
+describes our methods to evaluate practicality of the output approximation, and
 possible explanation theses.
@@
 \subsection{How to measure approximation success?} \label{heuristic}
-In the absence of ground truth for the approximated output we
-are bound to choose a \emph{heuristic approach} in order to measure the accuracy
-of the aproximation. We define a string literal to be an \emph{output candidate} if there exists an
-execution path reaching the string literal and passing it to an 
+For the evaluation of our output approximation we require a ground truth output model to compare against.
+In the absence of ground truth for the approximated output we are bound to choose a \emph{heuristic approach} in order to measure the accuracy of our aproximation.
+We define a string literal to be an \emph{output candidate} if there exists an execution path reaching the string literal and evantually passing it to an
 output-generating statement (e.g., \texttt{echo} or \texttt{print}).
 
-We approximate output candidates as those string literals containing the
-characters \texttt{<}, \texttt{>} or both. We evaluated this heuristic manually
-with a sample of 400 string literals from the entire corpus and measured a
-precision of 94\% and a recall of 50\%. This means, our classification has a
-rate of 6\% false positives and only half of the actual output candidates
-respond. These were mostly HTML-containing literals, whereas plain text did not
-respond.
+We approximate output candidates as those string literals containing the characters \texttt{<}, \texttt{>}, or both since we expect output to contain HTML tags. We evaluated this heuristic manually
+with a sample of 400 string literals randomly selected from the entire corpus.
+We measured for our heuristic classifier a precision of 94\%, and a recall of 50\%.
+
+This means that six percent of the string literals are classified incorrectly as false positives.
+In turn, the classifier is highly distinctive as 96 percent of the string literals are classified correctly.
+The recall of 50 percent means that half of the string literals, which we manually classified as output, actually were responsive to the classifier.
+We attempted to increase recall by looking for further distinctive properties to build a classifier from, but could not do so without decreasing precision.
+In spite of missing half of the output candidates, we decided to use a simple, yet distinctive classifier.
 
 \subsection{Experimental Evaluation} 
-For the evaluation we chose to selected a corpus of twelve PHP systems with regard to
-previous work, but chose to additionally scale system size and recency. The full list of twelve PHP systems is shown in Table \ref{corpus}. Our selection contains
-twelve systems:
-
-\begin{itemize}
-  \item Four small-scale/old systems are systems that have been subject to
-  related studies and approaches in order to have comparable results. These
-  projects do not make comprehensive use of PHP language features and output
-  approximation on these systems achieved fairly high precision
-  \cite{minamide2005static,Nguyen:2014:BCG:2635868.2635928}.
-
-  \item Three large-scale/modern systems that have been subject to related
-  studies, especially their use of language features as we aim to study
-  scalability of symbolic execution The feature usage of these systems has been
-  studied by Hills et al. \cite{Hills:2013:ESP:2483760.2483786}
-
-  \item Five small-scale/modern systems to compare the impact
-  of system size to the second part of the corpus. The systems are selected
-  from a list of recent content management systems (CMS) \cite{codegeekz}. We only selected five systems 	due to limited support of the Quercus parser used in our implementation.
-\end{itemize}
+For the evaluation we symbolically executed a selection of PHP systems.
+We selected a corpus of twelve PHP systems with regard to system size and recency as the corpus of the case study for \cite{Nguyen:2014:BCG:2635868.2635928} only contained small-scale systems that are not maintained any more.
+The full list of PHP systems is shown in Table \ref{corpus}.
+Our selection of PHP systems includes
+\begin{enumerate}[a)]
+\item four small-scale systems that selected from the previously mentioned case study in order to compare our results with the previous symbolic execution engine \cite{Nguyen:2014:BCG:2635868.2635928},
+\item three recent large-scale systems selected from the case study corpus of \cite{Hills:2013:ESP:2483760.2483786}, and
+\item four small-scale systems selected from a list of recent Content Management Systems \cite{codegeekz}.
+\end{enumerate}
+As for b) and c), we limited our selection since the parser we used did not support all language features used.
 

### 2016-12-22 paper.tex
21.12.2016

@@
 \usepackage{booktabs}
+\usepackage{paralist}
 
@@
 
+\lstset{
+language=PHP,
+basicstyle=\small\sffamily,
+numbers=left,
+numberstyle=\tiny,
+frame=tb,
+columns=fullflexible,
+showstringspaces=false
+}
+
 % ARIAL
@@
 
-\definecolor{RYB1}{HTML}{8F83DB}
-\definecolor{RYB2}{HTML}{DDDDDD}
-\definecolor{RYB3}{HTML}{FFFFFF}
+\definecolor{RYB1}{HTML}{DCDCDC}
+\definecolor{RYB2}{HTML}{C0C0C0}
+\definecolor{RYB3}{HTML}{808080}
 
-\definecolor{RYB4}{HTML}{A699FF}
-\definecolor{RYB5}{HTML}{FFBA49}
-\definecolor{RYB6}{HTML}{FFFFFF}
+\definecolor{RYB4}{HTML}{DCDCDC}
+\definecolor{RYB5}{HTML}{C0C0C0}
+\definecolor{RYB6}{HTML}{808080}
 
-\definecolor{RYB7}{HTML}{6A62A3}
-\definecolor{RYB8}{HTML}{BEB4FF}
-\definecolor{RYB9}{HTML}{E6E3FF}
+\definecolor{RYB7}{HTML}{DCDCDC}
+\definecolor{RYB8}{HTML}{C0C0C0}
+\definecolor{RYB9}{HTML}{808080}
 
-\definecolor{RYB10}{HTML}{FFB030}
-\definecolor{RYB11}{HTML}{FFCD7C}
+\definecolor{RYB10}{HTML}{D3D3D3}
+\definecolor{RYB11}{HTML}{A9A9A9}
 
@@
 Tien N. Nguyen\\
-       \affaddr{University of Texas at Dallas, USA}\\
+       \affaddr{University of Texas, USA}\\
 %      %\email{tien.n.nguyen@utdallas.edu}
@@
 \begin{abstract}
-Dynamic web applications have become widely popular and are to a large
-proportion based on the scripting language PHP. Output approximation of web
+Dynamic web applications have become widely popular and are to
+a large proportion based on the scripting language PHP. Output approximation of web
 applications enables a range of additional tool support as well as possibilities
@@
 applications.
-
 \end{abstract}
@@
 \section{Introduction}
-With the world wide web emerging, dynamic web applications have become widely popular.
-Various different implementation techniques have emerged, ranging from web application frameworks for programming (JSP, ASP .NET) and script languages (Ruby On Rails, Django) to languages tailored specifically to the domain of web applications.
-PHP \cite{phpNET} is a programming language focused on server-side application development.
-For 2012, it was used by 78.8 percent of the ten million most popular websites (according to Alexa popularity ranking) \cite{alexaPHP}, ranking 7th on the TIOBE programming community index \cite{tiobePHP}, and was the ranked as the 6th most popular language on GitHub \cite{githubPHP}.
+With the emerge of the world wide web, dynamic web applications have become
+widely popular. Various different implementation techniques have emerged,
+ranging from technologies for programming languages (JSP, ASP .NET) and web
+application frameworks for  script languages (Ruby On Rails, Django) to
+languages tailored specifically to the domain of web applications. PHP \cite{phpNET} is
+a programming language focused on server-side application development. As of 2012,
+it was used by 78.8 percent of the ten million most popular websites (according
+to Alexa popularity ranking) \cite{alexaPHP}, ranking 7th on the TIOBE
+programming community index \cite{tiobePHP}, and was the ranked as the 6th most
+popular language on GitHub \cite{githubPHP}.\\
 
-For all ways to implement web applications, one common ground is the nature of output generation: HTML output is generated dynamically during a session, for instance when a form is submitted to a web application.
-Thus, output can be dependent on different classes of inputs.
+One common property of all technologies for dynamic we b applications is the
+\emph{staged computation} of output, or simplified: dynamic code generation.
+Since based on both static values and runtime values, output code is generated step by step
+dynamically, output can be dependent on application inputs. Dynamic web
+applications incorporate various both server-side inputs (databases,
+configuration) and client-side inputs respectively. Static
+analysis of dynamic web applications, hence, has to consider different input
+sources.
 
-Symbolic execution is a static analysis technique between program testing and program proving.
-For a program, symbolic, i.e., abstract, yet fix inputs are applied instead of  concrete inputs \cite{Darringer1978,King1976}.
-The underlying concept is to map program input (symbolically) to program output.
-Different classes of inputs may result in variational output, representing  dependencies of output and input.
-Symbolic execution extends concrete execution semantics with semantics to handle symbolic values.
-Hence, concrete execution semantics is contained in symbolic execution semantics.
+\emph{Symbolic execution} is a static analysis technique between program testing
+and program proving. For a program, symbolic, i.e., abstract, yet fix inputs are
+applied instead of concrete inputs \cite{Darringer1978,King1976}. The underlying
+concept is to map program input (symbolically) to program output. Different classes of inputs may
+result in variational output, representing dependencies of output and input.
+Symbolic execution extends concrete execution semantics with semantics to
+handle symbolic values. Hence, concrete execution semantics is contained in
+symbolic execution semantics.\\
 
-Symbolic execution has been used and studied for generating test suites, but has more recently also been applied for approximating the output of web applications.
-Knowledge about different output variants enabled useful analyses for the domain of web applications, such as detecting and locating HTML validation errors \cite{Nguyen:2011:AFH:2190078.2190142}.
-As for developers, from an output model, outlines and callgraphs for easier IDE code navigation \cite{Nguyen:2014:BCG:2635868.2635928}, or program slices \cite{Nguyen:2015:CPS:2786805.2786872} can be computed.
-Existing work has been integrated in the Varis plugin for the Eclipse IDE \cite{Nguyen:2015:VIS:2819009.2819140}.
-Despite the various use cases for analyses based on an approximated output model, static output approximation for PHP web applications has so far only been tested for smaller systems that are not developed anymore.
+Symbolic execution has been used and studied for generating test suites, but
+has more recently also been applied for approximating the output of web
+applications. Knowledge about different output variants enabled useful analyses
+for the domain of web applications, such as detecting and locating HTML
+validation errors \cite{Nguyen:2011:AFH:2190078.2190142}. As for developers,
+from an output model, outlines and callgraphs for easier IDE code navigation
+\cite{Nguyen:2014:BCG:2635868.2635928}, or program slices
+\cite{Nguyen:2015:CPS:2786805.2786872} can be computed. Existing work has been
+integrated in the Varis plugin for the Eclipse IDE \cite{Nguyen:2015:VIS:2819009.2819140}.
 
-\subsection*{Can we have a practical tool?}
-We ask for the symbolic execution engine whether we can have a practical tool built from it.
-In the context of useful analyses stated earlier, we define a practical symbolic engine to be accurate and precise enough for the analyses' results to be useful.
-An accurate symbolic execution engine allows a single concrete execution to be retraced from a symbolic execution representing variability.
-For a precise symbolic execution engine, each concrete execution represented in a symbolic execution could have been performed by a concrete execution engine.
+Despite the various use cases for analyses based on an approximated output
+model, static output approximation for PHP web applications using symbolic
+execution has so far only been evaluated for smaller systems that are not
+developed anymore. The previously mentioned tools though are only practical, if
+for a given system the symbolic execution engine is scalable, i.e., will also
+approximate reliable output for larger and more recent systems with good time
+and space consumption.\\
 
-We see concerns for practicality in intrinsic properties of symbolic execution (abstraction of real input sources), and in the language properties of PHP.
-With no standardization the language has become successful one the one hand, but did not grew coordinately leaving constructs that are partially  functionally redundant, or inconsistent for different versions.
+To investigate the question, whether we can have practical tools built from the
+symbolic execution engine, we re-implemented the engine with the specifications
+of the previous symbolic execution semantics
+\cite{Nguyen:2014:BCG:2635868.2635928} for PHP and additional features, such as
+object-oriented programming, and extensive array support. We evaluated our
+symbolic execution engine for large-scale and modern PHP systems.
 
-\subsection*{We can have a practical tool, can't we?}
-To investigate this question, we build a symbolic execution engine for PHP that re-implements the specifications of previous symbolic execution semantics \cite{Nguyen:2014:BCG:2635868.2635928} for PHP with additional features, such as object-oriented programming, and extensive array support.
+During introduction of new semantics for object-orientation and advanced array
+we encountered two main challenges. First, we walked a boundary between a
+purely symbolic or mixed concrete and symbolic (concolic) execution engine
+since concrete execution, e.g. foreach loops for arrays, increased code
+coverage significantly. Second, we had to make a trade-off between the accuracy
+of method calls and the scalability of this language feature as a method call
+for different program state variants can have multiple targets.
 
-Based on our observations and evaluation results, we learn that for dynamic PHP constructs, expressions that are assembled dynamically, in combination with symbolic values symbolic execution can be impractical.
-For object-orientation and method calls, symbolic execution semantics can be a tightrope walk between state space explosion and imprecision.
-Moreover, the nature of PHP challenges a practical symbolic execution engine since  for it to be accurate, a large number of functions from the standard library needs to be supported.
+Based on our observations and evaluation results, we learned that for dynamic
+PHP constructs, expressions that are assembled dynamically, in combination with
+symbolic values symbolic execution can be impractical. For object-orientation
+and method calls, symbolic execution semantics can be a tightrope walk between
+state space explosion and imprecision. Moreover, the nature of PHP challenges a
+practical symbolic execution engine since for it to be accurate, a large number
+of functions from the standard library needs to be supported.\\
 
-Our key contributions in this paper include: (1) Ihe tool infrastructure to statically approximate the output for PHP web applications using symbolic execution and (2) observations as well as an anecdotal and empirical evaluation to identify and measure PHP constructs impeding symbolic execution of PHP code.
+Our key contributions in this paper include: (1) The tool infrastructure to
+statically approximate the output for PHP web applications using symbolic
+execution and (2) observations as well as an experience evaluation to identify
+and measure PHP constructs impeding symbolic execution of PHP code.
 
 \section{State Of the Art}
-\subsection{Symbolic Execution (and Choices?)}
-\subsection{Use Cases for Output Approximation}
 
-\begin{figure}
+\section{Towards scalable symbolic execution}%
 
-	\centering
-	\includegraphics[width=0.5\textwidth]{images-paper/approach}
-	\caption{Approach overview; the different analyses in Varis are based upon the
-	symbolic output. Step 1 is the \emph{symbolic execution}, step 2 is the
-	\emph{variability-aware parsing} and step enfolds subsequent analyses based
-	upon the symbolic output.}
-	\label{approach}
-\end{figure}
-
-\begin{itemize}
-  \item Steps: Symbolic Execution $\rightarrow$ Parsing $\rightarrow$
-  Analysis, see Figure \ref{approach}
-  \item Varis: IDE support \cite{Nguyen:2015:VIS:2819009.2819140} for PHP with
-  code slicing \cite{Nguyen:2015:CPS:2786805.2786872}, code navigation
-  \cite{Nguyen:2014:BCG:2635868.2635928} and
-  HTML validation \cite{Nguyen:2011:AFH:2190078.2190142}.
-\end{itemize}
-
-\section{Towards scalable symbolic execution}
-\subsection{Can we have a practical tool?}
-\begin{itemize}
-  \item Are the assumprions just engineering or do not matter for accuracy? 
-\end{itemize}
-\subsection{Oak: A Symbolic Interpreter}
-\begin{itemize}
-  \item \textsc{Oak}\footnote{\url{https://github.com/smba/oak}}:
-  Re-implementing and extending the symbolic execution engine with OOP and array support
-  \item Test-driven implementation strategy 
-  \item WordPress!
-\end{itemize}
-\subsection{Experience Report}
-\subsubsection{Problems}
-\subsubsection{Concrete/Symbolic Boundary}
-\subsubsection{Further experiments}
-\begin{itemize}
-  \item SPL approach for Oak in order to evaluate different execution modes
-  (feature model?)
-  \item ``\emph{Things we tried, some improvement but had no significant impact;
-  tradeoffs between accuracy and performance etc.}''
-\end{itemize}
-
-\section{How can we explain low code coverage?}
+\section{Evaluation}
 As already stated in the previous section, we encountered sevaral problems
@@
 describes our methods to evaluate practicality of the output approximation, and
-possible explanation theses.
+possible explanations.
 
-\subsection{How to measure approximation success?} \label{heuristic}
+\subsection{Measuring Approximation Success} \label{heuristic}
 For the evaluation of our output approximation we require a ground truth output model to compare against.
@@
 with a sample of 400 string literals randomly selected from the entire corpus.
-We measured for our heuristic classifier a precision of 94\%, and a recall of 50\%.
-
+We measured for our heuristic classifier a precision of 94 percent, and a recall
+of 50 percent.
 This means that six percent of the string literals are classified incorrectly as false positives.
@@
 
-\subsection{Experimental Evaluation} 
+\subsection{Experiment Setup} 
 For the evaluation we symbolically executed a selection of PHP systems.
@@
 Our selection of PHP systems includes
-\begin{enumerate}[a)]
+%\begin{enumerate}[a)]
+\begin{compactitem}
 \item four small-scale systems that selected from the previously mentioned case study in order to compare our results with the previous symbolic execution engine \cite{Nguyen:2014:BCG:2635868.2635928},
@@
 \item four small-scale systems selected from a list of recent Content Management Systems \cite{codegeekz}.
-\end{enumerate}
-As for b) and c), we limited our selection since the parser we used did not support all language features used.
+%umerate}
+
+\end{compactitem}
+As for for the last two cases, we limited our selection since the parser we
+used did not support all language features used.
 
@@
 
-\subsection{Metrics for Understanding Limitations} \label{metrics}
-The following metrics are used throughout the description of the results.
+\subsection{Measuring Approximation Accuracy}
+\label{HowAccurateIsOurApproximation} 
+In Section \ref{heuristic} we introduced the definition of output candidates as expected output.
+As a most accurate approximation contains all output candidates, we define two metrics to measure accuracy of our approximation.
 
-\subsubsection{Reach Coverage and Output Coverage} \label{coverage_section}
-In Section \ref{heuristic} we introduced the definition of output candidates. In
-order to describe how much of the expected output was actually reached by the
-symbolic interpreter or even was part of the output respectively.
-We define the metric \emph{reach coverage} as the ratio of output candidates
-that are reached by any execution path during symbolic execution and the total
-number of output candidates in the analyzed system. Moreover, the metric \emph{output coverage} describes the ratio of output
-candidates that are contained in the output of the symbolic interpreter and the total
-number of output candidates of the analyzed system.
+First, we measure how much of the expected output was actually processed by the symbolic interpreter.
+Once a line of code, a statement or expression containing an output candidate is actually an element in an execution path, we define this output candidate as reached.
+We define the metric \emph{reach coverage} as the ratio of output candidates that are reached and the total number of output candidates in the analyzed system.
+Although a reached output candidate is processed by the symbolic interpreter, it does not guarantee we will see that output candidate in the symbolic output.
+Second, we define the metric \emph{output coverage} as the ratio of output candidates that are contained in the output of the symbolic interpreter and the total number of output candidates in the analyzed system.
+For an imprecise approximation, we may see loss of output candidate information resulting in a output coverage lower than the reach coverage.
 
-\subsubsection{Include Expressions: Coverage and Success} \label{include_coverage_section}
-One metric to approach \emph{accuracy} is to measure (1) the ratio of reached
-include expressions and the total number of include expressionsm in a system
-respectively, and (2) the ratio of successfully resolved include expressions and
-reached include expressions in a system. We therefore measure the coverage
-of include expressions as well as the resolution success rate. For a most
-accurate result, all include expressions are reached and resolved.
-
-\subsubsection{String Literal Context} \label{context_section}
-Another approach to understand not-reached output candidates is to understand why their surrounding program code was not accessed/accessible.
-Of special interest are those output candidates which are located in HTML template files or are surrounded by a function definition.
-
-First, HTML templates are script files containing HTML and additional nested script snippets.
-Any HTML inside a template is part of the output once the template file is included.
-Consequently a not-reached output candidate residing in a template file means that the file is not included (dead code) or an include failed at runtime.
-Second, an output candidate located inside a function definition indicates that either the defined function is never called (dead code), or that the entire function definition is never reached, hence, it is not included or has failed to be included.
-
-Apparently the explanations for both two cases are not mutually exclusive as the surrounding code does not give any reference to why it has not been accessed or could not be accessed.
-Nevertheless, this classification of not-reached output candidates may give us some insight on the question whether code accessibility is a matter, and if so, where to look for further answers.
-We will refer to this two cases as \emph{function context} and \emph{template context}.
-
-\subsubsection{Dead Code Features} \label{dead_feature_section}
-A refinement of the previous metric is to explicitly measure those cases where (1) not-reached output candidates are located inside a function context and the corresponding function is defined at runtime.
-Here we can a priori exclude a failed include from the set of possible explanations.
-We are also interested in those cases where (2) the not-reached output candidate is located in a file that has never been included from any other script file at runtime).
-Consequently these script files are either never to be included, or every attempt to include these files failed.
-
-We refer to those two cases as dead code features, especially (1) \emph{dead functions} or (2) \emph{dead includes}.
-The explanations for these two cases are mutually exclusive since for (1) relevant script files need to be included and for (2) relevant script files must not be included.
-
-\subsubsection{Callback Candidates} \label{callback_candidate_section}
-Aside from failed or successful includes, this metric approaches the explanation of why functions containing not-reached output candidates are never called.
-The trivial explanation would be that there are no direct calls of those functions, yet PHP offers a number of ways to call a function indirectly, for instance using built-in functions (\texttt{call\_user\_func} or \texttt{call\_user\_func\_args}), or simply evaluating strings as PHP code using the \texttt{eval} function.
-
-Moreover we need to take into account that one missed function execution can result in missing even more function calls (direct or indirect) and executions.
-Thus, for functions containing not-reached output candidates we measure whether they can be (1) only, (2) partially or (3) never accessed transitively from functions that have no direct call sites.
-We refer to those functions (or access points) having no direct call site as callback candidates since the only way to reach them is through indirect mechanisms.
-
-Measuring how many not-reached output candidates can be explained by functions that are callback candidates helps to understand whether, and if so, indirect call mechanisms and their usage may impact our analysis' code coverage.
-
-The callgraph snippets in Figure \ref{callgraph} illustrate the idea of detecting the functions' access points: The green function definitions depict callback candidates, the gray ones depict function definitions containing not-reached output candidates.
-In Figure \ref{callgraph_a} function \texttt{x()} can only be accessed through callback candidates, as for Figure \ref{callgraph_b} function \texttt{y()} can be partially accessed by callback candidates, whereas function \texttt{z()} is never accessible through callback candidates.
-
-\begin{figure}
-	\centering
-	\begin{subfigure}[t]{0.18\textwidth}
-	\centering
-		\includegraphics[scale=0.4]{images-paper/callgraph_a}
-		\caption{Completely explained by callback candidates (no call from
-		\texttt{main})}
-		\label{callgraph_a}
-    \end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.25\textwidth}
-	\centering
-		\includegraphics[scale=0.4]{images-paper/callgraph_b}
-		\caption{Partially (left; one call from \texttt{main}) or not at all  (right)
-		explained by callback candidates}
-		\label{callgraph_b}
-    \end{subfigure}
-    \caption{Callgraph analysis. The blue nodes represent entry points to
-    relevant functions (gray).}
-    \label{callgraph}
-\end{figure}
-
-\subsection{Results} \label{results}
-
-% TODO Result description 
 The coverage results are illustrated in Figure \ref{coverage}.
-We could replicate high coverage results for the four small-scale systems with a reach coverage and output coverage over 80~\% respectively.
-For the three modern/large-scale systems we measured poor coverage with reach coverage ranging from 5~\% to 30~\% and output coverage ranging from 4~\% to 20~\%.
-We measured medium to high coverage for the more recent small-scale systems: reach coverage ranging from 43~\% to 85~\%, output coverage ranging from 40~\% to 85~\%.
-
+We could replicate high coverage results for the four small-scale systems with a reach coverage and output coverage over 80 percent respectively.
+For the three modern/large-scale systems we measured poor coverage with reach coverage ranging from 5 to 30 percent and output coverage ranging from 4 to 20 percent.
+We measured medium to high coverage for the more recent small-scale systems: reach coverage ranging from 43 to 85 percent, output coverage ranging from 40 to 85 percent.
 % RESULTS
@@
 
-\subsubsection{Output Candidates in Function Definitions}
-Based on these results we investigated why output candidates could not be reached.
-First, we started classifying not-reached output candidates by their string literal context (see Section \ref{context_section}).
-As the classification results in Figure \ref{contexts} illustrate, for almost all (except for AddressBook and TimeClock) systems not-reached output candidates had a function context.
-Note that SchoolMate was excluded from the diagram since our analysis reached all output candidates for that system.
+\subsection{Understanding Limitations}
+As we have seen in Section \ref{HowAccurateIsOurApproximation} both measured coverage metrics were poor for large-scale and small-scale/recent systems.
+To understand what output candidates we missed and why, we further investigated our approximation results and conducted three more measurements.
+
+\subsubsection{What literals did we miss?}
+\label{WhatLiteralsDidWeMiss?}
+Our initial approach to understand missed output candidates is to find out whether their surrounding program code was not accessed/accessible, and if so, why.
+Of special interest are those output candidates that are located in HTML files (possibly with nested PHP), or located in a function definition.
+Plain HTML files do not require any further execution (of course, unless PHP scripts are embedded) and represent output candidates that just need to be included properly to be part of the output.
+Aside, output candidates that are part of a function definition are only missed if the corresponding function is never called at runtime.
+
+We started classifying missed output candidates by their string literal context.
+As the classification statistics in Figure \ref{fig:contexts} illustrate, for almost all systems (except for \textsc{AddressBook} and \textsc{TimeClock}) not-reached output candidates had a function context.
+Note that \textsc{SchoolMate} was excluded from the diagram since our analysis reached all output candidates for that system.
 
@@
 	corpus.}
-	\label{contexts}
+	\label{fig:contexts}
 \end{figure}
 
-\subsubsection{Developing Hypotheses: Inaccessible Code}
-Furthermore, this lead us to investigate whether we could explain not-reached output candidates by inaccessible dead code features (see Section \ref{dead_feature_section}), especially either dead functions or dead includes.
-As Figure \ref{dead_explanations} illustrates, the two explanation hypotheses did explain a portion of not-reached output candidates completely.
-If we look at Anchor for example, we see that all not-reached output candidates had a function context, and that 30~\% were explained by dead functions and dead includes respectively.
-The missing 40~\% correspond to functions that were defined and called, but whose function body was not completely covered by the execution, i.e., certain branches were missed.
-Nevertheless, we see that inaccessible code is possibly one of the main reasons for missing output candidates.
+\subsubsection{Inaccessible Dynamic Features}
+\label{sec:inaccessible}
+For both cases of context, either inclusion of a script file failed, or a function call failed.
+If an HTML file is not part of the output, it also is never reached, i.e., included.
+An include can fail due to various reasons, such as an imprecise evaluation of the include expression returning a symbolic value, or simply a missing file.
 
-% DEAD FUNCTIONS and DEAD INCLUDES
-\begin{figure}
-	\input{diagrams/dead_features}
-	\caption{Explanation of never reached output candidates by dead code features.}
-	\label{dead_explanations}
-\end{figure}
+For a function to be never reached there are several scenarios: A function is undefined at runtime if the corresponding script file is never included.
+In turn, if the function is defined at runtime the function can either be dead code if there is no call site for this function, or the function call failed.
+In PHP there are several ways to call a function beside direct call sites.
+The language offers indirect call mechanisms like callback-commands or an evaluation function that parses and evaluates PHP source code as strings.
 
-\subsubsection{Why were code sections inaccessible?}  
-Eventually we started analyzing (1) why include expression may not be resolved/evaluated correctly, and (2) why a reasonable portion of defined functions is never called.
+Given a symbolic value, indirect call mechanisms like callback-commands are likely to fail since target function name and the symbolic value do not match.
+This also applies for the evaluation of include expressions as concrete values representing include targets can be included, but any include target containing symbolic values is ambiguous.
+
+\subsubsection{Why did includes fail?}
+\label{WhyDidIncludesFail}
+To better understand include expression evaluation, we use another metric to approach accuracy from a different angle: We measure (1) the ratio of reached include expressions and the total number of include expressions in a system respectively (coverage of include expressions), and (2) the ratio of successfully resolved include expressions and reached include expressions in a system (resolution success rate).
+For a most accurate result all include expressions are reached and resolved.
+
 For include expressions we sampled failed attempts and manually evaluated the cause of failure.
+We identified for (1) that all failed attempts were dynamic include expressions, i.e., those which were resolved at runtime and were assembled from fragments of environment-dependent information including database retrievals, user input, or files.
+Consequently, the resolved include expressions contained symbolic values which did not represent identifiable script files.
 
-We identified for (1) that all failed attempts were dynamic include expressions, i.e., those which are resolved at runtime and were assembled from fragments of environment-dependent information including database retrievals, user input, or files.
-Consequently, the resolved include expressions contained symbolic values which apparently did not represent identifiable script files.
-
-For (2) we again sampled dead function cases and identified the lack of direct calls to those functions.
-Interestingly, also functions despite having no direct call sites were called through indirect call features (see Section \ref{callback_candidate_section}).
-Again we identified dynamically assembled expressions (function names) to be responsible for failed attempts of calling a function indirectly as the resolved function name eventually contained symbolic values.
-These expressions, as well as include expressions, contained information dependent on the system environment.
-
-\subsubsection{Quantitative Analysis}
 Finally, we measured for include expressions the overall reach coverage and the overall resolution success rate per system.
-As illustrated in Figure \ref{include_coverage_results1}, except for the small-scale systems and two recent small-scale systems we did not even really reach a great portion of include expressions, although for most systems the resolution success rate was over 80~\% (see Figure \ref{include_coverage_results2}).
-As described in Section \ref{include_coverage_section} one failed attempt to include a script file possibly results in a cascade of missing more include expressions.
+As illustrated in Figure \ref{fig:include_coverage_results2}, except for the small-scale systems and two recent small-scale systems we did not even really reach a great portion of include expressions, although for most systems the resolution success rate was over 80 percent (see Figure \ref{fig:include_coverage_results1}).
+Nevertheless, one failed attempt to include a script file can result in a cascade of missing more include expressions as the include target script is not executed.
 
@@
         \caption{Proportion of reached include expressions per project.}
-        \label{include_coverage_results1}
+        \label{fig:include_coverage_results1}
     \end{subfigure}%
@@
         per project.}
-        \label{include_coverage_results2}
+        \label{fig:include_coverage_results2}
     \end{subfigure}
-    \label{include_coverage_results}
+    \label{fig:include_coverage_results}
     \caption{Include coverage and accuracy.}
@@
 
-For failed function calls to functions containing not-reached output candidates we measured how many of those functions were only, or partially accessibly through indirect calling mechanisms (see Section \ref{callback_candidate_section}).
-As Figure \ref{output_candidate_explanation} illustrates, around 80~\% (except for TimeClock) all missed output candidates of each system were only accessible through callback candidates, i.e., functions that are only accessible through indirect calling mechanisms.
+\subsubsection{Why did function calls fail?}
+\label{WhyDidFunctionCallsFail}
+Aside from failed or successful includes, this metric describes one scenario of why functions containing not-reached output candidates are never called: The trivial explanation would be that there are no direct calls of those functions, yet PHP offers a number of ways to call a function indirectly, for instance using built-in functions, or simply evaluating strings as PHP code using the \texttt{eval} function.
+
+Moreover we need to take into account that one missed function execution can result in missing even more function calls (direct or indirect) and executions.
+Thus, for functions containing not-reached output candidates we measure whether they can be (1) only, (2) partially or (3) never accessed transitively from functions that have no direct call sites.
+We refer to those functions (or access points) having no direct call site as callback candidates since the only way to reach them is through indirect mechanisms.
+Measuring how many not-reached output candidates can be explained by functions that are callback candidates helps to understand whether, and if so, indirect call mechanisms and their usage may impact our analysis’ code coverage.
+
+The callgraph snippets in Figure \ref{fig:callgraph} illustrate the idea of detecting the functions’ access points: The nodes at the top depict callback candidates, the nodes at the bottom with labels X, Y, and Z depict function definitions containing not-reached output candidates.
+In Figure \ref{fig:callgraph_a} function X can only be accessed through callback candidates, as for Figure \ref{fig:callgraph_b} function Y can be partially accessed by callback candidates, whereas function Z is never accessible through callback candidates.
+
+% Callback candidate description
+\begin{figure}
+	\centering
+	\begin{subfigure}[t]{0.18\textwidth}
+	\centering
+		\includegraphics[scale=0.4]{images-paper/callgraph_a}
+		\caption{Completely explained by callback candidates (no call from
+		\texttt{main})}
+		\label{fig:callgraph_a}
+    \end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.25\textwidth}
+	\centering
+		\includegraphics[scale=0.4]{images-paper/callgraph_b}
+		\caption{Partially (left; one call from \texttt{main}) or not at all  (right)
+		explained by callback candidates}
+		\label{fig:callgraph_b}
+    \end{subfigure}
+    \caption{Callgraph analysis: The bottom nodes represent functions containing missed output candidates, the top nodes depict entry points to them.}
+    \label{fig:callgraph}
+\end{figure}
+
+Like for include expressions, we again sampled dead function cases and identified the lack of direct call sites to those functions.
+As described in Section \ref{sec:inaccessible}, functions despite having no direct call sites were called through indirect call features.
+We identified dynamically assembled expressions (function names) to be responsible for failed attempts of calling a function indirectly as the resolved function name eventually contained symbolic values.
+These expressions, as well as include expressions, contained information dependent on the system environment.
+
+For failed function calls to functions containing not-reached output candidates we measured how many of those functions were only, or partially accessibly through indirect calling mechanisms (see Section \ref{sec:inaccessible}).
+As Figure \ref{fig:output_candidate_explanation} illustrates, around 80 percent (except for \textsc{TimeClock}) all missed output candidates of each system were only accessible through callback candidates, i.e., functions that are only accessible through indirect calling mechanisms.
 
 % CALLBACK CANDIDATE ACCESSIBILITY ANALYSIS
-\begin{figure}
+\begin{figure}[p]
 	\input{diagrams/output_candidates}
@@
 	completely explained due to callback candidates.}
-	\label{output_candidate_explanation}
+	\label{fig:output_candidate_explanation}
 \end{figure}
@@
 \section{Lessons Learned}
-\begin{itemize}
-  \item Certain language features are prone to be used in a way that symbolic
-  execution is impeded.
-  \item Tendency to use functions/includes with I/O, databases or configuration
-  files is problematic.
-\end{itemize}
 
-\begin{framed}
-\textcolor{RYB1}{The use of include expressions whose evaluation at runtime
-involves concatenation, especially of fragments from critical sources, impede
-symbolic execution due to inaccuracy.}
-\end{framed}
+		\begin{lstlisting}[caption=A misplaced and unused \texttt{listing} with no real value to this paper.]
+		function foo() {
+   		   echo "foo";
+		}
+		\end{lstlisting}
 
 \section{Related Work}
-\begin{itemize}
-  \item Static output approximation using context-free grammars
-  by Minamide \cite{minamide2005static} and detection of SQL
-  injection vulnerabilities \cite{wassermann2007sound}.
-  \item Static analysis of feature usage in PHP systems by Hills et al.
-  \cite{Hills:2013:ESP:2483760.2483786} and static resolution of include
-  expressions \cite{Hills:2013:ESP:2483760.2483786,hills2014static}
-  \item bener \cite{King1976} \cite{multiSE}
-\end{itemize}
 
-\newpage
+%References
 \bibliographystyle{abbrv}
-\bibliography{bibliography}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
+\bibliography{bibliography}
+
 \end{document}

### 2017-01-01 paper.tex
22.12.2016
01.01.2016

@@
 \usepackage{paralist}
-
+\usepackage{listings}
+\usepackage{lipsum}
+\usepackage{cite}
 \usepackage{enumerate}
@@
 language=PHP,
-basicstyle=\small\sffamily,
+basicstyle=\ttfamily\footnotesize,
+basewidth={0.5em,0.4em}, % for other columns settings
+%columns=fullflexible,
 numbers=left,
 numberstyle=\tiny,
-frame=tb,
-columns=fullflexible,
-showstringspaces=false
+frame=single,
+showstringspaces=false,
+xleftmargin=2em,
+framexleftmargin=1.5em
 }
@@
 \section{Introduction}
-With the emerge of the world wide web, dynamic web applications have become
+With the emerging world wide web, dynamic web applications have become
 widely popular. Various different implementation techniques have emerged,
@@
 programming community index \cite{tiobePHP}, and was the ranked as the 6th most
-popular language on GitHub \cite{githubPHP}.\\
+popular language on GitHub \cite{githubPHP}.
 
-One common property of all technologies for dynamic we b applications is the
-\emph{staged computation} of output, or simplified: dynamic code generation.
-Since based on both static values and runtime values, output code is generated step by step
-dynamically, output can be dependent on application inputs. Dynamic web
-applications incorporate various both server-side inputs (databases,
-configuration) and client-side inputs respectively. Static
-analysis of dynamic web applications, hence, has to consider different input
-sources.
+One common property of all technologies for dynamic web applications is \emph{staged computation} of output: A dynamic web application as a whole consists of both static code, such as scripts, and dynamically generated code, such as responses to HTTP requests.
+The latter code, although assembled at runtime, may contain client-side parts of the web application, such as JavaScript.
+So, to study dynamic web applications in its entirety, we need to consider both static as well as dynamic aspects of systems.
+
+%Since based on both static values and runtime values, output code is generated step by step
+%dynamically, output can be dependent on application inputs. Dynamic web
+%applications incorporate various both server-side inputs (databases,
+%configuration) and client-side inputs respectively. Static
+%analysis of dynamic web applications, hence, has to consider different input
+%sources.
 
@@
 result in variational output, representing dependencies of output and input.
-Symbolic execution extends concrete execution semantics with semantics to
-handle symbolic values. Hence, concrete execution semantics is contained in
-symbolic execution semantics.\\
+As symbolic execution of a program returns an input-output-mapping for symbolic inputs, this analysis technique helps unfold the \emph{staged} nature of web applications: Since every feasible path is executed all possible output is contained in the symbolic output, and all dynamically generated variational parts of the web application can be analyzed.
 
-Symbolic execution has been used and studied for generating test suites, but
-has more recently also been applied for approximating the output of web
-applications. Knowledge about different output variants enabled useful analyses
+%Symbolic execution extends concrete execution semantics with semantics to
+%handle symbolic values. Hence, concrete execution semantics is contained in
+%symbolic execution semantics. Symbolic execution has been used and studied for generating test suites, but
+%has more recently also been applied for approximating the output of web
+%applications.
+
+Knowledge about different output variants enabled useful analyses
 for the domain of web applications, such as detecting and locating HTML
@@
 execution has so far only been evaluated for smaller systems that are not
-developed anymore. The previously mentioned tools though are only practical, if
+maintained anymore. The previously mentioned tools though are only practical, if
 for a given system the symbolic execution engine is scalable, i.e., will also
-approximate reliable output for larger and more recent systems with good time
-and space consumption.\\
+approximate output accurately for larger and more recent systems with good time
+and space consumption.
 
-To investigate the question, whether we can have practical tools built from the
-symbolic execution engine, we re-implemented the engine with the specifications
+To investigate the question, whether we can build practical tools from the
+the symbolic execution engine, we re-implemented the engine with the specifications
 of the previous symbolic execution semantics
@@
 
-During introduction of new semantics for object-orientation and advanced array
-we encountered two main challenges. First, we walked a boundary between a
-purely symbolic or mixed concrete and symbolic (concolic) execution engine
+During introduction of new semantics for object-orientation we encountered two main challenges. 
+First, we walked a boundary between a purely symbolic or mixed concrete and symbolic execution engine
 since concrete execution, e.g. foreach loops for arrays, increased code
@@
 practical symbolic execution engine since for it to be accurate, a large number
-of functions from the standard library needs to be supported.\\
+of functions from the standard library needs to be supported.
 
-Our key contributions in this paper include: (1) The tool infrastructure to
-statically approximate the output for PHP web applications using symbolic
-execution and (2) observations as well as an experience evaluation to identify
-and measure PHP constructs impeding symbolic execution of PHP code.
+Our key contributions in this paper include (1) a new tool infrastructure to statically approximate the output for PHP web applications using symbolic execution, and (2) a report of our observations with state-of-the-art web applications as well as an explanation on conceptual limitations for symbolic execution of PHP applications.
 
 \section{State Of the Art}
+In this section we present the motivation for building tools that address the \textcolor{blue}{development, quality and xxx} of dynamic web applications, but require an accurate static output approximation.
+All tools proposed so far \cite{Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140,Nguyen:2011:AFH:2190078.2190142} leverage a representation of all possible HTML client page output.
+Since any variant can be analyzed in particular, this enables tool support addressing the web application as a whole despite its \emph{staged} nature, where parts of the web application (client-side output) is just generated dynamically.
 
-\section{Towards scalable symbolic execution}%
+The work-flow for all these tools is illustrated in Figure \ref{fig:approach}, where prior to any analysis, the web application is symbolically executed and all client page output variants are parsed.
+To illustrate the concept of symbolic execution for this purpose consider the following example:
+
+\begin{figure}[t!]
+	\centering
+	\includegraphics[width=0.45\textwidth]{images-paper/approach}
+
+	\caption{Bener}
+	\label{fig:approach}
+\end{figure}
+
+\begin{lstlisting}
+<?php
+
+echo 'Hello World!';
+\end{lstlisting}
+
+Based on this output representation, the so-called D-Model, subsequent analyses can be conducted:
+
+\begin{compactitem}
+\item \emph{PhpSync}: Using all variants of  HTML client page output, every single one can be statically checked for Markup Validity, i.e., if the web page conforms to syntactical specifications for HTML and other client-side languages.
+After validation errors are detected, auto-fixes are provided.
+Otherwise, the validation error can be traced back to source code responsible for the defect, which then can be refactored manually \cite{Nguyen:2011:AFH:2190078.2190142}.
+
+\item \emph{WebSlice}: Bener bener
+\item \emph{Varis}: Bener IDE bener
+\end{compactitem}
+
+\textcolor{blue}{\emph{D. Tool evaluation critique}
+\begin{compactitem}
+	\item Limitations/Assumptions
+	\item Too small sample
+\end{compactitem}
+}
+
+\section{Scalable Output Approximation}%
 
@@
 \centering 
-	\begin{tabular}{p{4cm}rrrl}
+	\begin{tabular}{lrp{4cm}rlrr}
 	\toprule
-	\textbf{System} & \textbf{Version} & \textbf{File Count} & \textbf{SLOC} &
-	\textbf{Description}
+	\textbf{System} & \textbf{Version} & \textbf{Classification} & \textbf{SLOC} & \textbf{\#files} & \textbf{\#OCs} & \textbf{\#includes}
 	\\
 	\midrule
-	AddressBook & 8.2.5.2 & 239 & 51,907 & Adress Administration \\
-	SchoolMate & 1.5.4 & 65 & 8,118 & School Administration \\
-	TimeClock & 1.04 & 63 & 20,800 & XXX \\
-	WebChess & 1.0.0 & 28 & 5,219 & XXX \\
+	\textsc{AddressBook} & 8.2.5.2 & 239 & 51,907 & Adress Administration  & XXX & XXX\\
+	\textsc{SchoolMate} & 1.5.4 & 65 & 8,118 & School Administration  & XXX & XXX\\
+	\textsc{TimeClock} & 1.04 & 63 & 20,800 & XXX  & XXX & XXX\\
+	\textsc{WebChess} & 1.0.0 & 28 & 5,219 & XXX  & XXX & XXX\\
 	\midrule
-	Drupal & 7.5.0 & 125 & 52,464 & Content Management System \\
-	phpBB & 3.1.9 & 1,398 & 327,371 & Bulletin Board \\
-	phpMyAdmin & 4.6.3 & 871 & 303,582 & Database Administration \\
+	\textsc{Drupal} & 7.5.0 & 125 & 52,464 & Content Management System  & XXX & XXX\\
+	\textsc{phpBB} & 3.1.9 & 1,398 & 327,371 & Bulletin Board  & XXX & XXX\\
+	\textsc{phpMyAdmin} & 4.6.3 & 871 & 303,582 & Database Administration  & XXX & XXX\\
 	\midrule
-	Anchor & 0.12.1 & 201 & 15,054 & Content Management System \\
-	Kirby & XXX & XXX & XXX & Content Management System \\
-	Automad & XXX & XXX & XXX & Content Management System \\
-	Monstra & XXX & XXX & XXX & Content Management System \\
-	Nibbleblog & XXX & XXX & XXX & Content Management System \\
+	\textsc{Anchor} & 0.12.1 & 201 & 15,054 & Content Management System  & XXX & XXX\\
+	\textsc{Kirby} & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
+	\textsc{Automad} & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
+	\textsc{Monstra} & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
+	\textsc{Nibbleblog} & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
 	\bottomrule
@@
 	\centering	
-	\resizebox{0.4\textwidth}{!}{%
-		\input{diagrams/coverage}	
+%	\resizebox{0.4\textwidth}{!}{%
+%		\input{diagrams/coverage}	
+%	}
+	\includegraphics[width=0.5\textwidth]{images-paper/coverage.eps}
+	\caption{
+		Results for output candidates: \emph{Coverage loss} repesents reached output candidates that were not part of the output.
+\emph{Simple file} and \emph{function definition} represent context classifications of missed output candidates.
 	}
-	\caption{Coverage results for the PHP corpus}
 	\label{coverage}
@@
 % - TEMPLATE CONTEXT
-\begin{figure}
-	\resizebox{0.4\textwidth}{!}{%
-		\input{diagrams/context_distribution}
-	}	
-	\caption{Context distribution of never reached output candidates for the PHP
-	corpus.}
-	\label{fig:contexts}
-\end{figure}
+%\begin{figure}
+%	\resizebox{0.4\textwidth}{!}{%
+%		\input{diagrams/context_distribution}
+%	}	
+%	\caption{Context distribution of never reached output candidates for the PHP
+%	corpus.}
+%	\label{fig:contexts}
+%\end{figure}
 
@@
 % INCLUDE COVERAGE / SUCCESS RATE
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.5 \textwidth}    	
-        \centering        
-        % Include Coverage Plot        
-		\input{diagrams/include_coverage}		
-        \caption{Proportion of reached include expressions per project.}
-        \label{fig:include_coverage_results1}
-    \end{subfigure}%
-	\hfill
-    \begin{subfigure}[t]{0.5\textwidth}    
-        \centering
-        % Success Rate Plot        
-        \input{diagrams/success_rate}        
-        \caption{Rate of successful resolutions of reached include expressions
-        per project.}
-        \label{fig:include_coverage_results2}
-    \end{subfigure}
+\begin{figure}
+%    \centering
+%    \begin{subfigure}[t]{0.5 \textwidth}    	
+%        \centering        
+%        % Include Coverage Plot        %
+%		\input{diagrams/include_coverage}		
+%        \caption{Proportion of reached include expressions per project.}
+%        \label{fig:include_coverage_results1}
+%    \end{subfigure}
+%	\hfill
+%    \begin{subfigure}[t]{0.5\textwidth}    
+%        \centering
+%        % Success Rate Plot        
+%        \input{diagrams/success_rate}        
+%        \caption{Rate of successful resolutions of reached include expressions
+%        per project.}
+%        \label{fig:include_coverage_results2}
+%    \end{subfigure}
+
+\includegraphics[width=0.5\textwidth]{images-paper/includes.eps}
+
     \label{fig:include_coverage_results}
     \caption{Include coverage and accuracy.}
-\end{figure*}
+\end{figure}
 
@@
 % Callback candidate description
-\begin{figure}
+\begin{figure}[t!]
 	\centering
-	\begin{subfigure}[t]{0.18\textwidth}
-	\centering
-		\includegraphics[scale=0.4]{images-paper/callgraph_a}
-		\caption{Completely explained by callback candidates (no call from
-		\texttt{main})}
-		\label{fig:callgraph_a}
-    \end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.25\textwidth}
-	\centering
-		\includegraphics[scale=0.4]{images-paper/callgraph_b}
-		\caption{Partially (left; one call from \texttt{main}) or not at all  (right)
-		explained by callback candidates}
-		\label{fig:callgraph_b}
-    \end{subfigure}
+	\includegraphics[scale=0.55]{images-paper/callgraph_cases}
     \caption{Callgraph analysis: The bottom nodes represent functions containing missed output candidates, the top nodes depict entry points to them.}
@@
 
-		\begin{lstlisting}[caption=A misplaced and unused \texttt{listing} with no real value to this paper.]
-		function foo() {
-   		   echo "foo";
-		}
-		\end{lstlisting}
-
 \section{Related Work}
+\begin{itemize}
+	\item Hills et al. \cite{Hills:2013:ESP:2483760.2483786}
+	\item Output approximation using context-free grammar construction; applications \cite{minamide2005static,wassermann2007sound}
+\end{itemize}
 

### 2017-01-01 paper.tex
01.01.2016 (2)

@@
 \cite{Nguyen:2014:BCG:2635868.2635928} for PHP and additional features, such as
-object-oriented programming, and extensive array support. We evaluated our
+object-oriented programming. We evaluated our
 symbolic execution engine for large-scale and modern PHP systems.
@@
 During introduction of new semantics for object-orientation we encountered two main challenges. 
-First, we walked a boundary between a purely symbolic or mixed concrete and symbolic execution engine
-since concrete execution, e.g. foreach loops for arrays, increased code
-coverage significantly. Second, we had to make a trade-off between the accuracy
+First, we were bound to introduce more concrete execution features to the symbolic execution engine, e.g.
+foreach loops for arrays, as this increased code coverage significantly.
+Second, we had to make a trade-off between the accuracy
 of method calls and the scalability of this language feature as a method call

### 2017-01-03 paper.tex
3.1.2016 es schneit

@@
 \documentclass[preprint]{sig-alternate-05-2015}
+
+\setlength\parindent{0pt}
 
@@
 
+\usepackage[zerostyle=d]{newtxtt}
+%\renewcommand*\familydefault{\ttfamily} %% Only if the base font of the document is to be typewriter style
+\usepackage[T1]{fontenc}
+
 \usepackage[labelfont=bf]{caption}
@@
 language=PHP,
-basicstyle=\ttfamily\footnotesize,
+basicstyle=\ttfamily\tiny,
 basewidth={0.5em,0.4em}, % for other columns settings
@@
     keepspaces = true,
-    keywordstyle = \bfseries
+    keywordstyle = \bfseries,
+    frame = tb
 }
@@
 \emph{Symbolic execution} is a static analysis technique between program testing
-and program proving. For a program, symbolic, i.e., abstract, yet fix inputs are
+and program proving. For a program, symbolic, i.e., abstract, yet fixed inputs are
 applied instead of concrete inputs \cite{Darringer1978,King1976}. The underlying
@@
 \section{State Of the Art}
-In this section we present the motivation for building tools that address the \textcolor{blue}{development, quality and xxx} of dynamic web applications, but require an accurate static output approximation.
-All tools proposed so far \cite{Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140,Nguyen:2011:AFH:2190078.2190142} leverage a representation of all possible HTML client page output.
-Since any variant can be analyzed in particular, this enables tool support addressing the web application as a whole despite its \emph{staged} nature, where parts of the web application (client-side output) is just generated dynamically.
+Symbolic execution is one of integral analysis techniques used by the tools whose practicality we aim to evaluate.
+This section recaps the idea of symbolic execution, the tools which make use of it, and both conceptual and language-related limitations  of symbolic execution applied to PHP so far.
 
-The work-flow for all these tools is illustrated in Figure \ref{fig:approach}, where prior to any analysis, the web application is symbolically executed and all client page output variants are parsed.
-To illustrate the concept of symbolic execution for this purpose consider the following example:
+\subsection{Symbolic Execution}
+Symbolic execution is a static analysis technique that was proposed by James C.
+King in 1976 \cite{King1976}.
+Symbolic execution allows to explore all feasible paths in a program.
+Based on normal program execution, the execution semantics is extended to handle symbolic values, which are abstract, but fixed; an execution becomes symbolic by introducing symbolic values as program inputs instead of concrete values \cite{King1976,Darringer1978}.
+These symbolic values are then propagated and used throughout the execution.
 
+Execution starts with a plain path condition, usually a tautology.
+Whenever the control-flow can be split into different branches, the corresponding path conditions (guards for instance) are conjoined with the previous path condition \cite{King1976}.
+Additionally, the conjoined path condition can be checked for satisfiability to exclude infeasible paths.
+Consequently, for a given program symbolic execution computes a mapping from (symbolic) input values to output values.
+
+% Introduction example for symbolic execution:
+\begin{figure}
+\begin{lstlisting}
+<?php
+$x = $_POST['name'];
+$y = $x + 1;
+\end{lstlisting}
+
+\begin{lstlisting}
+<?php
+function foo($arg) {
+   echo $arg;	
+   foo($arg);
+}
+$z = foo(3);
+\end{lstlisting}
+\begin{lstlisting}
+<?php
+if ($z < 0) {
+   $r = 'A';
+} else {
+   $r = 'B';
+}
+echo $r;
+\end{lstlisting}
+\begin{lstlisting}
+<?php
+while ($z > 0) {
+   echo 'B';
+   $z = $z - 1;
+}
+\end{lstlisting}
+\caption{Example for symbolic execution of PHP code\label{fig:symbolic}}
+\end{figure}
+
+\subsection{Workflow And Existing Tools} 
+The work-flow for all existing tools is illustrated in Figure \ref{fig:approach}.
+Prior to any analysis, the web application is symbolically executed which for every execution path yields its corresponding client page output variant.
+Next, all client page output variants are parsed and merged into a compact representation of all client page output variants.
+This output model, also referred to as \emph{VarDOM} (Variational Document Object Model), is similar to a DOM, yet it can contain symbolic value nodes and conditional node types to represent variants as alternatives.
+
+% Workflow Graphic
 \begin{figure}[t!]
@@
 
-\begin{lstlisting}
-<?php
+All tools proposed so far leverage a representation of all possible HTML client page output.
+Since any variant can be analyzed in particular, this enables tool support addressing the web application as a whole in spite of its staged nature, where parts of the web application (client-side output) are generated dynamically.
 
-echo 'Hello World!';
-\end{lstlisting}
-
-Based on this output representation, the so-called D-Model, subsequent analyses can be conducted:
+Based on the approximated output representation subsequent analyses can be conducted, which have been implemented in the following tools:
 
@@
 
-\item \emph{WebSlice}: Bener bener
-\item \emph{Varis}: Bener IDE bener
+\item \emph{WebSlice}: Program slices enable to extract and understand the impact of changes in an application. 
+To consider all of a dynamic web application for program slices, client page output needs to be taken into account.
+WebSlice combines PHP data-flow information with an output approximation to enable program slices across different languages \cite{Nguyen:2015:CPS:2786805.2786872}.
+
+\item \emph{Varis}: Editor services such as “jump to declaration” are nontrivial for dynamic web applications due to their staged nature.
+Varis provides editor services across stages for client-side code: Starting with an ouput approximation a callgraph with conditional edges is constructed, which allows navigation in client-site code (HTML, JS and CSS) although it is embedded in server-side code \cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
 \end{compactitem}
 
-\textcolor{blue}{\emph{D. Tool evaluation critique}
+\subsection{Assumptions and Limitations}
+\textcolor{blue}{
 \begin{compactitem}
-	\item Limitations/Assumptions
-	\item Too small sample
+	\item Design Decisions: Single loop and recursion execution due to conceptual limitations
+	\item Conceptual Limitations: Symbolic input as form input, interaction
+	\item Language-related or implementation-specific: No OOP, array support, standard lib
 \end{compactitem}
@@
 \subsection{Measuring Approximation Success} \label{heuristic}
-For the evaluation of our output approximation we require a ground truth output model to compare against.
-In the absence of ground truth for the approximated output we are bound to choose a \emph{heuristic approach} in order to measure the accuracy of our aproximation.
+For the evaluation of our output approximation we require a ground truth output model to compare our approximation against.
+Since output literals are embedded in server-side code and scattered, it is nontrivial to determine whether a string literal is part of the ouput or not.
+So, in the absence of ground truth for the approximated output we are bound to choose a \emph{heuristic approach} in order to measure the accuracy of our aproximation.
 We define a string literal to be an \emph{output candidate} if there exists an execution path reaching the string literal and evantually passing it to an
@@
 used did not support all language features used.
+For the experimental evaluation we symbolically executed per system each script file or entry point (files with a \texttt{.php}, \texttt{.inc}, \texttt{.bit} or \texttt{.module} extension).
+All following measurements are cumulated per system for each entry point.
 
@@
 	\midrule
-	\textsc{AddressBook} & 8.2.5.2 & 239 & 51,907 & Adress Administration  & XXX & XXX\\
-	\textsc{SchoolMate} & 1.5.4 & 65 & 8,118 & School Administration  & XXX & XXX\\
-	\textsc{TimeClock} & 1.04 & 63 & 20,800 & XXX  & XXX & XXX\\
-	\textsc{WebChess} & 1.0.0 & 28 & 5,219 & XXX  & XXX & XXX\\
+	AddressBook & 8.2.5.2 & 239 & 51,907 & Adress Administration  & XXX & XXX\\
+	SchoolMate & 1.5.4 & 65 & 8,118 & School Administration  & XXX & XXX\\
+	TimeClock & 1.04 & 63 & 20,800 & XXX  & XXX & XXX\\
+	WebChess & 1.0.0 & 28 & 5,219 & XXX  & XXX & XXX\\
 	\midrule
-	\textsc{Drupal} & 7.5.0 & 125 & 52,464 & Content Management System  & XXX & XXX\\
-	\textsc{phpBB} & 3.1.9 & 1,398 & 327,371 & Bulletin Board  & XXX & XXX\\
-	\textsc{phpMyAdmin} & 4.6.3 & 871 & 303,582 & Database Administration  & XXX & XXX\\
+	Drupal & 7.5.0 & 125 & 52,464 & Content Management System  & XXX & XXX\\
+	phpBB & 3.1.9 & 1,398 & 327,371 & Bulletin Board  & XXX & XXX\\
+	phpMyAdmin & 4.6.3 & 871 & 303,582 & Database Administration  & XXX & XXX\\
 	\midrule
-	\textsc{Anchor} & 0.12.1 & 201 & 15,054 & Content Management System  & XXX & XXX\\
-	\textsc{Kirby} & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
-	\textsc{Automad} & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
-	\textsc{Monstra} & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
-	\textsc{Nibbleblog} & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
+	Anchor & 0.12.1 & 201 & 15,054 & Content Management System  & XXX & XXX\\
+	Kirby & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
+	Automad & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
+	Monstra & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
+	Nibbleblog & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
 	\bottomrule
@@
 \end{table*}
-
-For the experimental analysis we symbolically executed per system each script file or entry point (files with a \texttt{.php}, \texttt{.inc}, \texttt{.bit} or \texttt{.module} extension).
-All following measurements are cumulated per system for each entry point.
 
@@
 
-\subsubsection{What literals did we miss?}
-\label{WhatLiteralsDidWeMiss?}
+\subsubsection{What literals did we miss?}\label{WhatLiteralsDidWeMiss?}
 Our initial approach to understand missed output candidates is to find out whether their surrounding program code was not accessed/accessible, and if so, why.
@@
 We started classifying missed output candidates by their string literal context.
-As the classification statistics in Figure \ref{fig:contexts} illustrate, for almost all systems (except for \textsc{AddressBook} and \textsc{TimeClock}) not-reached output candidates had a function context.
-Note that \textsc{SchoolMate} was excluded from the diagram since our analysis reached all output candidates for that system.
+As the classification statistics in Figure \ref{coverage} illustrate, for almost all systems (except for \textsf{AddressBook} and \textsf{TimeClock}) not-reached output candidates had a function context.
+Note that \textsf{SchoolMate} was excluded from the diagram since our analysis reached all output candidates for that system.
 
@@
 Finally, we measured for include expressions the overall reach coverage and the overall resolution success rate per system.
-As illustrated in Figure \ref{fig:include_coverage_results2}, except for the small-scale systems and two recent small-scale systems we did not even really reach a great portion of include expressions, although for most systems the resolution success rate was over 80 percent (see Figure \ref{fig:include_coverage_results1}).
+As illustrated in Figure \ref{fig:include_coverage_results}, except for the small-scale systems and two recent small-scale systems we did not even really reach a great portion of include expressions, although for most systems the resolution success rate was over 80 percent.
 Nevertheless, one failed attempt to include a script file can result in a cascade of missing more include expressions as the include target script is not executed.
@@
 
-    \label{fig:include_coverage_results}
-    \caption{Include coverage and accuracy.}
+    \caption{Include coverage and accuracy.\label{fig:include_coverage_results}}
 \end{figure}
@@
 The callgraph snippets in Figure \ref{fig:callgraph} illustrate the idea of detecting the functions’ access points: The nodes at the top depict callback candidates, the nodes at the bottom with labels X, Y, and Z depict function definitions containing not-reached output candidates.
-In Figure \ref{fig:callgraph_a} function X can only be accessed through callback candidates, as for Figure \ref{fig:callgraph_b} function Y can be partially accessed by callback candidates, whereas function Z is never accessible through callback candidates.
+In Figure \ref{fig:callgraph} function X can only be accessed through callback candidates, as for Figure \ref{fig:callgraph} function Y can be partially accessed by callback candidates, whereas function Z is never accessible through callback candidates.
 
@@
 For failed function calls to functions containing not-reached output candidates we measured how many of those functions were only, or partially accessibly through indirect calling mechanisms (see Section \ref{sec:inaccessible}).
-As Figure \ref{fig:output_candidate_explanation} illustrates, around 80 percent (except for \textsc{TimeClock}) all missed output candidates of each system were only accessible through callback candidates, i.e., functions that are only accessible through indirect calling mechanisms.
+As Figure \ref{fig:output_candidate_explanation} illustrates, around 80 percent (except for \textsf{TimeClock}) all missed output candidates of each system were only accessible through callback candidates, i.e., functions that are only accessible through indirect calling mechanisms.
 

### 2017-01-05 paper.tex
workflow refine
5.1

@@
 
-\section{State Of the Art}
+\section{State Of the Art}\label{sec:stateoftheart}
 Symbolic execution is one of integral analysis techniques used by the tools whose practicality we aim to evaluate.
-This section recaps the idea of symbolic execution, the tools which make use of it, and both conceptual and language-related limitations  of symbolic execution applied to PHP so far.
+This section recaps the idea of symbolic execution, the tools which make use of it, and both conceptual and language-related limitations.
 
@@
 
-% Introduction example for symbolic execution:
-\begin{figure}
-\begin{lstlisting}
-<?php
-$x = $_POST['name'];
-$y = $x + 1;
-\end{lstlisting}
-
-\begin{lstlisting}
-<?php
-function foo($arg) {
-   echo $arg;	
-   foo($arg);
-}
-$z = foo(3);
-\end{lstlisting}
-\begin{lstlisting}
-<?php
-if ($z < 0) {
-   $r = 'A';
-} else {
-   $r = 'B';
-}
-echo $r;
-\end{lstlisting}
-\begin{lstlisting}
-<?php
-while ($z > 0) {
-   echo 'B';
-   $z = $z - 1;
-}
-\end{lstlisting}
-\caption{Example for symbolic execution of PHP code\label{fig:symbolic}}
-\end{figure}
-
-\subsection{Workflow And Existing Tools} 
+\subsection{Workflow And Existing Tools} \label{sec:workflow}
 The work-flow for all existing tools is illustrated in Figure \ref{fig:approach}.
@@
 % Workflow Graphic
-\begin{figure}[t!]
-	\centering
-	\includegraphics[width=0.45\textwidth]{images-paper/approach}
-
-	\caption{Bener}
+\begin{figure*}[t!]
+	\centerline{
+	\includegraphics[width=0.95\textwidth]{images-paper/workflow}
+	}
+	\vspace{2mm}
+	\caption{
+		Overview of the workflow for tools based on static output approximation for dynamic web applications.	
+	}
 	\label{fig:approach}
-\end{figure}
+\end{figure*}
 
@@
 \subsection{Assumptions and Limitations}
-\textcolor{blue}{
-\begin{compactitem}
-	\item Design Decisions: Single loop and recursion execution due to conceptual limitations
-	\item Conceptual Limitations: Symbolic input as form input, interaction
-	\item Language-related or implementation-specific: No OOP, array support, standard lib
-\end{compactitem}
-}
+Symbolic execution remains an analysis technique not dedicated to a particular programming language.
+For the implementation of the symbolic execution engine in \cite{Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872}, several assumptions have been made to adapt to the PHP language.
+The following symbolic execution semantics for PHP features differ from concrete execution since symbolic values need to be handled without ambiguity.
+
+First, loops in the previous execution semantics are only executed once instead of an arbitrary number of times.
+Since a loop condition can contain symbolic values, it might not be feasible to determine when a loop terminates.
+Hence, any output constructed during the single loop iteration is highlighted to possibly be repeated  an arbitrary number of times.
+Second, recursive functions are also only executed once per call, and any subsequent call ot that function returns a symbolic value.
+This, again, is due to the difficulty to determine the exact recursion depth similar to the number of loop iterations.
+Finally, since the symbolic execution engine applies a static analysis, any form of user interaction providing input data as well as any form of input referred to from the deployment context of the web application, i.e.
+configuration files or databases, is substituted by a symbolic value.
+
+Language-wise the symbolic execution semantics for the tools presented in Section \ref{sec:workflow} currently does not support object-oriented programming.
+PHP provides a large number of built-in functions to manipulate strings as well as arrays.
+The current symbolic execution engine only provides implementations for a small range of PHP standard library functions, mainly string operations.
 
 \section{Scalable Output Approximation}%
+The main objective of this experience report is to investigate whether the tools described in Section \ref{sec:stateoftheart} are practical.
+In particular we want to find out whether symbolic execution as the underlying technique for static output approximation scales for PHP, and what may limitations restrict scalability.
+We may divide possible limitations into either matters of engineering or conceptual limitations where no precise execution is possible.
+
+In the following we describe in detail our implementation procedure, practical problems we encountered and whether these are limiting scalability conceptually or just require engineering effort to be tamed.
+
+\subsection{Towards Scalable Output Approximation?}
+We re-implement\footnote{\url{https://github.com/smba/oak}} the existing symbolic execution semantics of \cite{Nguyen:2014:BCG:2635868.2635928} with the goal to support modern large-scale dynamic web applications, or at least to better understand possible limitations.
+For reference, we choose to aim to support  \textsf{WordPress}\footnote{\url{https://github.com/WordPress/WordPress}}, a popular Content Management System of reasonable size that is still maintained.
+Moreover, our implementation uses parts of Quercus\footnote{\url{http://quercus.caucho.com/}}, a pure Java implementation of the PHP interpreter, which is tested against the \textsf{WordPress} test suite.
+We choose a test-driven approach for our implementation with regression tests and continuous integration.
+Our test cases are based on both custom examples for different language features as well as \textsf{SchoolMate}, a small-scale system, and, at least in parts, \textsf{WordPress}, whose output we manually validate and then automatically re-check for each test run.
+
+Our implementation basically follows the symbolic execution semantics of \cite{Nguyen:2014:BCG:2635868.2635928}, yet we extend it with additional rules for class definitions, class instantiation, and method invocation to support object-oriented programming.
+Furthermore, we add rules for do-while and  foreach loops since these are not supported as yet.
+Note that we execute all iterations of a foreach loop since the number of iterations is known since we are traversing all items of a given array.
 

### 2017-01-05 paper.tex
added code example for sec 2, maybe simplify

@@
 \usepackage{paralist}
-\usepackage{listings}
+\usepackage{listings,multicol}
 \usepackage{lipsum}
@@
 	basewidth={0.5em,0.4em}, % for other columns settings
-%columns=fullflexible,
+	columns=fullflexible,
 	numbers=left,
@@
 	xleftmargin=2em,
-framexleftmargin=1.5em
+	framexleftmargin=1.5em,
 }
+
+\newcommand{\dollar}{\mbox{\textdollar}}
 
@@
 
+\begin{figure*}
+	\centering
+	\begin{subfigure}[center]{0.49\textwidth}%%[multicols=1]
+\begin{lstlisting}
+<form>
+  <input type="text" name="name">
+  <input type="submit" value="Submit">
+</form>
+
+<?php
+  function titles($arg, $n) {
+    $is = array(1,2,3);
+    while ($n < shuffle($is)) {
+      echo "<h1>" . $arg . "</h1><br>";
+      $n += 1;
+    }
+  }
+
+  call_user_func("title", "Headline", 4);
+
+  $name = $_POST['name'];
+  if (!isset($name)) {
+    echo "No name entered!";
+  } else {
+    echo "Entered ". $name ."!";
+  }
+?>
+\end{lstlisting}
+		\caption{Snippet of server-side web application code}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[center]{0.49\textwidth}
+\begin{lstlisting}[mathescape]
+<form>
+  <input type="text" name="name">
+  <input type="submit" value="Submit">
+</form>
+
+// #repeat n < shuffle(is)
+  <h1>Headline</h1>
+// #endrepeat
+
+// #if !isset(name)
+  No name entered!
+// #else
+  Entered $\alpha$!
+// #endif
+
+\end{lstlisting}
+		\caption{Symbolic output approximation}
+    \end{subfigure}
+    \vspace{2mm}
+    \caption{Illustration of static ouptut approximation using symbolic
+    execution}
+\end{figure*}
+
 % Workflow Graphic
-\begin{figure*}[t!]
-	\centerline{
-	\includegraphics[width=0.95\textwidth]{images-paper/workflow}
-	}
-	\vspace{2mm}
-	\caption{
-		Overview of the workflow for tools based on static output approximation for dynamic web applications.	
-	}
-	\label{fig:approach}
-\end{figure*}
+%\begin{figure*}[t!]
+%	\centerline{
+%	\includegraphics[width=0.95\textwidth]{images-paper/workflow}
+%	}
+%	\vspace{2mm}
+%	\caption{
+%		Overview of the workflow for tools based on static output approximation for
+%		% dynamic web applications.
+%	}
+%	\label{fig:approach}
+%\end{figure*}
 
@@
 \begin{compactitem}
-\item \emph{PhpSync}: Using all variants of  HTML client page output, every single one can be statically checked for Markup Validity, i.e., if the web page conforms to syntactical specifications for HTML and other client-side languages.
-After validation errors are detected, auto-fixes are provided.
-Otherwise, the validation error can be traced back to source code responsible for the defect, which then can be refactored manually \cite{Nguyen:2011:AFH:2190078.2190142}.
+\item \emph{PhpSync}: Using all variants of  HTML client page output, every
+single one can be statically checked for Markup Validity, i.e., if the web page
+conforms to syntactical specifications for HTML and other client-side
+languages. After validation errors are detected, auto-fixes are provided.
+Otherwise, the validation error can be traced back to source code responsible
+for the defect, which then can be refactored manually
+\cite{Nguyen:2011:AFH:2190078.2190142}.
 
-\item \emph{WebSlice}: Program slices enable to extract and understand the impact of changes in an application. 
-To consider all of a dynamic web application for program slices, client page output needs to be taken into account.
-WebSlice combines PHP data-flow information with an output approximation to enable program slices across different languages \cite{Nguyen:2015:CPS:2786805.2786872}.
+\item \emph{WebSlice}: Program slices enable to extract and understand the
+impact of changes in an application. To consider all of a dynamic web
+application for program slices, client page output needs to be taken into
+account. WebSlice combines PHP data-flow information with an output
+approximation to enable program slices across different languages
+\cite{Nguyen:2015:CPS:2786805.2786872}.
 
-\item \emph{Varis}: Editor services such as “jump to declaration” are nontrivial for dynamic web applications due to their staged nature.
-Varis provides editor services across stages for client-side code: Starting with an ouput approximation a callgraph with conditional edges is constructed, which allows navigation in client-site code (HTML, JS and CSS) although it is embedded in server-side code \cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
+\item \emph{Varis}: Editor services such as “jump to declaration” are
+nontrivial for dynamic web applications due to their staged nature. Varis
+provides editor services across stages for client-side code: Starting with an
+ouput approximation a callgraph with conditional edges is constructed, which
+allows navigation in client-site code (HTML, JS and CSS) although it is
+embedded in server-side code
+\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
 \end{compactitem}

### 2017-01-07 diagrams/context_distribution.tex
Section 2 2nd iteration + back to tikz

@@
+    \pgfplotstableread{ % data 
+Label     Reached/Output OnlyReached FunctionContext Bener
+Adressbook 3 2 1.4 3
+Schoolmate 3 2 1 2
+Timeclock 3 2 1 2
+Webchess 3 2 1 2
+PhpBB 3 2 1 2
+Drupal 3 2 1 2
+phpMyAdmin 3 2 1 2
+Anchor 3 2 1 2
+Kirby 3 2 1 2
+Automad 3 2 1 2
+Monstra 3 2 1 2
+Nibbleblog 3 2 1 2
+    }\testdata
+
     \begin{tikzpicture}[thick,scale=0.8, every node/.style={transform shape}]
+
     \begin{axis}[
-    ybar stacked,
-	bar width=12pt,
+    enlargelimits=0.05,
+    xbar stacked,   % Stacked horizontal bars
+    xlabel={Proportion [\%]},
+    xmin=0,         % Start x axis at 0
+    ytick=data,     % Use as many tick labels as y coordinates
+    legend style={
+       at={(0.5,-0.28)}, 
+       anchor=north,
+       legend columns = -1,
+       cells={anchor=center, fill},
+       nodes={inner sep=1,below=-1.1ex}
+    },
+    yticklabels from table={\testdata}{Label}  % Get the labels from the Label column of the \datatable
+    ,bar width=2ex, y=3ex,
     nodes near coords = {%
        \pgfmathprintnumberto[fixed,assume math mode=true]{\pgfplotspointmeta}{\myval}%
-    \pgfmathparse{\myval<101?:\myval}\pgfmathresult%
+       \pgfmathparse{\myval<101?:}\pgfmathresult%
     },
-    enlargelimits=0.15,
-    legend style={at={(0.5,-0.28)},
-      anchor=north,legend columns=-1},
-    ylabel={Proportion [\%]},
-    symbolic x coords={AddressBook, TimelClock, WebChess, Drupal,
-    phpBB, phpMyAdmin, Anchor, Kirby, Automad, Monstra, Nibbleblog}, xtick=data,
-    x tick label style={rotate=45,anchor=east},
-    cycle list name=ColorListBar,
+    area legend
     ]
+    \addplot [fill=lightgrey!80] table [x=Reached/Output, meta=Label,y expr=\coordindex] {\testdata};
+    \addplot [fill=blue!60] table [x=OnlyReached, meta=Label,y expr=\coordindex] {\testdata};
+    \addplot [fill=red!60] table [x=FunctionContext, meta=Label,y expr=\coordindex] {\testdata};
+    \addplot [fill=red!99] table [x=Bener, meta=Label,y expr=\coordindex] {\testdata};
 
-\addplot+[ybar] plot coordinates {(AddressBook, 28) 
-  (TimelClock, 0) (WebChess, 86) (Drupal, 100) (phpBB, 96) (phpMyAdmin, 92)
-  (Anchor, 100) (Kirby, 96) (Automad, 93) (Monstra, 83) (Nibbleblog, 91)};
-\addplot+[ybar] plot coordinates {(AddressBook, 5) 
-  (TimelClock, 0) (WebChess, 0) (Drupal, 0) (phpBB, 0) (phpMyAdmin, 1)
-  (Anchor, 0) (Kirby, 4) (Automad, 0) (Monstra, 2) (Nibbleblog, 0)};
-\addplot+[ybar] plot coordinates {(AddressBook, 67) 
-  (TimelClock, 100) (WebChess, 14) (Drupal, 0) (phpBB, 4) (phpMyAdmin, 7)
-  (Anchor, 0) (Kirby, 0) (Automad, 7) (Monstra, 15) (Nibbleblog, 9)};
-
-\legend{Function Definition, HTML Template, Miscellaneous}
+    \legend{Reached/Output, OnlyReached, FunctionContext, Bener}
     \end{axis}
-%\caption{Relative distribution of contexts of not reached output candidates}
     \end{tikzpicture}

### 2017-01-07 diagrams/coveragencontext.tex
Section 2 2nd iteration + back to tikz

\documentclass{article}
\usepackage{tikz,pgfplots,pgfplotstable}

  \begin{document}

    \pgfplotstableread{ % data 
Label     Adequacy Sustainability Integrity
Austria   21.87797  5.566584 24.85545
Brazil    20.93009  7.937881 24.33203
Chile     20.41148 21.126529 27.56199
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
    }\testdata

    \begin{tikzpicture}

    \begin{axis}[
    xbar stacked,   % Stacked horizontal bars
    xmin=0,         % Start x axis at 0
    ytick=data,     % Use as many tick labels as y coordinates
    legend style={at={(axis cs:65,0.2)},anchor=south west},
    yticklabels from table={\testdata}{Label}  % Get the labels from the Label column of the \datatable
    ,bar width=1ex, y=3ex
    ]
    \addplot [fill=green!80] table [x=Adequacy, meta=Label,y expr=\coordindex] {\testdata};   % "First" column against the data index
    \addplot [fill=blue!60] table [x=Sustainability, meta=Label,y expr=\coordindex] {\testdata};
    \addplot [fill=red!60,
    point meta=x,
    nodes near coords,
    nodes near coords align={anchor=west},
    every node near coord/.append style={
        black,
        fill=white,
        fill opacity=0.75,
        text opacity=1,
        outer sep=\pgflinewidth % so the label fill doesn't overlap the plot
    }
    ] table [x=Integrity, meta=Label,y expr=\coordindex] {\testdata};
    \legend{Adequacy,Sustainability,Integrity}

    \end{axis}
    \end{tikzpicture}
\end{document}

### 2017-01-07 paper.tex
Section 2 2nd iteration + back to tikz

@@
 	framexleftmargin=1.5em,
+	backgroundcolor=\color{lightgrey}
 }
@@
 \usepackage{xcolor}
+
+\definecolor{lightgrey}{HTML}{FAFAFA}
 
@@
 
-\subsection{Workflow And Existing Tools} \label{sec:workflow}
-The work-flow for all existing tools is illustrated in Figure \ref{fig:approach}.
-Prior to any analysis, the web application is symbolically executed which for every execution path yields its corresponding client page output variant.
-Next, all client page output variants are parsed and merged into a compact representation of all client page output variants.
-This output model, also referred to as \emph{VarDOM} (Variational Document Object Model), is similar to a DOM, yet it can contain symbolic value nodes and conditional node types to represent variants as alternatives.
+\subsection{Static Output Approximation for PHP} \label{sec:workflow}
+Symbolic execution is an analysis technique not dedicated to a particular programming language.
+For dynamic web applications, in our case for web applications written in PHP, it can be used to approximate all possible client page output.
+Before we present in detail existing tools that aim to ease development and maintenance of PHP web applications, we illustrate how output approximation using symbolic execution is used by starting with an example.
 
-\begin{figure*}
+\begin{figure}[t]
 	\centering
-	\begin{subfigure}[center]{0.49\textwidth}%%[multicols=1]
+	\begin{subfigure}[center]{0.48\textwidth}%%[multicols=1]
 \begin{lstlisting}
@@
 <?php
-  function titles($arg, $n) {
-    $is = array(1,2,3);
-    while ($n < shuffle($is)) {
+function make_titles($arg, $n) {
+  $is = shuffle(array(1,2,3));
+  while ($n < $is[0]) {
     echo "<h1>" . $arg . "</h1><br>";
-      $n += 1;
+    $n = $n + 1;
   }
@@
 
-  call_user_func("title", "Headline", 4);
+call_user_func("make_titles", "Headline", 4);
 
-  $name = $_POST['name'];
+$name = "Hello " . $_POST['name'] .  "!";
 if (!isset($name)) {
@@
 } else {
-    echo "Entered ". $name ."!";
+  echo $name;
 }
@@
 \end{lstlisting}
-		\caption{Snippet of server-side web application code}
+		\caption{Snippet of HTML and PHP server-side web application code. \label{fig:example_code}}
     \end{subfigure}
+
+	\vspace{2mm}
+
     ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
       %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[center]{0.49\textwidth}
+    \begin{subfigure}[center]{0.48\textwidth}
 \begin{lstlisting}[mathescape]
@@
 \end{lstlisting}
-		\caption{Symbolic output approximation}
+		\caption{Approximation client paege output of the server-side web application code in Figure \ref{fig:example_code}.
+Variability as well as repetition annotations are represented by preprocessor directives.\label{fig:output_example}}
     \end{subfigure}
@@
     execution}
-\end{figure*}
+\end{figure}
 
@@
 
+Figure \ref{fig:} shows two code listings.
+The snippet of server-side web application code in Figure \ref{} contains both HTML and PHP.
+First, the function make\_titles defined in lines XX – XY prints the string arg passed as an argument between zero and an arbitrary number of times.
+Second, the function is indirectly called using the built-in function call\_user\_func in line XX with the string “Headline” and 4 as arguments.
+Finally, the input from the form in lines 1 – 4 is used: The variable greeting consists of three concatenated strings, “Hello ” the value of input field “name” and an exclamation mark.
+For this server-side code snippet there are two possible client page output variants: If the input field “name” is empty at runtime, line XX is executed and “No name entered!” is printed, or, if a name was entered, line XY is executed and the greeting constructed in line XZ is printed.
+
+The following symbolic execution semantics for PHP features  based on [BCG] illustrate the necessary modifications made to handle the extent of ambiguity that comes with handling symbolic values.
+In addition to normal output, the output approximation contains preprocessor-like directives to express repeatable output as well as output variants.
+All approximated client page output corresponding to Figure \ref{1a} consisting of two different variants is illustrated in Figure \ref{1b}.
+
+First, loops in the previous execution semantics are only executed once instead of an arbitrary number of times.
+Since a loop condition can contain symbolic values, it might not be feasible to determine when a loop terminates.
+Hence, any output constructed during the single loop iteration is highlighted (repeat … endrepeat) to possibly be repeated an arbitrary number of times.
+The output in line XX is marked to be repeated an arbitrary number of times, as the number of loop iterations cannot be determined statically.
+Second, a similar procedure applies to recursive functions.
+Once called, a recursive function is  also only executed to recursion depth one per call, and any subsequent invocation of that function returns a symbolic value.
+This, again, is due to the difficulty to determine the exact recursion depth similar to the number of loop iterations.
+Third, since the symbolic execution engine applies a static analysis, any form of user interaction providing input data as well as any form of input referred to from the deployment context of the
+Symbolic execution is an analysis technique not dedicated to a particular programming language.
+For dynamic web applications, in our case for web applications written in PHP, it can be used to approximate all possible client page output.
+Before we present in detail existing tools that aim to ease development and maintenance of PHP web applications, we illustrate how output approximation using symbolic execution is used by starting with an example.
+
+B. Explanation of Figure 1
+Figure \ref{fig:} shows two code listings.
+The snippet of server-side web application code in Figure \ref{} contains both HTML and PHP.
+First, the function make\_titles defined in lines XX – XY prints the string \$arg passed as an argument between zero and an arbitrary number of times.
+Second, the function is indirectly called using the built-in function call\_user\_func in line XX with the string “Headline” and 4 as arguments.
+Finally, the input from the form in lines 1 – 4 is used: The variable \$greeting consists of three concatenated strings, “Hello ” the value of input field “name” and an exclamation mark.
+For this server-side code snippet there are two possible client page output variants: If the input field “name” is empty at runtime, line XX is executed and “No name entered!” is printed, or, if a name was entered, line XY is executed and the greeting constructed in line XZ is printed.
+
+C. Introduction to symbolic execution of PHP code
+The following symbolic execution semantics for PHP features  based on [BCG] illustrate the necessary modifications made to handle the extent of ambiguity that comes with handling symbolic values.
+In addition to normal output, the output approximation contains preprocessor-like directives to express repeatable output as well as output variants.
+All approximated client page output corresponding to Figure \ref{1a} consisting of two different variants is illustrated in Figure \ref{1b}.
+First, loops in the previous execution semantics are only executed once instead of an arbitrary number of times.
+Since a loop condition can contain symbolic values, it might not be feasible to determine when a loop terminates.
+Hence, any output constructed during the single loop iteration is highlighted (repeat … endrepeat) to possibly be repeated an arbitrary number of times.
+The output in line XX is marked to be repeated an arbitrary number of times, as the number of loop iterations cannot be determined statically.
+Second, a similar procedure applies to recursive functions.
+Once called, a recursive function is  also only executed to recursion depth one per call, and any subsequent invocation of that function returns a symbolic value.
+This, again, is due to the difficulty to determine the exact recursion depth similar to the number of loop iterations.
+Third, since the symbolic execution engine applies a static analysis, any form of user interaction providing input data as well as any form of input referred to from the deployment context of the web application, i.e.
+configuration files or databases, is substituted by a symbolic value.
+The context variable \$\_POST represents input submitted to the server and, hence, is symbolic as every program input is per definition substituted by symbolic values.
+Finally, the output in lines and XX and XY respectively are marked as alternatives (if … else … endif), as symbolic execution of lines XX – YY in Figure 1a will explore both different branches.
+
+\subsection{Existing Tool Support for PHP}
 All tools proposed so far leverage a representation of all possible HTML client page output.
@@
 
-\subsection{Assumptions and Limitations}
-Symbolic execution remains an analysis technique not dedicated to a particular programming language.
+\textcolor{blue}{Bener}
+
+%\subsection{Assumptions and Limitations}
+%Symbolic execution remains an analysis technique not dedicated to a particular programming language.
 For the implementation of the symbolic execution engine in \cite{Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872}, several assumptions have been made to adapt to the PHP language.
@@
 
-First, loops in the previous execution semantics are only executed once instead of an arbitrary number of times.
+%First, loops in the previous execution semantics are only executed once instead of an arbitrary number of times.
 Since a loop condition can contain symbolic values, it might not be feasible to determine when a loop terminates.
@@
 
-Language-wise the symbolic execution semantics for the tools presented in Section \ref{sec:workflow} currently does not support object-oriented programming.
+%Language-wise the symbolic execution semantics for the tools presented in Section \ref{sec:workflow} currently does not support object-oriented programming.
 PHP provides a large number of built-in functions to manipulate strings as well as arrays.
@@
 
+% CALLBACK CANDIDATE ACCESSIBILITY ANALYSIS
+\begin{figure}[p]
+	\input{diagrams/context_distribution}
+	\caption{Distribution dead function output candidates that are partially or
+	completely explained due to callback candidates.}
+	\label{fig:coveragestuff}
+\end{figure}
+
 \section{Lessons Learned}

### 2017-01-08 paper.tex
Refactored sec 2 + new tikzed diagrams

@@
 \usepackage{xcolor}
+%\usepackage[dvipsnames]{xcolor}
 
-\definecolor{lightgrey}{HTML}{FAFAFA}
+\definecolor{lightgrey}{HTML}{EEEEEE}
 
@@
 result in variational output, representing dependencies of output and input.
-As symbolic execution of a program returns an input-output-mapping for symbolic inputs, this analysis technique helps unfold the \emph{staged} nature of web applications: Since every feasible path is executed all possible output is contained in the symbolic output, and all dynamically generated variational parts of the web application can be analyzed.
+As symbolic execution of a program returns an input-output-mapping for symbolic
+inputs, this analysis technique helps unfold the \emph{staged} nature of web
+applications: Since every feasible path is executed all possible output is
+contained in the symbolic output, and all dynamically generated variational
+parts of the web application can be analyzed.
 
@@
 
-During introduction of new semantics for object-orientation we encountered two main challenges. 
-First, we were bound to introduce more concrete execution features to the symbolic execution engine, e.g.
-foreach loops for arrays, as this increased code coverage significantly.
-Second, we had to make a trade-off between the accuracy
-of method calls and the scalability of this language feature as a method call
-for different program state variants can have multiple targets.
+During introduction of new semantics for object-orientation we encountered two
+main challenges. First, we were bound to introduce more concrete execution features to the
+symbolic execution engine, e.g. foreach loops for arrays, as this increased
+code coverage significantly. Second, we had to make a trade-off between the
+accuracy of method calls and the scalability of this language feature as a
+method call for different program state variants can have multiple targets.
 
@@
 
-Our key contributions in this paper include (1) a new tool infrastructure to statically approximate the output for PHP web applications using symbolic execution, and (2) a report of our observations with state-of-the-art web applications as well as an explanation on conceptual limitations for symbolic execution of PHP applications.
+Our key contributions in this paper include (1) a new tool infrastructure to
+statically approximate the output for PHP web applications using symbolic
+execution, and (2) a report of our observations with state-of-the-art web
+applications as well as an explanation on conceptual limitations for symbolic
+execution of PHP applications.
 
 \section{State Of the Art}\label{sec:stateoftheart}
-Symbolic execution is one of integral analysis techniques used by the tools whose practicality we aim to evaluate.
-This section recaps the idea of symbolic execution, the tools which make use of it, and both conceptual and language-related limitations.
+Symbolic execution is one of integral analysis techniques used by the tools
+whose practicality we aim to evaluate. This section recaps the idea of symbolic
+execution, the tools which make use of it, and both conceptual and
+language-related limitations.
 
@@
 Symbolic execution is a static analysis technique that was proposed by James C.
-King in 1976 \cite{King1976}.
-Symbolic execution allows to explore all feasible paths in a program.
-Based on normal program execution, the execution semantics is extended to handle symbolic values, which are abstract, but fixed; an execution becomes symbolic by introducing symbolic values as program inputs instead of concrete values \cite{King1976,Darringer1978}.
-These symbolic values are then propagated and used throughout the execution.
+King in 1976 \cite{King1976}. Symbolic execution allows to explore all feasible
+paths in a program. Based on normal program execution, the execution semantics
+is extended to handle symbolic values, which are abstract, but fixed; an
+execution becomes symbolic by introducing symbolic values as program inputs
+instead of concrete values \cite{King1976,Darringer1978}. These symbolic values
+are then propagated and used throughout the execution.
 
-Execution starts with a plain path condition, usually a tautology.
-Whenever the control-flow can be split into different branches, the corresponding path conditions (guards for instance) are conjoined with the previous path condition \cite{King1976}.
-Additionally, the conjoined path condition can be checked for satisfiability to exclude infeasible paths.
-Consequently, for a given program symbolic execution computes a mapping from (symbolic) input values to output values.
+Execution starts with a plain path condition, usually a tautology. Whenever the
+control-flow can be split into different branches, the corresponding path
+conditions (guards for instance) are conjoined with the previous path condition
+\cite{King1976}. Additionally, the conjoined path condition can be checked for
+satisfiability to exclude infeasible paths. Consequently, for a given program
+symbolic execution computes a mapping from (symbolic) input values to output values.
 
 \subsection{Static Output Approximation for PHP} \label{sec:workflow}
-Symbolic execution is an analysis technique not dedicated to a particular programming language.
-For dynamic web applications, in our case for web applications written in PHP, it can be used to approximate all possible client page output.
-Before we present in detail existing tools that aim to ease development and maintenance of PHP web applications, we illustrate how output approximation using symbolic execution is used by starting with an example.
+Symbolic execution is an analysis technique not dedicated to a particular
+programming language. For dynamic web applications, in our case for web
+applications written in PHP, it can be used to approximate all possible client
+page output. Before we present in detail existing tools that aim to ease
+development and maintenance of PHP web applications, we illustrate how output
+approximation using symbolic execution is used by starting with an example.
 
@@
   while ($n < $is[0]) {
-    echo "<h1>" . $arg . "</h1><br>";
+    echo "<h1>" . $arg . "</h1><br />";
     $n = $n + 1;
@@
 
-$name = "Hello " . $_POST['name'] .  "!";
-if (!isset($name)) {
+$greeting = "Hello " . $_POST['name'] .  "!";
+if (!isset($_POST['name'])) {
   echo "No name entered!";
 } else {
-  echo $name;
+  echo $greeting;
 }
@@
 \end{lstlisting}
-		\caption{Snippet of HTML and PHP server-side web application code. \label{fig:example_code}}
+		\caption{Snippet of HTML and PHP server-side web application code.
+		\label{fig:web_application}}
     \end{subfigure}
@@
 // #repeat n < shuffle(is)
-  <h1>Headline</h1>
+<h1>Headline</h1><br />
 // #endrepeat
@@
 \end{lstlisting}
-		\caption{Approximation client paege output of the server-side web application code in Figure \ref{fig:example_code}.
-Variability as well as repetition annotations are represented by preprocessor directives.\label{fig:output_example}}
+		\caption{
+			Approximation client paege output of the server-side web application code in
+			Figure \ref{fig:web_application}. Variability as well as repetition annotations
+			are represented by preprocessor directives.
+			\label{fig:output_example}
+		}
     \end{subfigure}
     \vspace{2mm}
-    \caption{Illustration of static ouptut approximation using symbolic
-    execution}
+    \caption{
+    	Illustration of static ouptut approximation using symbolic
+    	execution
+    	\label{fig:code_example}
+    }
 \end{figure}
 
-% Workflow Graphic
-%\begin{figure*}[t!]
-%	\centerline{
-%	\includegraphics[width=0.95\textwidth]{images-paper/workflow}
-%	}
-%	\vspace{2mm}
-%	\caption{
-%		Overview of the workflow for tools based on static output approximation for
-%		% dynamic web applications.
-%	}
-%	\label{fig:approach}
-%\end{figure*}
+Figure \ref{fig:code_example} shows two code listings. The snippet of server-side web
+application code in Figure \ref{fig:web_application} contains both HTML and PHP. First, the
+function \texttt{make\_titles} defined in lines 7 – 13 prints the string
+\texttt{\$arg} passed as an argument between zero and an arbitrary number of
+times. Second, the function is indirectly called using the built-in function
+\texttt{call\_user\_func} in line 15 with the string \texttt{Headline} and
+\texttt{4} as arguments.  Finally, the input from the form in lines 1 – 4 is
+used: The variable \texttt{\$greeting} consists of three concatenated strings,
+\texttt{Hello} the value of input field \texttt{name}, and an exclamation mark.
+For this server-side code snippet there are two possible client page output
+variants: If the input field \texttt{name} is empty at runtime, line 19 is
+executed and \texttt{No name entered!} is printed, or, if a name was entered,
+line 21 is executed and the greeting constructed in line 17 is printed.
 
-Figure \ref{fig:} shows two code listings.
-The snippet of server-side web application code in Figure \ref{} contains both HTML and PHP.
-First, the function make\_titles defined in lines XX – XY prints the string arg passed as an argument between zero and an arbitrary number of times.
-Second, the function is indirectly called using the built-in function call\_user\_func in line XX with the string “Headline” and 4 as arguments.
-Finally, the input from the form in lines 1 – 4 is used: The variable greeting consists of three concatenated strings, “Hello ” the value of input field “name” and an exclamation mark.
-For this server-side code snippet there are two possible client page output variants: If the input field “name” is empty at runtime, line XX is executed and “No name entered!” is printed, or, if a name was entered, line XY is executed and the greeting constructed in line XZ is printed.
-
-The following symbolic execution semantics for PHP features  based on [BCG] illustrate the necessary modifications made to handle the extent of ambiguity that comes with handling symbolic values.
-In addition to normal output, the output approximation contains preprocessor-like directives to express repeatable output as well as output variants.
-All approximated client page output corresponding to Figure \ref{1a} consisting of two different variants is illustrated in Figure \ref{1b}.
-
-First, loops in the previous execution semantics are only executed once instead of an arbitrary number of times.
-Since a loop condition can contain symbolic values, it might not be feasible to determine when a loop terminates.
-Hence, any output constructed during the single loop iteration is highlighted (repeat … endrepeat) to possibly be repeated an arbitrary number of times.
-The output in line XX is marked to be repeated an arbitrary number of times, as the number of loop iterations cannot be determined statically.
-Second, a similar procedure applies to recursive functions.
-Once called, a recursive function is  also only executed to recursion depth one per call, and any subsequent invocation of that function returns a symbolic value.
-This, again, is due to the difficulty to determine the exact recursion depth similar to the number of loop iterations.
-Third, since the symbolic execution engine applies a static analysis, any form of user interaction providing input data as well as any form of input referred to from the deployment context of the
-Symbolic execution is an analysis technique not dedicated to a particular programming language.
-For dynamic web applications, in our case for web applications written in PHP, it can be used to approximate all possible client page output.
-Before we present in detail existing tools that aim to ease development and maintenance of PHP web applications, we illustrate how output approximation using symbolic execution is used by starting with an example.
-
-B. Explanation of Figure 1
-Figure \ref{fig:} shows two code listings.
-The snippet of server-side web application code in Figure \ref{} contains both HTML and PHP.
-First, the function make\_titles defined in lines XX – XY prints the string \$arg passed as an argument between zero and an arbitrary number of times.
-Second, the function is indirectly called using the built-in function call\_user\_func in line XX with the string “Headline” and 4 as arguments.
-Finally, the input from the form in lines 1 – 4 is used: The variable \$greeting consists of three concatenated strings, “Hello ” the value of input field “name” and an exclamation mark.
-For this server-side code snippet there are two possible client page output variants: If the input field “name” is empty at runtime, line XX is executed and “No name entered!” is printed, or, if a name was entered, line XY is executed and the greeting constructed in line XZ is printed.
-
-C. Introduction to symbolic execution of PHP code
-The following symbolic execution semantics for PHP features  based on [BCG] illustrate the necessary modifications made to handle the extent of ambiguity that comes with handling symbolic values.
-In addition to normal output, the output approximation contains preprocessor-like directives to express repeatable output as well as output variants.
-All approximated client page output corresponding to Figure \ref{1a} consisting of two different variants is illustrated in Figure \ref{1b}.
-First, loops in the previous execution semantics are only executed once instead of an arbitrary number of times.
-Since a loop condition can contain symbolic values, it might not be feasible to determine when a loop terminates.
-Hence, any output constructed during the single loop iteration is highlighted (repeat … endrepeat) to possibly be repeated an arbitrary number of times.
-The output in line XX is marked to be repeated an arbitrary number of times, as the number of loop iterations cannot be determined statically.
-Second, a similar procedure applies to recursive functions.
-Once called, a recursive function is  also only executed to recursion depth one per call, and any subsequent invocation of that function returns a symbolic value.
-This, again, is due to the difficulty to determine the exact recursion depth similar to the number of loop iterations.
-Third, since the symbolic execution engine applies a static analysis, any form of user interaction providing input data as well as any form of input referred to from the deployment context of the web application, i.e.
-configuration files or databases, is substituted by a symbolic value.
-The context variable \$\_POST represents input submitted to the server and, hence, is symbolic as every program input is per definition substituted by symbolic values.
-Finally, the output in lines and XX and XY respectively are marked as alternatives (if … else … endif), as symbolic execution of lines XX – YY in Figure 1a will explore both different branches.
+%C. Introduction to symbolic execution of PHP code
+The following symbolic execution semantics for PHP features  based on \cite{Nguyen:2014:BCG:2635868.2635928}
+illustrate the necessary modifications made to handle the extent of ambiguity
+that comes with handling symbolic values. In addition to normal output, the
+output approximation contains preprocessor directives to express
+repeatable output as well as output variants. All approximated client page
+output corresponding to Figure \ref{fig:web_application} consisting of two different variants
+is illustrated in Figure \ref{fig:output_example}.
+First, loops in the previous execution semantics are only executed once instead
+of an arbitrary number of times. Since a loop condition can contain symbolic
+values, it might not be feasible to determine when a loop terminates. Hence,
+any output constructed during the single loop iteration is highlighted
+(\texttt{repeat \ldots  endrepeat}) to possibly be repeated an arbitrary number
+of times. The output in line 7 is marked to be repeated an arbitrary number of times, as the number
+of loop iterations cannot be determined statically. Second, a similar procedure
+applies to recursive functions. Once called, a recursive function is  also only
+executed to recursion depth one per call, and any subsequent invocation of that
+function returns a symbolic value. This, again, is due to the difficulty to
+determine the exact recursion depth similar to the number of loop iterations.
+Third, since the symbolic execution engine applies a static analysis, any form
+of user interaction providing input data as well as any form of input referred
+to from the deployment context of the web application, i.e. configuration files
+or databases, is substituted by a symbolic value. The context variable
+\texttt{\$\_POST} represents input submitted to the server and, hence, is
+symbolic as every program input is per definition substituted by symbolic values. Finally, the
+output in lines 11 and 13 respectively are marked as alternatives
+(\texttt{if \ldots else \ldots endif}), as symbolic execution of lines 18 – 22
+in Figure \ref{fig:web_application} will explore both different branches.
 
 \subsection{Existing Tool Support for PHP}
-All tools proposed so far leverage a representation of all possible HTML client page output.
-Since any variant can be analyzed in particular, this enables tool support addressing the web application as a whole in spite of its staged nature, where parts of the web application (client-side output) are generated dynamically.
+All tools proposed so far leverage a representation of all possible HTML client
+page output. Since any variant can be analyzed in particular, this enables tool
+support addressing the web application as a whole in spite of its staged
+nature, where parts of the web application (client-side output) are generated
+dynamically.
 
-Based on the approximated output representation subsequent analyses can be conducted, which have been implemented in the following tools:
+Based on the approximated output representation subsequent analyses can be
+conducted, which have been implemented in the following tools:
 
@@
 
-\textcolor{blue}{Bener}
+All three tools though have only been applied to systems that are small and not
+maintained any more. Despite the functional benefits these tools provide to
+developers, their practicality for large and modern systems has not been
+studied.
 
@@
 \section{Scalable Output Approximation}%
-The main objective of this experience report is to investigate whether the tools described in Section \ref{sec:stateoftheart} are practical.
-In particular we want to find out whether symbolic execution as the underlying technique for static output approximation scales for PHP, and what may limitations restrict scalability.
-We may divide possible limitations into either matters of engineering or conceptual limitations where no precise execution is possible.
+The main objective of this experience report is to investigate whether the
+tools described in Section \ref{sec:stateoftheart} are practical. In particular
+we want to find out whether symbolic execution as the underlying technique for
+static output approximation scales for PHP, and what may limitations restrict
+scalability. We may divide possible limitations into either matters of
+engineering or conceptual limitations where no precise execution is possible.
 
-In the following we describe in detail our implementation procedure, practical problems we encountered and whether these are limiting scalability conceptually or just require engineering effort to be tamed.
+In the following we describe in detail our implementation procedure, practical
+problems we encountered and whether these are limiting scalability conceptually
+or just require en gineering effort to be tamed.
 
 \subsection{Towards Scalable Output Approximation?}
-We re-implement\footnote{\url{https://github.com/smba/oak}} the existing symbolic execution semantics of \cite{Nguyen:2014:BCG:2635868.2635928} with the goal to support modern large-scale dynamic web applications, or at least to better understand possible limitations.
-For reference, we choose to aim to support  \textsf{WordPress}\footnote{\url{https://github.com/WordPress/WordPress}}, a popular Content Management System of reasonable size that is still maintained.
-Moreover, our implementation uses parts of Quercus\footnote{\url{http://quercus.caucho.com/}}, a pure Java implementation of the PHP interpreter, which is tested against the \textsf{WordPress} test suite.
-We choose a test-driven approach for our implementation with regression tests and continuous integration.
-Our test cases are based on both custom examples for different language features as well as \textsf{SchoolMate}, a small-scale system, and, at least in parts, \textsf{WordPress}, whose output we manually validate and then automatically re-check for each test run.
+We re-implement\footnote{\url{https://github.com/smba/oak}} the existing
+symbolic execution semantics of \cite{Nguyen:2014:BCG:2635868.2635928} with the
+goal to support modern large-scale dynamic web applications, or at least to
+better understand possible limitations. For reference, we choose to aim to
+support  \textsf{WordPress}
+\footnote{\url{https://github.com/WordPress/WordPress }}, a popular Content
+Management System of reasonable size that is still maintained. Moreover, our
+implementation uses parts of
+Quercus\footnote{\url{http://quercus.caucho.com/}}, a pure Java implementation
+of the PHP interpreter, which is tested against the \textsf{WordPress} test
+suite. We choose a test-driven approach for our implementation with regression
+tests and continuous integration. Our test cases are based on both custom
+examples for different language features as well as \textsf{SchoolMate}, a 
+small-scale system, and, at least in parts, \textsf{WordPress}, whose output we
+manually validate and then automatically re-check for each test run.
 
-Our implementation basically follows the symbolic execution semantics of \cite{Nguyen:2014:BCG:2635868.2635928}, yet we extend it with additional rules for class definitions, class instantiation, and method invocation to support object-oriented programming.
-Furthermore, we add rules for do-while and  foreach loops since these are not supported as yet.
-Note that we execute all iterations of a foreach loop since the number of iterations is known since we are traversing all items of a given array.
+Our implementation basically follows the symbolic execution semantics of
+\cite{Nguyen:2014:BCG:2635868.2635928}, yet we extend it with additional rules
+for class definitions, class instantiation, and method invocation to support
+object-oriented programming. Furthermore, we add rules for do-while and 
+foreach loops since these are not supported as yet. Note that we execute all
+iterations of a foreach loop since the number of iterations is known since we
+ are traversing all items of a given array.
 
@@
 \subsection{Measuring Approximation Success} \label{heuristic}
-For the evaluation of our output approximation we require a ground truth output model to compare our approximation against.
-Since output literals are embedded in server-side code and scattered, it is nontrivial to determine whether a string literal is part of the ouput or not.
-So, in the absence of ground truth for the approximated output we are bound to choose a \emph{heuristic approach} in order to measure the accuracy of our aproximation.
-We define a string literal to be an \emph{output candidate} if there exists an execution path reaching the string literal and evantually passing it to an
-output-generating statement (e.g., \texttt{echo} or \texttt{print}).
+For the evaluation of our output approximation we require a ground truth output
+model to compare our approximation against. Since output literals are embedded
+in server-side code and scattered, it is nontrivial to determine whether a
+string literal is part of the ouput or not. So, in the absence of ground truth
+for the approximated output we are bound to choose a \emph{heuristic approach}
+in order to measure the accuracy of our aproximation. We define a string
+literal to be an \emph{output candidate} if there exists an execution path
+reaching the string literal and evantually passing it to an output-generating
+statement (e.g., \texttt{echo} or \texttt{print}).
 
-We approximate output candidates as those string literals containing the characters \texttt{<}, \texttt{>}, or both since we expect output to contain HTML tags. We evaluated this heuristic manually
-with a sample of 400 string literals randomly selected from the entire corpus.
-We measured for our heuristic classifier a precision of 94 percent, and a recall
-of 50 percent.
-This means that six percent of the string literals are classified incorrectly as false positives.
-In turn, the classifier is highly distinctive as 96 percent of the string literals are classified correctly.
-The recall of 50 percent means that half of the string literals, which we manually classified as output, actually were responsive to the classifier.
-We attempted to increase recall by looking for further distinctive properties to build a classifier from, but could not do so without decreasing precision.
-In spite of missing half of the output candidates, we decided to use a simple, yet distinctive classifier.
+We approximate output candidates as those string literals containing the
+characters \texttt{<}, \texttt{>}, or both since we expect output to contain
+HTML tags. We evaluated this heuristic manually with a sample of 400 string
+literals randomly selected from the entire corpus.
+We measured for our heuristic classifier a precision of 94 percent, and a
+recall of 50 percent. This means that six percent of the string literals are
+classified incorrectly as false positives. In turn, the classifier is highly
+distinctive as 96 percent of the string literals are classified correctly. The
+recall of 50 percent means that half of the string literals, which we manually 
+classified as output, actually were responsive to the classifier. We attempted
+to increase recall by looking for further distinctive properties to build a
+classifier from, but could not do so without decreasing precision. In spite of
+missing half of the output candidates, we decided to use a simple, yet
+distinctive classifier.
 
 \subsection{Experiment Setup} 
-For the evaluation we symbolically executed a selection of PHP systems.
-We selected a corpus of twelve PHP systems with regard to system size and recency as the corpus of the case study for \cite{Nguyen:2014:BCG:2635868.2635928} only contained small-scale systems that are not maintained any more.
-The full list of PHP systems is shown in Table \ref{corpus}.
-Our selection of PHP systems includes
-%\begin{enumerate}[a)]
+For the evaluation we symbolically executed a selection of PHP systems. We
+selected a corpus of twelve PHP systems with regard to system size and recency
+as the corpus of the case study for \cite{Nguyen:2014:BCG:2635868.2635928} only
+contained small-scale systems that are not maintained any more. The full list
+of PHP systems is shown in Table \ref{corpus}. Our selection of PHP systems
+includes
+
+%
 \begin{compactitem}
-\item four small-scale systems that selected from the previously mentioned case study in order to compare our results with the previous symbolic execution engine \cite{Nguyen:2014:BCG:2635868.2635928},
-\item three recent large-scale systems selected from the case study corpus of \cite{Hills:2013:ESP:2483760.2483786}, and
-\item four small-scale systems selected from a list of recent Content Management Systems \cite{codegeekz}.
-%umerate}
+	\item four small-scale systems that selected from the previously mentioned case
+	study in order to compare our results with the previous symbolic execution
+	engine \cite{Nguyen:2014:BCG:2635868.2635928},
 
+	\item three recent large-scale systems selected from the case study corpus of
+	\cite{Hills:2013:ESP:2483760.2483786}, and
+
+	\item four small-scale systems selected from a list of recent Content
+	Management Systems \cite{codegeekz}.
 \end{compactitem}
+
 As for for the last two cases, we limited our selection since the parser we
@@
 \centering 
-	\begin{tabular}{lrp{4cm}rlrr}
+	\begin{tabular}{lrp{4cm}rrrr}
 	\toprule
-	\textbf{System} & \textbf{Version} & \textbf{Classification} & \textbf{SLOC} & \textbf{\#files} & \textbf{\#OCs} & \textbf{\#includes}
+	\textbf{System} & \textbf{Version} & \textbf{Classification} & \textbf{SLOC} &
+	\textbf{\#files} & \textbf{\#OCs} & \textbf{\#includes}
 	\\
 	\midrule
-	AddressBook & 8.2.5.2 & 239 & 51,907 & Adress Administration  & XXX & XXX\\
-	SchoolMate & 1.5.4 & 65 & 8,118 & School Administration  & XXX & XXX\\
-	TimeClock & 1.04 & 63 & 20,800 & XXX  & XXX & XXX\\
-	WebChess & 1.0.0 & 28 & 5,219 & XXX  & XXX & XXX\\
+	AddressBook & 8.2.5.2 &  & 51,907 & 239  & 1009 & 186\\
+	SchoolMate & 1.5.4 &  & 8,118 & 65  & 853 & 88\\
+	TimeClock & 1.04 &  & 20,800 & 63  & 7920 & 306\\
+	WebChess & 1.0.0 &  & 5,219 & 28  & 470 & 56\\
 	\midrule
-	Drupal & 7.5.0 & 125 & 52,464 & Content Management System  & XXX & XXX\\
-	phpBB & 3.1.9 & 1,398 & 327,371 & Bulletin Board  & XXX & XXX\\
-	phpMyAdmin & 4.6.3 & 871 & 303,582 & Database Administration  & XXX & XXX\\
+	Drupal & 7.5.0 & CMS & 52,464 & 125  & 3569 & 749\\
+	phpBB & 3.1.9 &  & 327,371 & 1,398  & 3606 & 206\\
+	phpMyAdmin & 4.6.3 &  & 303,582 & 871  & 7103 & 571\\
 	\midrule
-	Anchor & 0.12.1 & 201 & 15,054 & Content Management System  & XXX & XXX\\
-	Kirby & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
-	Automad & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
-	Monstra & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
-	Nibbleblog & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
+	Anchor & 0.12.1 &  & 15,054 & 201 & 987 & 32\\
+	Kirby & XXX & CMS & XXX & XXX  & 654 & 23\\
+	Automad & XXX & CMS & XXX & XXX  & 655 & 8\\
+	Monstra & XXX & CMS & XXX & XXX  & 1934 & 48\\
+	Nibbleblog & XXX & CMS & XXX & XXX  & 1013 & 28\\
 	\bottomrule
@@
 \label{HowAccurateIsOurApproximation} 
-In Section \ref{heuristic} we introduced the definition of output candidates as expected output.
-As a most accurate approximation contains all output candidates, we define two metrics to measure accuracy of our approximation.
+In Section \ref{heuristic} we introduced the definition of output candidates as
+expected output. As a most accurate approximation contains all output
+candidates, we define two metrics to measure accuracy of our approximation.
 
-First, we measure how much of the expected output was actually processed by the symbolic interpreter.
-Once a line of code, a statement or expression containing an output candidate is actually an element in an execution path, we define this output candidate as reached.
-We define the metric \emph{reach coverage} as the ratio of output candidates that are reached and the total number of output candidates in the analyzed system.
-Although a reached output candidate is processed by the symbolic interpreter, it does not guarantee we will see that output candidate in the symbolic output.
-Second, we define the metric \emph{output coverage} as the ratio of output candidates that are contained in the output of the symbolic interpreter and the total number of output candidates in the analyzed system.
-For an imprecise approximation, we may see loss of output candidate information resulting in a output coverage lower than the reach coverage.
+First, we measure how much of the expected output was actually processed by the
+symbolic interpreter.  Once a line of code, a statement or expression
+containing an output candidate is actually an element in an execution path, we
+define this output candidate as reached. We define the metric \emph{reach
+coverage} as the ratio of output candidates that are reached and the total
+number of output candidates in the analyzed system. Although a reached output
+candidate is processed by the symbolic interpreter, it does not guarantee we
+will see that output candidate in the symbolic output. Second, we define the
+metric \emph{output coverage} as the ratio of output candidates that are
+contained in the output of the symbolic interpreter and the total number of
+output candidates in the analyzed system.  For an imprecise approximation, we
+may see loss of output candidate information resulting in a output coverage
+lower than the reach coverage.
 
-The coverage results are illustrated in Figure \ref{coverage}.
-We could replicate high coverage results for the four small-scale systems with a reach coverage and output coverage over 80 percent respectively.
-For the three modern/large-scale systems we measured poor coverage with reach coverage ranging from 5 to 30 percent and output coverage ranging from 4 to 20 percent.
-We measured medium to high coverage for the more recent small-scale systems: reach coverage ranging from 43 to 85 percent, output coverage ranging from 40 to 85 percent.
-% RESULTS
-% REACH COVERAGE and
-% OUTPUT COVERAGE
-\begin{figure}
-	\centering	
-%	\resizebox{0.4\textwidth}{!}{%
-%		\input{diagrams/coverage}	
-%	}
-	\includegraphics[width=0.5\textwidth]{images-paper/coverage.eps}
-	\caption{
-		Results for output candidates: \emph{Coverage loss} repesents reached output candidates that were not part of the output.
-\emph{Simple file} and \emph{function definition} represent context classifications of missed output candidates.
-	}
-	\label{coverage}
-\end{figure}
+The coverage results are illustrated in Figure \ref{coverage}. We could
+replicate high coverage results for the four small-scale systems with a reach
+coverage and output coverage over 80 percent respectively. For the three
+modern/large-scale systems we measured poor coverage with reach coverage
+ranging from 5 to 30 percent and output coverage ranging from 4 to 20 percent.
+We measured medium to high coverage for the more recent small-scale systems:
+reach coverage ranging from 43 to 85 percent, output coverage ranging from 40
+to 85 percent.
 
 \subsection{Understanding Limitations}
-As we have seen in Section \ref{HowAccurateIsOurApproximation} both measured coverage metrics were poor for large-scale and small-scale/recent systems.
-To understand what output candidates we missed and why, we further investigated our approximation results and conducted three more measurements.
+As we have seen in Section \ref{HowAccurateIsOurApproximation} both measured
+coverage metrics were poor for large-scale and small-scale/recent systems. To
+understand what output candidates we missed and why, we further investigated
+our approximation results and conducted three more measurements.
 
 \subsubsection{What literals did we miss?}\label{WhatLiteralsDidWeMiss?}
-Our initial approach to understand missed output candidates is to find out whether their surrounding program code was not accessed/accessible, and if so, why.
-Of special interest are those output candidates that are located in HTML files (possibly with nested PHP), or located in a function definition.
-Plain HTML files do not require any further execution (of course, unless PHP scripts are embedded) and represent output candidates that just need to be included properly to be part of the output.
-Aside, output candidates that are part of a function definition are only missed if the corresponding function is never called at runtime.
+Our initial approach to understand missed output candidates is to find out
+whether their surrounding program code was not accessed/accessible, and if so,
+why. Of special interest are those output candidates that are located in HTML
+files (possibly with nested PHP), or located in a function definition. Plain
+HTML files do not require any further execution (of course, unless PHP scripts
+are embedded) and represent output candidates that just need to be included
+properly to be part of the output. Aside, output candidates that are part of a
+function definition are only missed if the corresponding function is never
+called at runtime.
 
-We started classifying missed output candidates by their string literal context.
-As the classification statistics in Figure \ref{coverage} illustrate, for almost all systems (except for \textsf{AddressBook} and \textsf{TimeClock}) not-reached output candidates had a function context.
-Note that \textsf{SchoolMate} was excluded from the diagram since our analysis reached all output candidates for that system.
-
-% CONTEXT DISTRIBUTION
-% - FUNCTION CONTEXT
-% - TEMPLATE CONTEXT
-%\begin{figure}
-%	\resizebox{0.4\textwidth}{!}{%
-%		\input{diagrams/context_distribution}
-%	}	
-%	\caption{Context distribution of never reached output candidates for the PHP
-%	corpus.}
-%	\label{fig:contexts}
-%\end{figure}
+We started classifying missed output candidates by their string literal
+context. As the classification statistics in Figure \ref{coverage} illustrate,
+for almost all systems (except for \textsf{AddressBook} and \textsf{TimeClock})
+not-reached output candidates had a function context. Note that
+\textsf{SchoolMate} was excluded from the diagram since our analysis reached
+all output candidates for that system.
 
@@
 \label{sec:inaccessible}
-For both cases of context, either inclusion of a script file failed, or a function call failed.
-If an HTML file is not part of the output, it also is never reached, i.e., included.
-An include can fail due to various reasons, such as an imprecise evaluation of the include expression returning a symbolic value, or simply a missing file.
+For both cases of context, either inclusion of a script file failed, or a
+function call failed. If an HTML file is not part of the output, it also is
+never reached, i.e., included. An include can fail due to various reasons, such
+as an imprecise evaluation of the include expression returning a symbolic
+value, or simply a missing file.
 
-For a function to be never reached there are several scenarios: A function is undefined at runtime if the corresponding script file is never included.
-In turn, if the function is defined at runtime the function can either be dead code if there is no call site for this function, or the function call failed.
-In PHP there are several ways to call a function beside direct call sites.
-The language offers indirect call mechanisms like callback-commands or an evaluation function that parses and evaluates PHP source code as strings.
+For a function to be never reached there are several scenarios: A function is
+undefined at runtime if the corresponding script file is never included. In
+turn, if the function is defined at runtime the function can either be dead
+code if there is no call site for this function, or the function call failed.
+In PHP there are several ways to call a function beside direct call sites. The
+language offers indirect call mechanisms like callback-commands or an
+evaluation function that parses and evaluates PHP source code as strings.
 
-Given a symbolic value, indirect call mechanisms like callback-commands are likely to fail since target function name and the symbolic value do not match.
-This also applies for the evaluation of include expressions as concrete values representing include targets can be included, but any include target containing symbolic values is ambiguous.
+Given a symbolic value, indirect call mechanisms like callback-commands are
+likely to fail since target function name and the symbolic value do not match.
+This also applies for the evaluation of include expressions as concrete values
+representing include targets can be included, but any include target containing
+symbolic values is ambiguous.
 
@@
 \label{WhyDidIncludesFail}
-To better understand include expression evaluation, we use another metric to approach accuracy from a different angle: We measure (1) the ratio of reached include expressions and the total number of include expressions in a system respectively (coverage of include expressions), and (2) the ratio of successfully resolved include expressions and reached include expressions in a system (resolution success rate).
-For a most accurate result all include expressions are reached and resolved.
+To better understand include expression evaluation, we use another metric to
+approach accuracy from a different angle: We measure (1) the ratio of reached
+include expressions and the total number of include expressions in a system
+respectively (coverage of include expressions), and (2) the ratio of
+successfully resolved include expressions and reached include expressions in a
+system (resolution success rate). For a most accurate result all include
+expressions are reached and resolved.
 
-For include expressions we sampled failed attempts and manually evaluated the cause of failure.
-We identified for (1) that all failed attempts were dynamic include expressions, i.e., those which were resolved at runtime and were assembled from fragments of environment-dependent information including database retrievals, user input, or files.
-Consequently, the resolved include expressions contained symbolic values which did not represent identifiable script files.
+For include expressions we sampled failed attempts and manually evaluated the
+cause of failure. We identified for (1) that all failed attempts were dynamic
+include expressions, i.e., those which were resolved at runtime and were
+assembled from fragments of environment-dependent information including
+database retrievals, user input, or files. Consequently, the resolved include
+expressions contained symbolic values which did not represent identifiable
+script files.
 
-Finally, we measured for include expressions the overall reach coverage and the overall resolution success rate per system.
-As illustrated in Figure \ref{fig:include_coverage_results}, except for the small-scale systems and two recent small-scale systems we did not even really reach a great portion of include expressions, although for most systems the resolution success rate was over 80 percent.
-Nevertheless, one failed attempt to include a script file can result in a cascade of missing more include expressions as the include target script is not executed.
-
-% INCLUDE COVERAGE / SUCCESS RATE
-\begin{figure}
-%    \centering
-%    \begin{subfigure}[t]{0.5 \textwidth}    	
-%        \centering        
-%        % Include Coverage Plot        %
-%		\input{diagrams/include_coverage}		
-%        \caption{Proportion of reached include expressions per project.}
-%        \label{fig:include_coverage_results1}
-%    \end{subfigure}
-%	\hfill
-%    \begin{subfigure}[t]{0.5\textwidth}    
-%        \centering
-%        % Success Rate Plot        
-%        \input{diagrams/success_rate}        
-%        \caption{Rate of successful resolutions of reached include expressions
-%        per project.}
-%        \label{fig:include_coverage_results2}
-%    \end{subfigure}
-
-\includegraphics[width=0.5\textwidth]{images-paper/includes.eps}
-
-    \caption{Include coverage and accuracy.\label{fig:include_coverage_results}}
-\end{figure}
+Finally, we measured for include expressions the overall reach coverage and the
+overall resolution success rate per system. As illustrated in Figure
+\ref{fig:include_coverage_results}, except for the small-scale systems and two
+recent small-scale systems we did not even really reach a great portion of
+include expressions, although for most systems the resolution success rate was
+over 80 percent. Nevertheless, one failed attempt to include a script file can
+result in a cascade of missing more include expressions as the include target
+script is not executed.
 
@@
 \label{WhyDidFunctionCallsFail}
-Aside from failed or successful includes, this metric describes one scenario of why functions containing not-reached output candidates are never called: The trivial explanation would be that there are no direct calls of those functions, yet PHP offers a number of ways to call a function indirectly, for instance using built-in functions, or simply evaluating strings as PHP code using the \texttt{eval} function.
+Aside from failed or successful includes, this metric describes one scenario of
+why functions containing not-reached output candidates are never called: The
+trivial explanation would be that there are no direct calls of those functions,
+yet PHP offers a number of ways to call a function indirectly, for instance
+using built-in functions, or simply evaluating strings as PHP code using the
+\texttt{eval} function.
 
-Moreover we need to take into account that one missed function execution can result in missing even more function calls (direct or indirect) and executions.
-Thus, for functions containing not-reached output candidates we measure whether they can be (1) only, (2) partially or (3) never accessed transitively from functions that have no direct call sites.
-We refer to those functions (or access points) having no direct call site as callback candidates since the only way to reach them is through indirect mechanisms.
-Measuring how many not-reached output candidates can be explained by functions that are callback candidates helps to understand whether, and if so, indirect call mechanisms and their usage may impact our analysis’ code coverage.
+Moreover we need to take into account that one missed function execution can
+result in missing even more function calls (direct or indirect) and executions.
+Thus, for functions containing not-reached output candidates we measure whether
+they can be (1) only, (2) partially or (3) never accessed transitively from
+functions that have no direct call sites. We refer to those functions (or
+access points) having no direct call site as callback candidates since the only
+way to reach them is through indirect mechanisms. Measuring how many
+not-reached output candidates can be explained by functions that are callback
+candidates helps to understand whether, and if so, indirect call mechanisms and
+their usage may impact our analysis’ code coverage.
 
-The callgraph snippets in Figure \ref{fig:callgraph} illustrate the idea of detecting the functions’ access points: The nodes at the top depict callback candidates, the nodes at the bottom with labels X, Y, and Z depict function definitions containing not-reached output candidates.
-In Figure \ref{fig:callgraph} function X can only be accessed through callback candidates, as for Figure \ref{fig:callgraph} function Y can be partially accessed by callback candidates, whereas function Z is never accessible through callback candidates.
+The callgraph snippets in Figure \ref{fig:callgraph} illustrate the idea of
+detecting the functions’ access points: The nodes at the top depict callback
+candidates, the nodes at the bottom with labels X, Y, and Z depict function
+definitions containing not-reached output candidates. In Figure
+\ref{fig:callgraph} function X can only be accessed through callback
+candidates, as for Figure \ref{fig:callgraph} function Y can be partially
+accessed by callback candidates, whereas function Z is never accessible through
+callback candidates.
 
@@
 \begin{figure}[t!]
-	\centering
+	%\centering
 	\includegraphics[scale=0.55]{images-paper/callgraph_cases}
     \caption{Callgraph analysis: The bottom nodes represent functions containing missed output candidates, the top nodes depict entry points to them.}
-    \label{fig:callgraph}
+    \label{fig:callgraph} output candidates we measured how many of those functions were only, or partially accessibly through indirect calling mechanisms (see Section \ref{sec:inaccessible}).
+As Figure \ref{fig:output_candidate_explanation} illustrates, around 80 percent (except for \textsf{TimeClock}) all missed output candidates of each system were only accessible through callback candidates, i.e., functions that are only accessible through indirect calling mechanisms.
+
 \end{figure}
 
-Like for include expressions, we again sampled dead function cases and identified the lack of direct call sites to those functions.
-As described in Section \ref{sec:inaccessible}, functions despite having no direct call sites were called through indirect call features.
-We identified dynamically assembled expressions (function names) to be responsible for failed attempts of calling a function indirectly as the resolved function name eventually contained symbolic values.
-These expressions, as well as include expressions, contained information dependent on the system environment.
+Like for include expressions, we again sampled dead function cases and
+identified the lack of direct call sites to those functions. As described in
+Section \ref{sec:inaccessible}, functions despite having no direct call sites
+were called through indirect call features. We identified dynamically assembled
+expressions (function names) to be responsible for failed attempts of calling a
+function indirectly as the resolved function name eventually contained symbolic
+values. These expressions, as well as include expressions, contained
+information dependent on the system environment.
 
-For failed function calls to functions containing not-reached output candidates we measured how many of those functions were only, or partially accessibly through indirect calling mechanisms (see Section \ref{sec:inaccessible}).
-As Figure \ref{fig:output_candidate_explanation} illustrates, around 80 percent (except for \textsf{TimeClock}) all missed output candidates of each system were only accessible through callback candidates, i.e., functions that are only accessible through indirect calling mechanisms.
+For failed function calls to functions containing not-reached
 
-% CALLBACK CANDIDATE ACCESSIBILITY ANALYSIS
-\begin{figure}[p]
-	\input{diagrams/output_candidates}
-	\caption{Distribution dead function output candidates that are partially or
-	completely explained due to callback candidates.}
+	% RESULT DIAGRAMS
+\begin{figure}[h!]
+	\begin{subfigure}[center]{0.48\textwidth}
+		\input{new_diagrams/not_reached}
+		\caption{\label{coverage}}
+	\end{subfigure}
+
+	\begin{subfigure}[center]{0.48\textwidth}
+		\input{new_diagrams/includes}
+		\caption{
+			Include coverage and accuracy.
+			\label{fig:include_coverage_results}
+		}
+	\end{subfigure}
+
+	% Callback candidate explanation
+	\begin{subfigure}[center]{0.48\textwidth}
+		\input{new_diagrams/callbacks}
+		\caption{
+			Distribution dead function output candidates that are par-
+			tially or completely explained due to callback candidates.
 			\label{fig:output_candidate_explanation}
+		}
+
+	\end{subfigure}
+	\caption{Results}
 \end{figure}
-
-% CALLBACK CANDIDATE ACCESSIBILITY ANALYSIS
-\begin{figure}[p]
-	\input{diagrams/context_distribution}
-	\caption{Distribution dead function output candidates that are partially or
-	completely explained due to callback candidates.}
-	\label{fig:coveragestuff}
-\end{figure}
-
 \section{Lessons Learned}

### 2017-01-09 paper.tex
update
refactored sec 3

@@
 
-\subsection{Existing Tool Support for PHP}
+\subsection{Existing Tool Support for PHP}\label{sec:existing_tools}
 All tools proposed so far leverage a representation of all possible HTML client
@@
 
-All three tools though have only been applied to systems that are small and not
-maintained any more. Despite the functional benefits these tools provide to
-developers, their practicality for large and modern systems has not been
-studied.
-
-%\subsection{Assumptions and Limitations}
-%Symbolic execution remains an analysis technique not dedicated to a particular programming language.
-For the implementation of the symbolic execution engine in \cite{Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872}, several assumptions have been made to adapt to the PHP language.
-The following symbolic execution semantics for PHP features differ from concrete execution since symbolic values need to be handled without ambiguity.
-
-%First, loops in the previous execution semantics are only executed once instead of an arbitrary number of times.
-Since a loop condition can contain symbolic values, it might not be feasible to determine when a loop terminates.
-Hence, any output constructed during the single loop iteration is highlighted to possibly be repeated  an arbitrary number of times.
-Second, recursive functions are also only executed once per call, and any subsequent call ot that function returns a symbolic value.
-This, again, is due to the difficulty to determine the exact recursion depth similar to the number of loop iterations.
-Finally, since the symbolic execution engine applies a static analysis, any form of user interaction providing input data as well as any form of input referred to from the deployment context of the web application, i.e.
-configuration files or databases, is substituted by a symbolic value.
-
-%Language-wise the symbolic execution semantics for the tools presented in Section \ref{sec:workflow} currently does not support object-oriented programming.
-PHP provides a large number of built-in functions to manipulate strings as well as arrays.
-The current symbolic execution engine only provides implementations for a small range of PHP standard library functions, mainly string operations.
+\subsection{Assumptions And Limitations}\label{sec:limitations}
+In addition to the limitations introduced by  symbolic execution, such as symbolic inputs and one-time loop execution, the symbolic execution described in \cite{Nguyen:2014:BCG:2635868.2635928} does not support advanced language features, including object-oriented programming.
+Moreover, the range of standard library functions supported are mostly string operations, yet PHP provides a large collection of functions for array operations.
+All three tools described in Section \ref{sec:existing_tools} though have only been applied to systems that are small and not maintained any more.
+Despite the functional benefits these tools provide to developers, their practicality for large and modern systems with respect to advanced language support has not been evaluated.
 
 \section{Scalable Output Approximation}%
-The main objective of this experience report is to investigate whether the
-tools described in Section \ref{sec:stateoftheart} are practical. In particular
-we want to find out whether symbolic execution as the underlying technique for
-static output approximation scales for PHP, and what may limitations restrict
-scalability. We may divide possible limitations into either matters of
-engineering or conceptual limitations where no precise execution is possible.
+The main objective of this experience report is to investigate whether we can have built practical tools that are on output approximations like the tools in Section \ref{sec:existing_tools}.
+In particular, we want to find out whether symbolic execution as the underlying technique for static output approximation is scalable, i.e., it computes an accurate approximation for large and modern PHP systems with reasonable time and space consumption.
 
-In the following we describe in detail our implementation procedure, practical
-problems we encountered and whether these are limiting scalability conceptually
-or just require en gineering effort to be tamed.
+Also, we are interested in the limitations of scalable symbolic execution for PHP web applications.
+Limitations may either be conceptional, where symbolic execution of PHP code, in contrast to concrete execution, is inaccurate: For example an include expression evaluated to a symbolic value simply provides to little information to include a script.
+Or, in spite of infeasible effort required, limitations could allow an accurate symbolic execution.
+For example, exhaustively exploring all feasible paths in an application may result in an infeasible number of program states to store; nevertheless, this so called state explosion problem may be addressed with additional engineering.
 
 \subsection{Towards Scalable Output Approximation?}
-We re-implement\footnote{\url{https://github.com/smba/oak}} the existing
-symbolic execution semantics of \cite{Nguyen:2014:BCG:2635868.2635928} with the
-goal to support modern large-scale dynamic web applications, or at least to
-better understand possible limitations. For reference, we choose to aim to
-support  \textsf{WordPress}
-\footnote{\url{https://github.com/WordPress/WordPress }}, a popular Content
-Management System of reasonable size that is still maintained. Moreover, our
-implementation uses parts of
-Quercus\footnote{\url{http://quercus.caucho.com/}}, a pure Java implementation
-of the PHP interpreter, which is tested against the \textsf{WordPress} test
-suite. We choose a test-driven approach for our implementation with regression
-tests and continuous integration. Our test cases are based on both custom
-examples for different language features as well as \textsf{SchoolMate}, a 
-small-scale system, and, at least in parts, \textsf{WordPress}, whose output we
-manually validate and then automatically re-check for each test run.
+As our goal is to have practical tools support, we approach this question by re-implementing the symbolic execution semantics described in \cite{Nguyen:2014:BCG:2635868.2635928} and extending it with support for additional language features.
+We choose to aim to support symbolic execution for a real-world example system of reasonable size that incorporates the missing language features described in Section \ref{sec:limitations} and is still developed and maintained: \textsf{WordPress} is a popular open-source Content Management System (CMS) with around 300k lines that provides a vivid plug-in environment and is maintained by a large community.
+For our implementation of a symbolic execution engine for PHP, \textsf{Oak}, we have chosen a test-driven approach with continuous integration using regression tests based on \textsf{WordPress} and \textsf{SchoolMate} (see Section \ref{sec:experiment_setup}).
+Oak incorporates the existing execution semantics of \cite{Nguyen:2014:BCG:2635868.2635928} and extends it with support for object-oriented orientation (class definition, class instantiation and method invocation) as well as PHP array operations.
+A list of the functions provided by the PHP standard library can be found at \cite{PHP}.
+Throughout the implementation we did not achieve high code coverage for WordPress, which was the main motivation to further investigate limitations in the wild.
 
-Our implementation basically follows the symbolic execution semantics of
-\cite{Nguyen:2014:BCG:2635868.2635928}, yet we extend it with additional rules
-for class definitions, class instantiation, and method invocation to support
-object-oriented programming. Furthermore, we add rules for do-while and 
-foreach loops since these are not supported as yet. Note that we execute all
-iterations of a foreach loop since the number of iterations is known since we
- are traversing all items of a given array.
+\subsection{What Went Wrong? Experience Report}
 
@@
 
-\subsection{Experiment Setup} 
+\subsection{Experiment Setup} \label{sec:experiment_setup}
 For the evaluation we symbolically executed a selection of PHP systems. We

### 2017-01-11 paper.tex

@@
-\documentclass[preprint]{sig-alternate-05-2015}
+\documentclass[sigconf]{acmart}
 
 \setlength\parindent{0pt}
-
 \usepackage[utf8]{inputenc}
@@
 \usepackage{hyperref}
+\usepackage{blindtext}
 \usepackage{booktabs}
@@
 \usepackage{lipsum}
-\usepackage{cite}
+%\usepackage{cite}
 \usepackage{enumerate}
@@
 
+\renewcommand{\sf}[1]{\textsf{#1}}
+\renewcommand{\tt}[1]{\texttt{#1}}
+
 \usepackage[zerostyle=d]{newtxtt}
@@
 
-\usepackage[labelfont=bf]{caption}
+%\usepackage[labelfont=bf]{caption}
+
+\definecolor{lightgrey}{HTML}{FFFFFF}
 
@@
 	language=PHP,
-	basicstyle=\ttfamily\tiny,
+	extendedchars = \true,
+	basicstyle=\ttfamily\footnotesize,
 	basewidth={0.5em,0.4em}, % for other columns settings
+	commentstyle = \color{gray},
 	columns=fullflexible,
@@
 	numberstyle=\tiny,
-	frame=single,
+	frame=tb,
 	showstringspaces=false,
@@
 	framexleftmargin=1.5em,
-	backgroundcolor=\color{lightgrey}
+	backgroundcolor=\color{lightgrey},
+	inputencoding = utf8x,
+    keepspaces = true,
+    keywordstyle = \bfseries,
 }
@@
 \newcommand{\dollar}{\mbox{\textdollar}}
+
+\newcommand*{\origrightarrow}{}
+\let\oldarrow\textrightarrow
+\renewcommand*{\textrightarrow}{\fontfamily{cmr}\selectfont\origrightarrow}
 
@@
 % TIMES
-\usepackage{mathptmx}% http://ctan.org/pkg/mathptmx
+%\usepackage{mathptmx}% http://ctan.org/pkg/mathptmx
 
@@
 \usepackage{xcolor}
-%\usepackage[dvipsnames]{xcolor}
-
-\definecolor{lightgrey}{HTML}{EEEEEE}
-
-\definecolor{orange}{RGB}{230, 100, 20}
-\definecolor{gruen}{RGB}{255, 190, 69}
-
-\definecolor{RYB1}{HTML}{DCDCDC}
-\definecolor{RYB2}{HTML}{C0C0C0}
-\definecolor{RYB3}{HTML}{808080}
-
-\definecolor{RYB4}{HTML}{DCDCDC}
-\definecolor{RYB5}{HTML}{C0C0C0}
-\definecolor{RYB6}{HTML}{808080}
-
-\definecolor{RYB7}{HTML}{DCDCDC}
-\definecolor{RYB8}{HTML}{C0C0C0}
-\definecolor{RYB9}{HTML}{808080}
-
-\definecolor{RYB10}{HTML}{D3D3D3}
-\definecolor{RYB11}{HTML}{A9A9A9}
-
-\pgfplotscreateplotcyclelist{ColorListBar}{
-{fill=RYB1},
-{fill=RYB2},
-{fill=RYB3},
-}
-
-\pgfplotscreateplotcyclelist{ColorListBar2}{
-{fill=RYB4},
-{fill=RYB5},
-{fill=RYB6},
-}
-
-\pgfplotscreateplotcyclelist{ColorListBar3}{
-{fill=RYB7},
-{fill=RYB8},
-{fill=RYB9},
-}
-
-\begin{document}
-\lstset{
-	language=PHP,
-	basicstyle=\footnotesize\ttfamily,
-    commentstyle = \color{gray},
-    extendedchars = \true,
-    inputencoding = utf8x,
-    keepspaces = true,
-    keywordstyle = \bfseries,
-    frame = tb
-}
 
 % Copyright
-\setcopyright{acmcopyright}
+%\setcopyright{none}
+%\setcopyright{acmcopyright}
 %\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
+\setcopyright{rightsretained}
 %\setcopyright{usgov}
@@
 % DOI
-\doi{10.475/123_4}
+\acmDOI{10.475/123_4}
 
 % ISBN
-\isbn{123-4567-24-567/08/06}
+\acmISBN{123-4567-24-567/08/06}
 
 %Conference
-\conferenceinfo{ISSTA '17}{July 9--13, 2013, Santa Barbara, CA, USA}
+\acmConference[ISSTA'17]{ACM Woodstock conference}{July 2017}{Santa Barbara,
+California, USA}
+\acmYear{1997}
+\copyrightyear{2016}
 
-\acmPrice{\$15.00}
+\acmPrice{15.00}
 
-%
-% --- Author Metadata here ---
-\conferenceinfo{ISSTA '17}{July 9--13, 2013, Santa Barbara, CA, USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data
-% (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-% --- End of Author Metadata ---
-
+\begin{document}
 \title{Does It Scale? Static Output Approximation of PHP Web Applications Using
 Symbolic Execution}
+\titlenote{Produces the permission block, and copyright information}
+%\subtitle{Extended Abstract}
+%\subtitlenote{The full version of the author's guide is available as  \texttt{acmart.pdf} document}
 
-%
+\author{Stefan Mühlbauer}
+\affiliation{%
+  \institution{Technische Universität Braunschweig}
+  %\streetaddress{P.O. Box 1212}
+  %\city{Braunschweig} 
+  %\state{Germany} 
+  %\postcode{43017-6221}
+}
+%\email{s.muehlbauer@tu-bs.de}
 
-\numberofauthors{3} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
-Stefan Mühlbauer\\
-       \affaddr{TU Braunschweig, Germany}\\
+\author{Christian Kästner}
+%\orcid{1234-5678-9012}
+\affiliation{%
+  \institution{Carnegie Mellon University}
+  %\streetaddress{5000 Forbes Avenue}
+  %\city{Pittsburgh} 
+  %\state{PA} 
+  %\postcode{15217}
+}
 %\email{s.muehlbauer@tu-bs.de}
-% 2nd. author
-\alignauthor
-Christian Kästner\\
-      \affaddr{Carnegie Mellon University, USA}\\
-       %\email{ckaestner@cs.cmu.edu}
-% 3rd. author
-\alignauthor
-Tien N. Nguyen\\
-       \affaddr{University of Texas, USA}\\
-%      %\email{tien.n.nguyen@utdallas.edu}
+
+\author{Tien N. Nguyen}
+%\orcid{1234-5678-9012}
+\affiliation{%
+  \institution{University of Texas at Dallas}
+  %\streetaddress{P.O. Box 1212}
+  %\city{Braunschweig} 
+  %\state{Germany} 
+  %\postcode{43017-6221}
 }
-% There's nothing stopping you putting the seventh, eighth, etc.
-% author on the opening page (as the 'third row') but we ask,
-% for aesthetic reasons that you place these 'additional authors'
-% in the \additional authors block, viz.
-\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
-email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
-(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
-\date{30 July 1999}
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
+%\email{s.muehlbauer@tu-bs.de}
 
-\maketitle
+% The default list of authors is too long for headers}
+\renewcommand{\shortauthors}{S. Mühlbauer et al.}
+
 \begin{abstract}
@@
 
-%
-%  Use this command to print the description
-%
-\printccsdesc
-
 % We no longer use \terms command
@@
 
-\keywords{Static program behavior, Static metrics, Dynamic language features,
-Static Analysis, PHP}
+\keywords{ACM proceedings, \LaTeX, text tagging}
+
+\maketitle
 
@@
 application frameworks for  script languages (Ruby On Rails, Django) to
-languages tailored specifically to the domain of web applications. PHP \cite{phpNET} is
-a programming language focused on server-side application development. As of 2012,
-it was used by 78.8 percent of the ten million most popular websites (according
+languages tailored specifically to the domain of web applications. PHP
+cite{phpNET} is a programming language focused on server-side application
+development. As of 2012, it was used by 78.8 percent of the ten million most popular websites (according
 to Alexa popularity ranking) \cite{alexaPHP}, ranking 7th on the TIOBE
-programming community index \cite{tiobePHP}, and was the ranked as the 6th most
-popular language on GitHub \cite{githubPHP}.
+programming community index \cite{tiobePHP}, and was the ranked as the 6th
+most popular language on GitHub \cite{githubPHP}.
 
-One common property of all technologies for dynamic web applications is \emph{staged computation} of output: A dynamic web application as a whole consists of both static code, such as scripts, and dynamically generated code, such as responses to HTTP requests.
-The latter code, although assembled at runtime, may contain client-side parts of the web application, such as JavaScript.
-So, to study dynamic web applications in its entirety, we need to consider both static as well as dynamic aspects of systems.
+One common property of all technologies for dynamic web applications is
+\emph{staged computation} of output: A dynamic web application as a whole
+consists of both static code, such as scripts, and dynamically generated code,
+such as responses to HTTP requests. The latter code, although assembled at
+runtime, may contain client-side parts of the web application, such as
+JavaScript. So, to study dynamic web applications in its entirety, we need to
+consider both static as well as dynamic aspects of systems.
 
@@
 parts of the web application can be analyzed.
-
-%Symbolic execution extends concrete execution semantics with semantics to
-%handle symbolic values. Hence, concrete execution semantics is contained in
-%symbolic execution semantics. Symbolic execution has been used and studied for generating test suites, but
-%has more recently also been applied for approximating the output of web
-%applications.
 
@@
 of the previous symbolic execution semantics
-\cite{Nguyen:2014:BCG:2635868.2635928} for PHP and additional features, such as
-object-oriented programming. We evaluated our
+\cite{Nguyen:2014:BCG:2635868.2635928} for PHP and additional features, such
+as object-oriented programming. We evaluated our
 symbolic execution engine for large-scale and modern PHP systems.
@@
 
-\begin{compactitem}
-	\item \emph{PhpSync}: Using all variants of  HTML client page output, every
+\begin{itemize}
+	\item {\emph{PhpSync}: Using all variants of  HTML client page output, every
 	single one can be statically checked for Markup Validity, i.e., if the web page
@@
 	for the defect, which then can be refactored manually
-	\cite{Nguyen:2011:AFH:2190078.2190142}.
+	\cite{Nguyen:2011:AFH:2190078.2190142}. }
 
@@
 	\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
-\end{compactitem}
+\end{itemize}
 
 \subsection{Assumptions And Limitations}\label{sec:limitations}
-In addition to the limitations introduced by  symbolic execution, such as symbolic inputs and one-time loop execution, the symbolic execution described in \cite{Nguyen:2014:BCG:2635868.2635928} does not support advanced language features, including object-oriented programming.
-Moreover, the range of standard library functions supported are mostly string operations, yet PHP provides a large collection of functions for array operations.
-All three tools described in Section \ref{sec:existing_tools} though have only been applied to systems that are small and not maintained any more.
-Despite the functional benefits these tools provide to developers, their practicality for large and modern systems with respect to advanced language support has not been evaluated.
+In addition to the limitations introduced by  symbolic execution, such as
+symbolic inputs and one-time loop execution, the symbolic execution described
+in \cite{Nguyen:2014:BCG:2635868.2635928} does not support advanced language
+features, including object-oriented programming. Moreover, the range of
+standard library functions supported are mostly string operations, yet PHP
+provides a large collection of functions for array operations.
+All three tools described in Section \ref{sec:existing_tools} though have only
+been applied to systems that are small and not maintained any more. Despite the
+functional benefits these tools provide to developers, their practicality for
+large and modern systems with respect to advanced language support has not been
+evaluated.
 
 \section{Scalable Output Approximation}%
-The main objective of this experience report is to investigate whether we can have built practical tools that are on output approximations like the tools in Section \ref{sec:existing_tools}.
-In particular, we want to find out whether symbolic execution as the underlying technique for static output approximation is scalable, i.e., it computes an accurate approximation for large and modern PHP systems with reasonable time and space consumption.
+The main objective of this experience report is to investigate whether we can
+have built practical tools that are on output approximations like the tools in
+Section \ref{sec:existing_tools}. In particular, we want to find out whether
+symbolic execution as the underlying technique for static output approximation
+is scalable, i.e., it computes an accurate approximation for large and modern
+PHP systems with reasonable time and space consumption.
 
-Also, we are interested in the limitations of scalable symbolic execution for PHP web applications.
-Limitations may either be conceptional, where symbolic execution of PHP code, in contrast to concrete execution, is inaccurate: For example an include expression evaluated to a symbolic value simply provides to little information to include a script.
-Or, in spite of infeasible effort required, limitations could allow an accurate symbolic execution.
-For example, exhaustively exploring all feasible paths in an application may result in an infeasible number of program states to store; nevertheless, this so called state explosion problem may be addressed with additional engineering.
+Also, we are interested in the limitations of scalable symbolic execution for
+PHP web applications. Limitations may either be conceptional, where symbolic
+execution of PHP code, in contrast to concrete execution, is inaccurate: For
+example an include expression evaluated to a symbolic value simply provides to
+little information to include a script. Or, in spite of infeasible effort
+required, limitations could allow an accurate symbolic execution. For example,
+exhaustively exploring all feasible paths in an application may result in an
+infeasible number of program states to store; nevertheless, this so called
+state explosion problem may be addressed with additional engineering.
 
 \subsection{Towards Scalable Output Approximation?}
-As our goal is to have practical tools support, we approach this question by re-implementing the symbolic execution semantics described in \cite{Nguyen:2014:BCG:2635868.2635928} and extending it with support for additional language features.
-We choose to aim to support symbolic execution for a real-world example system of reasonable size that incorporates the missing language features described in Section \ref{sec:limitations} and is still developed and maintained: \textsf{WordPress} is a popular open-source Content Management System (CMS) with around 300k lines that provides a vivid plug-in environment and is maintained by a large community.
-For our implementation of a symbolic execution engine for PHP, \textsf{Oak}, we have chosen a test-driven approach with continuous integration using regression tests based on \textsf{WordPress} and \textsf{SchoolMate} (see Section \ref{sec:experiment_setup}).
-Oak incorporates the existing execution semantics of \cite{Nguyen:2014:BCG:2635868.2635928} and extends it with support for object-oriented orientation (class definition, class instantiation and method invocation) as well as PHP array operations.
-A list of the functions provided by the PHP standard library can be found at \cite{PHP}.
-Throughout the implementation we did not achieve high code coverage for WordPress, which was the main motivation to further investigate limitations in the wild.
+As our goal is to have practical tools support, we approach this question by
+re-implementing the symbolic execution semantics described in
+\cite{Nguyen:2014:BCG:2635868.2635928} and extending it with support for
+additional language features. We choose to aim to support symbolic execution
+for a real-world example system of reasonable size that incorporates the
+missing language features described in Section \ref{sec:limitations} and is
+still developed and maintained: \textsf{WordPress} is a popular open-source
+Content Management System (CMS) with around 300k lines that provides a vivid
+plug-in environment and is maintained by a large community.
+For our implementation of a symbolic execution engine for PHP, \textsf{Oak}, we
+have chosen a test-driven approach with continuous integration using regression
+tests based on \textsf{WordPress} and \textsf{SchoolMate} (see Section
+\ref{sec:experiment_setup}). Oak incorporates the existing execution semantics
+of \cite{Nguyen:2014:BCG:2635868.2635928} and extends it with support for
+object-oriented orientation (class definition, class instantiation and method
+invocation) as well as PHP array operations. A list of the functions provided
+by the PHP standard library can be found at \cite{PHP}.
+Throughout the implementation we did not achieve high code coverage for
+WordPress, which was the main motivation to further investigate limitations in
+the wild.
 
-\subsection{What Went Wrong? Experience Report}
+\subsection{Things We Learned The Hard Way}
+\textcolor{blue}{\blindtext}
+
+\subsubsection{Symbolic vs Concrete Execution}
+\textcolor{blue}{\blindtext}
+
+\subsubsection{Multi-Target Method Calls}
+The first non-conceptual problem we encountered, in particular for
+\sf{WordPress} is related to object-oriented programming. While exploring
+different execution paths when the control flow is split, a variable can have
+different values depending on the path condition under which a value has been
+assigned to it (see the code example in Figure \ref{fig:output_example}). Similar, a variable
+can point to different objects depending on the path condition as the snippet in
+Figure \ref{fig:multi_target} illustrates. The variable \tt{\$person} can either
+point to object \tt{Alice} or \tt{Bob}. In addition, for any method invocated on
+that variable, the actual target is ambiguous as there are two, depending on the
+path condition. In order to have a sound symbolic execution, both method calls
+need to be executed. Nevertheless, the number of contexts for a method can grow
+rapidly, and easily become infeasible to execute.
+
+We implemented several heuristic approaches to tame this state space explosion
+for multi-target method calls, including executing a method only for a subset or
+just a single object. This trade-off between accuracy and soundness
+of the symbolic execution and performance time required.
+
+\subsubsection{Dynamic Language Features}
+Conceptually, symbolic execution failed for two types of language features. As
+stated in Section \ref{sec:workflow}, dynamic web applications are staged and client page
+output is computed at runtime. Symbolic execution hand has shown to be a useful
+technique to approximate dynamically generated output, it is challenged by
+language features that require concrete information rather than symbolic
+information at runtime. In particular, this applies to language features that
+provide access to reusable code of the web application: include expressions and
+function calls.
+
+First, include expressions in PHP usually require a string value representing
+the path to the script file or template to include. If this string value is
+statically provided by a string literal or can be constructed unambiguously, the
+symbolic execution engine was able to resolve the include expression and include
+the desired file. However, it is also common to assemble include expressions not
+only static string literals, but from several sources including user input
+information, database query results or configuration files. Figure
+\ref{fig:dynamic_includes} shows a small example of a dynamic include containing
+information retrieved from a database. Since any input for a symbolically
+executed application is symbolic this information leads to include expressions
+containing symbolic values. Also, configuration files represent only one
+particular configuration or even just default values or placeholders. Hence, 
+symbolic execuion is prone to fail for dynamic includes if symbolic information
+is contained.
+
+Second, function calls can either be direct by statically providing the name of
+the function (or method), or indirect. Indirect function calls in PHP are
+enabled by using built-in functions like \tt{call\_user\_func},
+\tt{call\_user\_func\_array} or simply passing a string value with a direct
+function call to the \tt{eval} function. The first two functions take as
+arguments the name of the function two call, and all arguments to that functions either as additional arguments or
+array respectively. The eval function evaluates a given PHP expression passed
+as a string literal.
+Again, we encountered similar issues to dynamic include expressions with
+indirect function call mechanisms. The example in Figure
+\ref{fig:indirect_calls} illustrates two functions \tt{header\_serif} and
+\tt{header\_sserif} which both print a HTML headline. Depending on whether the
+value of the variable \tt{\$style} retrieved from a database is \tt{serif} or
+\tt{sserif}, a different function is called. Given that the function name is 
+concrete at runtime, the desired function can be called. Otherwise, if the
+function name is assembled at runtime and contains symbolic values, no function
+can be called as the name is ambiguous or unknown. For this example, symbolic
+execution is not able to determine which function to call indirectly.
+
+% code examples for 3.3 experience report
+\begin{figure}[t]
+	\begin{subfigure}[b]{0.48\textwidth}
+
+\begin{lstlisting}
+<?php
+class Person {
+   function __construct__($name) {
+      $this->name = $name;
+   }
+   function greet($greeting) {
+      echo $greeting . $this->name;
+   }
+}
+
+if (...) {
+   $person = new Person("Alice");
+} else {
+   $person = new Person("Bob");
+}
+$person->greet("Good Morning, ");
+\end{lstlisting}
+	\caption{
+		Dynamic dispatch: Multiple targets for method calls
+		\label{fig:multi_target}
+	}
+	\end{subfigure}
+	\vspace{2mm}
+	\hfill
+	\begin{subfigure}[b]{0.48\textwidth}
+
+\begin{lstlisting}
+<?php
+define("ROOT", getcwd());
+define("TEMPLATES", ROOT . "templates/")
+
+$template = mysql_result(...);
+
+require_once TEMPLATE . $template;
+\end{lstlisting}
+	\caption{
+		Dynamic Include Resolution
+		\label{fig:dynamic_includes}
+	}
+	\end{subfigure}
+	\vspace{2mm}
+	\hfill
+	\begin{subfigure}[b]{0.48\textwidth}
+
+\begin{lstlisting}
+<?php
+function header_serif($title) {
+   $style = "font-family: serif;";
+   echo "<h1 style = '$style'>$title</h1>";
+}
+function header_sserif($title) {
+   $style = "font-family: sans-serif;";
+   echo "<h1 style = '$style'>$title</h1>";
+}
+\end{lstlisting}
+	\caption{
+		Function calls by indirect invocation mechanisms
+		\label{fig:indirect_calls}
+	}
+
+	\end{subfigure}
+	\caption{
+		Code Examples of defective code features
+		\label{fig:defective_code_features}
+	}
+\end{figure}
+
+\subsection{Things We Tried\ldots}
+\textcolor{blue}{\blindtext\blindtext}
 
@@
 %
-\begin{compactitem}
+\begin{itemize}
 	\item four small-scale systems that selected from the previously mentioned case
@@
 	Management Systems \cite{codegeekz}.
-\end{compactitem}
+\end{itemize}
 
@@
 \end{figure}
+
 \section{Lessons Learned}
+\textcolor{blue}{\Blindtext}
 
 \section{Related Work}
-\begin{itemize}
-	\item Hills et al. \cite{Hills:2013:ESP:2483760.2483786}
-	\item Output approximation using context-free grammar construction; applications \cite{minamide2005static,wassermann2007sound}
-\end{itemize}
+\textcolor{blue}{\blindtext\blindtext}
 
 %References
-\bibliographystyle{abbrv}
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{bibliography}

### 2017-01-11 paper.tex
added patrs of 3

@@
 
-\setlength\parindent{0pt}
+% just in case
+\usepackage{etex}
+
+% does not seem to work with sigconf template
+%\usepackage{cite}
+
 \usepackage[utf8]{inputenc}
@@
 \usepackage{lipsum}
-%\usepackage{cite}
+
 \usepackage{enumerate}
@@
 \usepackage{framed}
+
+\setlength\parindent{0pt}
 
@@
     keepspaces = true,
-    keywordstyle = \bfseries,
+    keywordstyle = \bfseries
 }
 
-\newcommand{\dollar}{\mbox{\textdollar}}
+%\newcommand{\dollar}{\mbox{\textdollar}}
 
@@
 conducted, which have been implemented in the following tools:
-
 \begin{itemize}
@@
 plug-in environment and is maintained by a large community.
-For our implementation of a symbolic execution engine for PHP, \textsf{Oak}, we
-have chosen a test-driven approach with continuous integration using regression
-tests based on \textsf{WordPress} and \textsf{SchoolMate} (see Section
-\ref{sec:experiment_setup}). Oak incorporates the existing execution semantics
+For our implementation of a symbolic execution engine for PHP,
+\textsf{Oak}\footnote{The implementation can be found
+at \url{github.com/smba/oak}.}, we have chosen a test-driven approach with
+continuous integration using regression tests based on \textsf{WordPress} and \textsf{SchoolMate} (see Section \ref{sec:experiment_setup}). Oak incorporates the existing execution semantics
 of \cite{Nguyen:2014:BCG:2635868.2635928} and extends it with support for
@@
 
-\subsection{Things We Learned The Hard Way}
+\subsection{Experience Report}
+This sections sums up 
 \textcolor{blue}{\blindtext}
@@
 \subsubsection{Symbolic vs Concrete Execution}
-\textcolor{blue}{\blindtext}
+Throughout the implementation, we basically followed the symbolic execution
+semantics described in \cite{Nguyen:2014:BCG:2635868.2635928}. Nevertheless, for additional language features as
+well as standard library functions we experienced it to be beneficial to the
+approximation to execute them as concretely as possible.
+For instance, \sf{WordPress} uses a text filter mechanism, where different
+filter functions are applied sequentially to a string literal. All filters can
+be linked to a hook using  associative arrays so that when the hook is
+triggered, all filters linked to it are applied. As described in Section \ref{}
+we execute loops, in particular while loops, only once due to an possibly
+indeterminable number of iterations. PHP yet offers more loop constructs: the
+do-while loop and the foreach loop. While for the do-while loop as well as for
+the while loop the number of iterations is possibly indeterminable, the foreach
+loop iterates over all items of a given array whose number of items is known at
+runtime. For \sf{WordPress}, we are able to achieve a more accurate
+approximation and higher code coverage results by concretely executing foreach
+loops as more server-side code (filter functions in our case) are accessed and
+executed. In turn, exhaustively executing foreach loops and filter functions
+concretely is a trade-off between accuracy and performance, as single hook can
+be called for different contexts as we are exploring multiple execution paths.
+
+An additional aspect on the boundary of concrete and symbolic execution, or
+effort and accuracy respectively, is the extent to which program inputs are
+assumed to be symbolic. As shown in Section \ref{} introducing symbolic values
+can be problematic if they are propagated to language features requiring
+concrete information, such as include expressions or indirect function calls.
+We manually scaled the number of sources of symbolic inputs by providing static
+dummy inputs such as a concrete form input, file name or configuration entry.
+For specific cases, where concrete information is inevitably required, we can
+achieve higher code coverage and a more accurate approximation, as more code
+becomes accessible. Nevertheless, while detecting dynamic features can be
+automated, it requires manual inspection and understanding of the web
+application to provide appropriate dummy input data. Again, this is a trade-off
+between the effort to provide input data and accuracy of the approximation and
+code coverage respectively.
 
@@
 of PHP systems is shown in Table \ref{corpus}. Our selection of PHP systems
-includes
-
-%
+includes%
 \begin{itemize}
@@
 \section{Related Work}
+\cite{Hills:2013:ESP:2483760.2483786}
+\cite{wassermann2007sound,minamide2005static}
 \textcolor{blue}{\blindtext\blindtext}
 
+\newpage
 %References

### 2017-01-11 paper.tex
completed sec 3

@@
 \subsection{Experience Report}
-This sections sums up 
-\textcolor{blue}{\blindtext}
+We identified a number of trade-offs between accuracy and performance that may
+cause the approximation to be impractical, yet possible. Also, we identified
+limitations impeding symbolic execution conceptually. For certain language
+features, concrete information rather than symbolic values may be required to
+actually be executed properly as this defective execution can impact subsequent
+parts of the program which may become inaccessible and may be missed.
 
@@
 }
+
+$font_style = mysql_result(...);
+call_user_func('header_' . $font_style, 'Title');
 \end{lstlisting}
@@
 
-\subsection{Things We Tried\ldots}
-\textcolor{blue}{\blindtext\blindtext}
+\subsubsection{Interpreter Customization}
+We have approached issues we encountered in different ways, including
+multiple-target method calls and foreach loops, as these represent a trade-off
+between accuracy of execution and performance.
+For method calls having ambiguous targets, we implemented different modes of
+execution, ranging from a complete mode where a method is executed for all
+possible targets to abstract modes with either a subset of targets or only a
+single target. In the latter cases where targets are discarded, we assumed
+symbolic values as method return values. Our symbolic interpreter can be
+customized to work in the desired mode to scale accuracy or performance.
+Nevertheless, for \sf{WordPress} these variations did not have significant
+impact on the output approximation as symbolic execution was rather impeded by
+dynamic features.
+Also, for foreach loops we are able to toggle between a single iteration mode
+and a complete iteration mode. Although we were able to increase code coverage
+by accessing more code, it did not significantly increase our output
+approximation.
+Note that these variants have only been tested for \sf{WordPress} to better
+understand possible trade-offs or simply achieve any output approximation in a
+reasonable period of time. We believe based on our experience dynamic features
+to be the main conceptual limitation impeding code coverage as well as output
+approximation quality. In the next Section we analyze in detail the impact of
+dynamic features with concrete foreach loop execution and complete
+multiple-target execution for method calls.
 

### 2017-01-13 paper.tex
refactored introduction

@@
 One common property of all technologies for dynamic web applications is
-\emph{staged computation} of output: A dynamic web application as a whole
+\emph{staged computation}: A dynamic web application as a whole
 consists of both static code, such as scripts, and dynamically generated code,
-such as responses to HTTP requests. The latter code, although assembled at
+such as client page output. The latter code, although assembled at
 runtime, may contain client-side parts of the web application, such as
@@
 
-%Since based on both static values and runtime values, output code is generated step by step
-%dynamically, output can be dependent on application inputs. Dynamic web
-%applications incorporate various both server-side inputs (databases,
-%configuration) and client-side inputs respectively. Static
-%analysis of dynamic web applications, hence, has to consider different input
-%sources.
-
 \emph{Symbolic execution} is a static analysis technique between program testing
-and program proving. For a program, symbolic, i.e., abstract, yet fixed inputs are
-applied instead of concrete inputs \cite{Darringer1978,King1976}. The underlying
-concept is to map program input (symbolically) to program output. Different classes of inputs may
-result in variational output, representing dependencies of output and input.
-As symbolic execution of a program returns an input-output-mapping for symbolic
-inputs, this analysis technique helps unfold the \emph{staged} nature of web
-applications: Since every feasible path is executed all possible output is
-contained in the symbolic output, and all dynamically generated variational
-parts of the web application can be analyzed.
-
-Knowledge about different output variants enabled useful analyses
-for the domain of web applications, such as detecting and locating HTML
-validation errors \cite{Nguyen:2011:AFH:2190078.2190142}. As for developers,
-from an output model, outlines and callgraphs for easier IDE code navigation
-\cite{Nguyen:2014:BCG:2635868.2635928}, or program slices
-\cite{Nguyen:2015:CPS:2786805.2786872} can be computed. Existing work has been
-integrated in the Varis plugin for the Eclipse IDE \cite{Nguyen:2015:VIS:2819009.2819140}.
+and program proving. For a program, symbolic values are
+used instead of concrete inputs \cite{Darringer1978,King1976}. The underlying
+concept is to map program input to program output: Symbolic values are used and
+propagated throughout the symbolic execution and keep dependencies for program
+output traceable. This analysis technique helps unfold the \emph{staged} nature
+of dynamic web applications. Since every feasible path can be executed the
+symbolic output contains both invariant output as well as output variants that
+depend on program input.
+
+Knowledge about different output output variants is leveraged by a number of
+analyses for the domain of web applications. Previous work presented tools
+for detecting and locating HTML validation errors
+\cite{Nguyen:2011:AFH:2190078.2190142}, computing program slices across
+server-side and client-page code \cite{Nguyen:2015:CPS:2786805.2786872}, or
+easing development and maintenance by extending IDE support for web applications
+with code navigation
+\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
 
@@
 model, static output approximation for PHP web applications using symbolic
-execution has so far only been evaluated for smaller systems that are not
-maintained anymore. The previously mentioned tools though are only practical, if
-for a given system the symbolic execution engine is scalable, i.e., will also
-approximate output accurately for larger and more recent systems with good time
-and space consumption.
-
-To investigate the question, whether we can build practical tools from the
-the symbolic execution engine, we re-implemented the engine with the specifications
-of the previous symbolic execution semantics
-\cite{Nguyen:2014:BCG:2635868.2635928} for PHP and additional features, such
-as object-oriented programming. We evaluated our
+execution has so far only been evaluated for small systems that are not
+maintained anymore. The tools presented by previous work
+\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2011:AFH:2190078.2190142}
+though are only practical, if for a given system the symbolic execution engine
+is scalable, i.e., will also approximate output accurately for larger and more
+recent systems with acceptable time and space consumption.
+
+To investigate the question, whether we can have a practical and scalable
+symbolic executoin engine, we re-implemented the engine with the specifications
+of the previous symbolic execution semantics \cite{Nguyen:2014:BCG:2635868.2635928} for PHP and additional features, such
+as object-oriented programming. We evaluate our
 symbolic execution engine for large-scale and modern PHP systems.
 
-During introduction of new semantics for object-orientation we encountered two
-main challenges. First, we were bound to introduce more concrete execution features to the
-symbolic execution engine, e.g. foreach loops for arrays, as this increased
-code coverage significantly. Second, we had to make a trade-off between the
-accuracy of method calls and the scalability of this language feature as a
-method call for different program state variants can have multiple targets.
-
-Based on our observations and evaluation results, we learned that for dynamic
-PHP constructs, expressions that are assembled dynamically, in combination with
-symbolic values symbolic execution can be impractical. For object-orientation
-and method calls, symbolic execution semantics can be a tightrope walk between
-state space explosion and imprecision. Moreover, the nature of PHP challenges a
-practical symbolic execution engine since for it to be accurate, a large number
-of functions from the standard library needs to be supported.
-
-Our key contributions in this paper include (1) a new tool infrastructure to
-statically approximate the output for PHP web applications using symbolic
-execution, and (2) a report of our observations with state-of-the-art web
-applications as well as an explanation on conceptual limitations for symbolic
-execution of PHP applications.
+During the introduction of new semantics, we addressed two  trade-offs between
+accuracy of our e xecution and performance: For method calls with ambiguous
+targets as well as concrete execution of loops the number of program states can
+become infeasible. This may require additional engineering adjustments to tame
+the state space explosion at cost of accuracy.
+
+Based on out observations we identify conceptual limitations of symbolic
+execution for PHP.  Dynamic features such as indirect function calls and
+include expression require concrete information to be evaluated properly. Since
+these expressions may be assembled at runtime and can contain symbolic
+information, for many cases this restricts our engine from further execution.
+Although some effort to statically approximate include expression can be spent,
+it is often non-trivial since expressions can contain information coming from
+various inputs including databases, user inputs or configuration files. We
+empirically evaluate our observations for a corpus including large PHP systems that are still maintained.
+
+Our key contribution in this paper include (1)  a new tool infrastructure to
+statically approximate the client-page output for PHP web applications using
+symbolic execution and (2) a report of out observations for state-of-the-art
+web applications as well as an empirical evaluation of conceptual limitations
+of symbolic executions for PHP web applications.
 

### 2017-01-13 paper.tex
refactored 2 w/o 2.2

@@
 \section{State Of the Art}\label{sec:stateoftheart}
-Symbolic execution is one of integral analysis techniques used by the tools
+Symbolic execution is an analysis techniques used by the tools
 whose practicality we aim to evaluate. This section recaps the idea of symbolic
@@
 Symbolic execution is a static analysis technique that was proposed by James C.
-King in 1976 \cite{King1976}. Symbolic execution allows to explore all feasible
+King in 1976 \cite{King1976}. Symbolic execution allows to explore feasible
 paths in a program. Based on normal program execution, the execution semantics
-is extended to handle symbolic values, which are abstract, but fixed; an
-execution becomes symbolic by introducing symbolic values as program inputs
-instead of concrete values \cite{King1976,Darringer1978}. These symbolic values
-are then propagated and used throughout the execution.
-
-Execution starts with a plain path condition, usually a tautology. Whenever the
-control-flow can be split into different branches, the corresponding path
-conditions (guards for instance) are conjoined with the previous path condition
-\cite{King1976}. Additionally, the conjoined path condition can be checked for
-satisfiability to exclude infeasible paths. Consequently, for a given program
-symbolic execution computes a mapping from (symbolic) input values to output values.
+is extended to handle symbolic values. An execution becomes symbolic by
+assuming symbolic values as program inputs rather than obtaining concrete
+values \cite{King1976,Darringer1978}. These symbolic values are then propagated
+and used throughout the execution.
+
+Execution starts with a tautology path
+condition. The control-flow can be split whenever the decision between branches
+is ambiguous. This is the case, when conditional expressions, for example for
+if-statements, evaluate to a symbolic value rather than an actual result.
+The corresponding path conditions (guards for instance) are conjoined with the
+previous path condition \cite{King1976}. Additionally, the conjoined path
+condition can be checked for satisfiability in order to exclude infeasible
+paths.
 
@@
 All tools proposed so far leverage a representation of all possible HTML client
-page output. Since any variant can be analyzed in particular, this enables tool
-support addressing the web application as a whole in spite of its staged
-nature, where parts of the web application (client-side output) are generated
-dynamically.
-
-Based on the approximated output representation subsequent analyses can be
-conducted, which have been implemented in the following tools:
+page output. Since any single client page output variant can be analyzed this
+enables tool support addressing the web application as a whole in spite of its
+staged nature, where parts of the web application are
+generated dynamically.
+Based on the approximated output representation, subsequent analyses can be
+conducted:
 \begin{itemize}
-	\item {\emph{PhpSync}: Using all variants of  HTML client page output, every
-	single one can be statically checked for Markup Validity, i.e., if the web page
+	\item \emph{PhpSync}: Using all variants of  HTML client page output, every
+	single one can be statically checked for markup validity, i.e., if the web page
 	conforms to syntactical specifications for HTML and other client-side
-	languages. After validation errors are detected, auto-fixes are provided.
-	Otherwise, the validation error can be traced back to source code responsible
-	for the defect, which then can be refactored manually
-	\cite{Nguyen:2011:AFH:2190078.2190142}. }
+	languages. After the tool detects validation errors, auto-fixes
+	can be provided. Otherwise, it traces validation errors back to
+	source code responsible for the defect, which then can be refactored manually
+	\cite{Nguyen:2011:AFH:2190078.2190142}.
 
@@
 \subsection{Assumptions And Limitations}\label{sec:limitations}
-In addition to the limitations introduced by  symbolic execution, such as
-symbolic inputs and one-time loop execution, the symbolic execution described
-in \cite{Nguyen:2014:BCG:2635868.2635928} does not support advanced language
-features, including object-oriented programming. Moreover, the range of
+In addition to the assumptions introduced by symbolic execution, such as
+symbolic inputs and one-time loop execution, the symbolic execution
+engine described in previous work \cite{Nguyen:2014:BCG:2635868.2635928} does
+not support advanced language features, including object-oriented programming. Moreover, the range of
 standard library functions supported are mostly string operations, yet PHP
 provides a large collection of functions for array operations.
-All three tools described in Section \ref{sec:existing_tools} though have only
-been applied to systems that are small and not maintained any more. Despite the
+All three analyses described in Section \ref{sec:existing_tools} though have
+only been applied to systems that are small and not maintained any more. Despite the
 functional benefits these tools provide to developers, their practicality for

### 2017-01-13 paper.tex
2.2 refactored

@@
 % The default list of authors is too long for headers}
-\renewcommand{\shortauthors}{S. Mühlbauer et al.}
+\renewcommand{\shortauthors}{\textcolor{blue}{put shortauthors here}}
 
@@
 \subsection{Static Output Approximation for PHP} \label{sec:workflow}
-Symbolic execution is an analysis technique not dedicated to a particular
+Symbolic execution is not dedicated to a particular
 programming language. For dynamic web applications, in our case for web
 applications written in PHP, it can be used to approximate all possible client
-page output. Before we present in detail existing tools that aim to ease
-development and maintenance of PHP web applications, we illustrate how output
-approximation using symbolic execution is used by starting with an example.
+page output variants. Before we present in detail existing tools that aim to
+ease development and maintenance of PHP web applications, we illustrate how output
+approximation using symbolic execution is pursued by starting with a small
+example.
 
@@
 Figure \ref{fig:code_example} shows two code listings. The snippet of server-side web
-application code in Figure \ref{fig:web_application} contains both HTML and PHP. First, the
-function \texttt{make\_titles} defined in lines 7 – 13 prints the string
-\texttt{\$arg} passed as an argument between zero and an arbitrary number of
-times. Second, the function is indirectly called using the built-in function
+application code in Figure \ref{fig:web_application} contains both HTML code
+(input form in lines 1 -- 4) as well as PHP code: First, the function
+\texttt{make\_titles} defined in lines 7 – 13 prints the string \texttt{\$arg} passed as an argument between zero and an arbitrary number of
+times. Second, the function is indirectly called using the built-in mechnism
 \texttt{call\_user\_func} in line 15 with the string \texttt{Headline} and
@@
 used: The variable \texttt{\$greeting} consists of three concatenated strings,
-\texttt{Hello} the value of input field \texttt{name}, and an exclamation mark.
-For this server-side code snippet there are two possible client page output
+\texttt{Hello}, the value of input field \texttt{name}, and an exclamation mark.
+For the greeting there are two possible client page output
 variants: If the input field \texttt{name} is empty at runtime, line 19 is
@@
 
-%C. Introduction to symbolic execution of PHP code
-The following symbolic execution semantics for PHP features  based on \cite{Nguyen:2014:BCG:2635868.2635928}
-illustrate the necessary modifications made to handle the extent of ambiguity
-that comes with handling symbolic values. In addition to normal output, the
+The following symbolic execution semantics for PHP features based on
+previous work \cite{Nguyen:2014:BCG:2635868.2635928} illustrate the necessary
+modifications and extensions made to handle the level of ambiguity that comes with handling symbolic values. In addition to normal output, the
 output approximation contains preprocessor directives to express
-repeatable output as well as output variants. All approximated client page
+repeatable parts as well as output variants. All approximated client page
 output corresponding to Figure \ref{fig:web_application} consisting of two different variants
 is illustrated in Figure \ref{fig:output_example}.
-First, loops in the previous execution semantics are only executed once instead
-of an arbitrary number of times. Since a loop condition can contain symbolic
-values, it might not be feasible to determine when a loop terminates. Hence,
-any output constructed during the single loop iteration is highlighted
-(\texttt{repeat \ldots  endrepeat}) to possibly be repeated an arbitrary number
-of times. The output in line 7 is marked to be repeated an arbitrary number of times, as the number
-of loop iterations cannot be determined statically. Second, a similar procedure
+First, for loops in the previous execution semantics
+\cite{Nguyen:2014:BCG:2635868.2635928} only one iteration is executed instead of
+an arbitrary number of iterations. Since a loop condition can contain symbolic
+values, it might not be feasible to determine whether, and if so, when a loop
+execution terminates. Hence, any output constructed during the single loop
+iteration is highlighted (\texttt{repeat \ldots endrepeat}) to possibly be repeated an arbitrary number of times.
+The output in line 7 is marked to be repeated an arbitrary number of times, as the number of loop iterations cannot be determined statically.
+Second, a similar procedure
 applies to recursive functions. Once called, a recursive function is  also only
@@
 determine the exact recursion depth similar to the number of loop iterations.
-Third, since the symbolic execution engine applies a static analysis, any form
-of user interaction providing input data as well as any form of input referred
+Third, any form of user interaction providing input data as well as any form of input referred
 to from the deployment context of the web application, i.e. configuration files
-or databases, is substituted by a symbolic value. The context variable
+or databases, is assumed to be symbolic. The context variable
 \texttt{\$\_POST} represents input submitted to the server and, hence, is
 symbolic as every program input is per definition substituted by symbolic values. Finally, the
-output in lines 11 and 13 respectively are marked as alternatives
-(\texttt{if \ldots else \ldots endif}), as symbolic execution of lines 18 – 22
+output in lines 11 and 13 respectively are marked as alternative variants
+(\texttt{if \ldots else \ldots endif}) as symbolic execution of lines 18 – 22
 in Figure \ref{fig:web_application} will explore both different branches.

### 2017-01-14 paper.tex
refactored sec 3

@@
 
-\definecolor{lightgrey}{HTML}{FFFFFF}
+\definecolor{lightgrey}{HTML}{FFFFFC}
 
@@
 The main objective of this experience report is to investigate whether we can
-have built practical tools that are on output approximations like the tools in
-Section \ref{sec:existing_tools}. In particular, we want to find out whether
-symbolic execution as the underlying technique for static output approximation
-is scalable, i.e., it computes an accurate approximation for large and modern
-PHP systems with reasonable time and space consumption.
+have a scalable symbolic execution engine as it is the underlying technique for
+static output approximation in previous work
+\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140}.
+Our vision is a practical symbolic execution, i.e., it computes an accurate approximation for large and modern PHP systems with reasonable time and space consumption.
 
@@
 \subsection{Towards Scalable Output Approximation?}
-As our goal is to have practical tools support, we approach this question by
-re-implementing the symbolic execution semantics described in
-\cite{Nguyen:2014:BCG:2635868.2635928} and extending it with support for
-additional language features. We choose to aim to support symbolic execution
+As our goal is to have a scalable symbolic execution engine, we approach this
+question by re-implementing the symbolic execution semantics described in
+previous work \cite{Nguyen:2014:BCG:2635868.2635928} and extending it with
+support for additional language features. We aim to support symbolic execution
 for a real-world example system of reasonable size that incorporates the
 missing language features described in Section \ref{sec:limitations} and is
-still developed and maintained: \textsf{WordPress} is a popular open-source
-Content Management System (CMS) with around 300k lines that provides a vivid
-plug-in environment and is maintained by a large community.
+still developed and maintained. \textsf{WordPress} is a popular open-source
+Content Management System (CMS) with around \textcolor{blue}{300.000} lines of
+code that provides a vivid plug-in environment and is maintained by a large
+community.
 For our implementation of a symbolic execution engine for PHP,
@@
 object-oriented orientation (class definition, class instantiation and method
-invocation) as well as PHP array operations. A list of the functions provided
-by the PHP standard library can be found at \cite{PHP}.
+invocation) as well as most PHP array operations. A list of the functions
+provided by the PHP standard library can be found at \cite{PHP}.
 Throughout the implementation we did not achieve high code coverage for
-WordPress, which was the main motivation to further investigate limitations in
-the wild.
+\sf{WordPress}. In the following we present our observations we made thorughout
+the implementation regarding issues impeding symbolic execution.
 
 \subsection{Experience Report}
-We identified a number of trade-offs between accuracy and performance that may
-cause the approximation to be impractical, yet possible. Also, we identified
-limitations impeding symbolic execution conceptually. For certain language
-features, concrete information rather than symbolic values may be required to
+We identified trade-offs between accuracy and performance that may
+cause the approximation to be impractical, yet possible. Additionally, we
+detected limitations impeding symbolic execution conceptually: For certain
+language features, concrete information rather than symbolic values may be required to
 actually be executed properly as this defective execution can impact subsequent
@@
 Throughout the implementation, we basically followed the symbolic execution
-semantics described in \cite{Nguyen:2014:BCG:2635868.2635928}. Nevertheless, for additional language features as
-well as standard library functions we experienced it to be beneficial to the
-approximation to execute them as concretely as possible.
+semantics described in previous work \cite{Nguyen:2014:BCG:2635868.2635928}.
+Nevertheless, for additional language features we experienced it to be beneficial to the
+approximation to execute them as precisely as possible.
 For instance, \sf{WordPress} uses a text filter mechanism, where different
-filter functions are applied sequentially to a string literal. All filters can
+filter functions are applied sequentially to a string literal. Filters can
 be linked to a hook using  associative arrays so that when the hook is
-triggered, all filters linked to it are applied. As described in Section \ref{}
-we execute loops, in particular while loops, only once due to an possibly
+triggered, all filters that are linked to it are applied. As described in
+Section \ref{} we execute loops only once due to an possibly
 indeterminable number of iterations. PHP yet offers more loop constructs: the
@@
 executed. In turn, exhaustively executing foreach loops and filter functions
-concretely is a trade-off between accuracy and performance, as single hook can
-be called for different contexts as we are exploring multiple execution paths.
-
-An additional aspect on the boundary of concrete and symbolic execution, or
-effort and accuracy respectively, is the extent to which program inputs are
+concretely is a trade-off between accuracy and performance, as a single hook
+call site can be called from various contexts since we are exploring multiple
+execution paths.
+
+An additional aspect on the trade-off between concrete and symbolic execution,
+or effort and accuracy respectively, is the extent to which program inputs are
 assumed to be symbolic. As shown in Section \ref{} introducing symbolic values
@@
 \subsubsection{Multi-Target Method Calls}
-The first non-conceptual problem we encountered, in particular for
-\sf{WordPress} is related to object-oriented programming. While exploring
+A non-conceptual problem we encountere is related to object-oriented programming. While exploring
 different execution paths when the control flow is split, a variable can have
 different values depending on the path condition under which a value has been
-assigned to it (see the code example in Figure \ref{fig:output_example}). Similar, a variable
-can point to different objects depending on the path condition as the snippet in
-Figure \ref{fig:multi_target} illustrates. The variable \tt{\$person} can either
+assigned to it (see the code example in Figure \ref{fig:output_example}).
+Similarly, a variable can point to different objects depending on the path
+condition as the snippet in Figure \ref{fig:multi_target} illustrates. The variable \tt{\$person} can either
 point to object \tt{Alice} or \tt{Bob}. In addition, for any method invocated on
-that variable, the actual target is ambiguous as there are two, depending on the
+that variable the actual target is ambiguous as there are two, depending on the
 path condition. In order to have a sound symbolic execution, both method calls
-need to be executed. Nevertheless, the number of contexts for a method can grow
-rapidly, and easily become infeasible to execute.
+need to be executed. Nevertheless, the number of contexts for which a method can
+be called can grow rapidly, and easily become infeasible to execute
+exhaustively.
 
 We implemented several heuristic approaches to tame this state space explosion
-for multi-target method calls, including executing a method only for a subset or
-just a single object. This trade-off between accuracy and soundness
-of the symbolic execution and performance time required.
+for multi-target method calls, including executing a method only for a subset of
+target objects, or just a single object. This is a trade-off between accuracy
+and soundness of the symbolic execution and performance time required.
 
 \subsubsection{Dynamic Language Features}
-Conceptually, symbolic execution failed for two types of language features. As
-stated in Section \ref{sec:workflow}, dynamic web applications are staged and client page
-output is computed at runtime. Symbolic execution hand has shown to be a useful
-technique to approximate dynamically generated output, it is challenged by
-language features that require concrete information rather than symbolic
-information at runtime. In particular, this applies to language features that
-provide access to reusable code of the web application: include expressions and
-function calls.
+Conceptually, we identified two language features for which symbolic execution
+failed frequently. As stated in Section \ref{sec:workflow}, dynamic web
+applications are staged and client page output is computed at runtime. Symbolic
+execution has shown to be a useful technique to approximate dynamically
+generated output in theory, though it is challenged by language features that
+require concrete information rather than symbolic information at runtime. In particular, this
+applies to language features that provide access to reusable code of the web
+application: include expressions and function calls.
 
@@
 the desired file. However, it is also common to assemble include expressions not
-only static string literals, but from several sources including user input
+only from static string literals, but from several sources including user input
 information, database query results or configuration files. Figure
@@
 information retrieved from a database. Since any input for a symbolically
-executed application is symbolic this information leads to include expressions
-containing symbolic values. Also, configuration files represent only one
-particular configuration or even just default values or placeholders. Hence, 
-symbolic execuion is prone to fail for dynamic includes if symbolic information
-is contained.
-
-Second, function calls can either be direct by statically providing the name of
-the function (or method), or indirect. Indirect function calls in PHP are
+executed application is symbolic this information can be propagated to include
+expressions. Also, configuration files represent only one particular
+configuration or even just default values or placeholders. Hence, symbolic
+execuion is prone to fail for dynamic includes if symbolic information is contained.
+
+Second, function can either be called directly by statically providing the name
+of the function (or method), or indirectly. Indirect function calls in PHP are
 enabled by using built-in functions like \tt{call\_user\_func},
@@
 function call to the \tt{eval} function. The first two functions take as
-arguments the name of the function two call, and all arguments to that functions either as additional arguments or
-array respectively. The eval function evaluates a given PHP expression passed
-as a string literal.
-Again, we encountered similar issues to dynamic include expressions with
-indirect function call mechanisms. The example in Figure
+arguments the name of the function two call, and all arguments to that
+functions either as additional arguments or a single associative array
+respectively.
+The \tt{eval} function evaluates a given PHP expression passed as a string
+literal.
+Again, we encountered issues similar to dynamic include expressions with
+indirect function call mechanisms: The example in Figure
 \ref{fig:indirect_calls} illustrates two functions \tt{header\_serif} and
@@
 \subsubsection{Interpreter Customization}
-We have approached issues we encountered in different ways, including
-multiple-target method calls and foreach loops, as these represent a trade-off
-between accuracy of execution and performance.
+We have approached issues we encountered, including
+multiple-target method calls and foreach loops, in different ways as these
+represent a trade-off between accuracy of execution and performance.
 For method calls having ambiguous targets, we implemented different modes of

### 2017-01-14 paper.tex
4.0 4.1 and 4.2 refactored

@@
 
-\definecolor{lightgrey}{HTML}{FFFFFC}
+\definecolor{lightgrey}{HTML}{CCFFEE}
 
@@
 
-\subsection{Experience Report}
+\subsection{Experience Report}\label{sec:experience_report}
 We identified trade-offs between accuracy and performance that may
@@
 \section{Evaluation}
-As already stated in the previous section, we encountered sevaral problems
-during the development of the symbolic interpreter. Therefore this section
-describes our methods to evaluate practicality of the output approximation, and
-possible explanations.
-
-\subsection{Measuring Approximation Success} \label{heuristic}
-For the evaluation of our output approximation we require a ground truth output
-model to compare our approximation against. Since output literals are embedded
-in server-side code and scattered, it is nontrivial to determine whether a
-string literal is part of the ouput or not. So, in the absence of ground truth
-for the approximated output we are bound to choose a \emph{heuristic approach}
-in order to measure the accuracy of our aproximation. We define a string
-literal to be an \emph{output candidate} if there exists an execution path
-reaching the string literal and evantually passing it to an output-generating
-statement (e.g., \texttt{echo} or \texttt{print}).
-
-We approximate output candidates as those string literals containing the
-characters \texttt{<}, \texttt{>}, or both since we expect output to contain
-HTML tags. We evaluated this heuristic manually with a sample of 400 string
-literals randomly selected from the entire corpus.
-We measured for our heuristic classifier a precision of 94 percent, and a
-recall of 50 percent. This means that six percent of the string literals are
-classified incorrectly as false positives. In turn, the classifier is highly
-distinctive as 96 percent of the string literals are classified correctly. The
-recall of 50 percent means that half of the string literals, which we manually 
-classified as output, actually were responsive to the classifier. We attempted
-to increase recall by looking for further distinctive properties to build a
-classifier from, but could not do so without decreasing precision. In spite of
-missing half of the output candidates, we decided to use a simple, yet
-distinctive classifier.
+As already stated in the previous Section \ref{sec:experience_report}, we
+encountered several problems during the implementation of our symbolic
+execution engine for \sf{WordPress}. The main items of critique of the symbolic
+execution engine and tools described in previous work
+\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140}
+are the missing support for widely-used language features, and the evaluation
+only using small-scale systems.
+Although evaluating a tool for a small systems can be a valuable proof-of-concept, it does not necessarily cover the question of whether a tool is practical on a larger-scale.
+Therefore in the following we present our
+methods to evaluate practicality of our symbolic output approximation.
+Moreover, we investigate to which extent the conceptual limitations encountered
+with \sf{WordPress} apply for a wider choice of PHP systems.
 
 \subsection{Experiment Setup} \label{sec:experiment_setup}
-For the evaluation we symbolically executed a selection of PHP systems. We
-selected a corpus of twelve PHP systems with regard to system size and recency
-as the corpus of the case study for \cite{Nguyen:2014:BCG:2635868.2635928} only
-contained small-scale systems that are not maintained any more. The full list
-of PHP systems is shown in Table \ref{corpus}. Our selection of PHP systems
-includes%
-\begin{itemize}
-	\item four small-scale systems that selected from the previously mentioned case
-	study in order to compare our results with the previous symbolic execution
-	engine \cite{Nguyen:2014:BCG:2635868.2635928},
-
-	\item three recent large-scale systems selected from the case study corpus of
-	\cite{Hills:2013:ESP:2483760.2483786}, and
-
-	\item four small-scale systems selected from a list of recent Content
-	Management Systems \cite{codegeekz}.
-\end{itemize}
-
-As for for the last two cases, we limited our selection since the parser we
-used did not support all language features used.
-For the experimental evaluation we symbolically executed per system each script file or entry point (files with a \texttt{.php}, \texttt{.inc}, \texttt{.bit} or \texttt{.module} extension).
-All following measurements are cumulated per system for each entry point.
+We have encountered several issues for \sf{WordPress} that conceptually limit
+the scalability of symbolic execution. Since our observations so far only considers
+one example system an, we evaluate our symbolic execution engine for a wider
+range of PHP systems with regard to the following questions. The metrics to
+approach those questions are explained in Sections
+\ref{HowAccurateIsOurApproximation} and \ref{sec:understanding_limitations}.
+
+First, the evaluation of the output approximation using the symbolic execution
+engine, \sf{Symex}, in previous work \cite{Nguyen:2014:BCG:2635868.2635928} has
+shown high code coverage for small systems. As our symbolic execution
+re-implements \sf{Symex} and extends with support for additional language
+features, we ask whether we can replicate high coverage results at least on a small-scale. Therefore, our PHP
+corpus contains four small-scale systems used in previous work \cite{Nguyen:2014:BCG:2635868.2635928}.
+
+Second, we ask whether system size and usage of modern language features are
+factors affecting the practicality of our symbolic execution engine. We
+selected both a range of large-scale and small-scale modern PHP systems to
+better separate and understand the circumstances under which conceptual
+limitations might affect practicality. For the selection of large-scale systems
+we borrow three systems from a case study addressing the feature usage in PHP
+systems. Some more detailed description of this case study can be found in
+Section \ref{sec:related_work}. For the small-scale systems we selected five systems from
+a list of recent Content Management Systems (CMS) written in PHP \cite{codegeekz}. For both
+selections we were bound to select a smaller number of system than we intended
+to since the parser used in our implementation, \sf{Quercus}\footnote{Quercus
+footnote}, did not support a number of language features. The full list of PHP
+systems for that we evaluate our symbolic execution engine along with
+descriptive code metrics can be found in Table \ref{corpus}.
 
@@
 	\textbf{System} & \textbf{Version} & \textbf{Classification} & \textbf{SLOC} &
-	\textbf{\#files} & \textbf{\#OCs} & \textbf{\#includes}
-	\\
+	\textbf{\#files} & \textbf{\#OCs} & \textbf{\#includes}\\
 	\midrule
@@
 
+\subsection{Measuring Approximation Success} \label{heuristic}
+In order to accurately answer the question how good our symbolic execution
+engine approximates client page output, we would require knowledge of all
+possible variants to compare our approximation against. Since we do not have
+any ground truth information alike, we approach this demand by using a
+heuristic. Rather than having knowledge about all client page output variants
+we say our approximation is sound if all string literals (which are embedded in
+and  scattered across server-side code) that may eventually become part of any
+client page output variant (in the following referred to as \emph{output
+candidate}) is contained in our symbolic output approximation.
+It is non-trivial to determine whether a string literal is an output candidate,
+so we heuristically classify string literals as output candidates if they
+contain the characters \tt{<}, \tt{>}, or both since we expect output to contain
+HTML tags.
+
+We evaluated this heuristic manually with a sample of 400 string
+literals randomly selected from our corpus of PHP systems (see Table
+\ref{corpus}). We measured for our heuristic classifier a precision of 94 percent, and a
+recall of 50 percent. This means that six percent of the string literals are
+classified incorrectly as false positives. In turn, the classifier is highly
+distinctive as 96 percent of the string literals are classified correctly. The
+recall of 50 percent means that half of the string literals, which we manually 
+classified as output candidates, actually were responsive to the classifier. We
+attempted to increase recall by looking for further distinctive properties to build a
+classifier from, but could not do so without decreasing precision. 
+\textcolor{blue}{In spite of missing half of the output candidates, we decided to use a simple, yet
+precise and distinctive classifier.}
+
 \subsection{Measuring Approximation Accuracy}
@@
 
-\subsection{Understanding Limitations}
+\subsection{Understanding Limitations} \label{sec:understanding_limitations}
 As we have seen in Section \ref{HowAccurateIsOurApproximation} both measured
@@
 
-\section{Related Work}
+\section{Related Work} \label{sec:related_work}
 \cite{Hills:2013:ESP:2483760.2483786}

### 2017-01-14 paper.tex
monochromatic diagrams

@@
 %\usepackage{cite}
+
+\usepackage{tikz}
+\usetikzlibrary{patterns}
 
@@
 
-\definecolor{lightgrey}{HTML}{CCFFEE}
+\definecolor{lightgrey}{HTML}{FFFFFF}
 
@@
 are the missing support for widely-used language features, and the evaluation
-only using small-scale systems.
-Although evaluating a tool for a small systems can be a valuable proof-of-concept, it does not necessarily cover the question of whether a tool is practical on a larger-scale.
-Therefore in the following we present our
-methods to evaluate practicality of our symbolic output approximation.
-Moreover, we investigate to which extent the conceptual limitations encountered
+only using small-scale systems. Although evaluating a tool for a small systems
+can be a valuable proof-of-concept, it does not necessarily cover the question
+of whether a tool is practical on a larger-scale. Therefore in the following we
+present our methods to evaluate practicality of our symbolic output
+approximation. Moreover, we investigate to which extent the conceptual limitations encountered
 with \sf{WordPress} apply for a wider choice of PHP systems.
@@
 heuristic. Rather than having knowledge about all client page output variants
-we say our approximation is sound if all string literals (which are embedded in
-and  scattered across server-side code) that may eventually become part of any
+we say our approximation is accurate if all string literals (which are embedded
+in and  scattered across server-side code) that may eventually become part of any
 client page output variant (in the following referred to as \emph{output
@@
 In Section \ref{heuristic} we introduced the definition of output candidates as
-expected output. As a most accurate approximation contains all output
-candidates, we define two metrics to measure accuracy of our approximation.
-
-First, we measure how much of the expected output was actually processed by the
+expected output. For an approximation to be most accurate, it needs to contain
+all output candidates of a system analyzed. Based on whether output
+candidates are reached by an execution path and part of the output
+approximatiopn define two metrics to approach measuring accuracy.
+
+We measure how much of the expected output was actually processed by the
 symbolic interpreter.  Once a line of code, a statement or expression
 containing an output candidate is actually an element in an execution path, we
-define this output candidate as reached. We define the metric \emph{reach
+define this output candidate as \emph{reached}. We define the metric \emph{reach
 coverage} as the ratio of output candidates that are reached and the total
@@
 candidate is processed by the symbolic interpreter, it does not guarantee we
-will see that output candidate in the symbolic output. Second, we define the
+will see that output candidate in the symbolic output. Therefore, we define the
 metric \emph{output coverage} as the ratio of output candidates that are
 contained in the output of the symbolic interpreter and the total number of
-output candidates in the analyzed system.  For an imprecise approximation, we
-may see loss of output candidate information resulting in a output coverage
-lower than the reach coverage.
-
-The coverage results are illustrated in Figure \ref{coverage}. We could
-replicate high coverage results for the four small-scale systems with a reach
+output candidates in the analyzed system.  
+%For an imprecise approximation, we may see loss of output candidate information
+% resulting in a output coverage lower than the reach coverage.
+
+The coverage results for our case study are illustrated in Figure
+\ref{coverage}. We could replicate high coverage results for the four small-scale systems with a reach
 coverage and output coverage over 80 percent respectively. For the three

### 2017-01-14 paper.tex
4.4.1 refactored

@@
 whether their surrounding program code was not accessed/accessible, and if so,
-why. Of special interest are those output candidates that are located in HTML
+why. We are interested in those output candidates that are located in HTML
 files (possibly with nested PHP), or located in a function definition. Plain
-HTML files do not require any further execution (of course, unless PHP scripts
-are embedded) and represent output candidates that just need to be included
-properly to be part of the output. Aside, output candidates that are part of a
-function definition are only missed if the corresponding function is never
-called at runtime.
+HTML files do not require any further execution and represent output candidates
+that just need to be included properly to be part of the output. Aside, output
+candidates that are part of a function definition are only missed if the
+corresponding function is never called at runtime.
 
@@
 for almost all systems (except for \textsf{AddressBook} and \textsf{TimeClock})
-not-reached output candidates had a function context. Note that
-\textsf{SchoolMate} was excluded from the diagram since our analysis reached
-all output candidates for that system.
+not-reached output candidates had a function context.
+
+%Note that \textsf{SchoolMate} was excluded from the diagram since our analysis
+% reached all output candidates for that system.
 
@@
 
+\section{Conclusion}
+
 \newpage

### 2017-01-14 paper.tex
refactored 4 completely

@@
 \subsubsection{Dynamic Language Features}
+\label{sec:experiment_dynamicfeaturess} 
 Conceptually, we identified two language features for which symbolic execution
@@
 
+% RESULT DIAGRAMS
+\begin{figure}[h!]
+	\begin{subfigure}[center]{0.48\textwidth}
+		\input{new_diagrams/not_reached}
+		\caption{\label{coverage}}
+	\end{subfigure}
+
+	\begin{subfigure}[center]{0.48\textwidth}
+		\input{new_diagrams/includes}
+		\caption{
+			Include coverage and accuracy.
+			\label{fig:include_coverage_results}
+		}
+	\end{subfigure}
+
+	% Callback candidate explanation
+	\begin{subfigure}[center]{0.48\textwidth}
+		\input{new_diagrams/callbacks}
+		\caption{
+			Distribution dead function output candidates that are par-
+			tially or completely explained due to callback candidates.
+			\label{fig:output_candidate_explanation}
+		}
+
+	\end{subfigure}
+	\caption{Results}
+\end{figure}
+
 \subsubsection{What literals did we miss?}\label{WhatLiteralsDidWeMiss?}
@@
 context. As the classification statistics in Figure \ref{coverage} illustrate,
-for almost all systems (except for \textsf{AddressBook} and \textsf{TimeClock})
-not-reached output candidates had a function context.
+for almost all systems tested (except for \textsf{AddressBook} and
+\textsf{TimeClock}) not-reached output candidates had a function context.
 
@@
 
-\subsubsection{Inaccessible Dynamic Features}
-\label{sec:inaccessible}
+%\subsubsection{Inaccessible Dynamic Features} \label{sec:inaccessible}
 For both cases of context, either inclusion of a script file failed, or a
@@
 undefined at runtime if the corresponding script file is never included. In
-turn, if the function is defined at runtime the function can either be dead
-code if there is no call site for this function, or the function call failed.
-In PHP there are several ways to call a function beside direct call sites. The
-language offers indirect call mechanisms like callback-commands or an
-evaluation function that parses and evaluates PHP source code as strings.
+turn, if the function is defined at runtime, the function is dead
+code if there is no call site for this function, or the function is only called
+indirectly and the function name could not be resolved accurately. In PHP there
+are several ways to call a function beside direct call sites. 
+%The language offers indirect call mechanisms like callback-commands or an
+% evaluation function that parses and evaluates PHP source code as strings.
 
@@
 representing include targets can be included, but any include target containing
-symbolic values is ambiguous.
+symbolic values is ambiguous. Since it is unlikely that all functions containing
+missed output candidates are dead code having no call site (direct or
+indirect) at all, in Section \ref{WhyDidFunctionCallsFail} we further
+investigate how many functions fail due to inaccurately resolved indirect
+function calls. Moreover, we evaluate the accuracy of evaluation of include
+expressions in the next section.
 
@@
 \label{WhyDidIncludesFail}
-To better understand include expression evaluation, we use another metric to
-approach accuracy from a different angle: We measure (1) the ratio of reached
-include expressions and the total number of include expressions in a system
-respectively (coverage of include expressions), and (2) the ratio of
-successfully resolved include expressions and reached include expressions in a
-system (resolution success rate). For a most accurate result all include
-expressions are reached and resolved.
-
-For include expressions we sampled failed attempts and manually evaluated the
-cause of failure. We identified for (1) that all failed attempts were dynamic
-include expressions, i.e., those which were resolved at runtime and were
-assembled from fragments of environment-dependent information including
-database retrievals, user input, or files. Consequently, the resolved include
-expressions contained symbolic values which did not represent identifiable
-script files.
-
-Finally, we measured for include expressions the overall reach coverage and the
-overall resolution success rate per system. As illustrated in Figure
-\ref{fig:include_coverage_results}, except for the small-scale systems and two
-recent small-scale systems we did not even really reach a great portion of
-include expressions, although for most systems the resolution success rate was
-over 80 percent. Nevertheless, one failed attempt to include a script file can
-result in a cascade of missing more include expressions as the include target
-script is not executed.
+From Section \ref{sec:experiment_dynamicfeaturess} we learn that include
+expressions can not be evaluated accurately for symbolic values, and from the
+previous Section \ref{WhatLiteralsDidWeMiss?} that very little of missed output
+candidates can be explained by simply missing to include a file. To better
+understand the impact of include expression resolution throughout analyzing
+systems, we construct two additional metrics. First, we are interested in how
+many include expressions throughout the execution we actually reach in a system.
+Therefore we introduce the metric \emph{reach coverage for include expressions}
+as the ratio of reached include expressions and the total number of reach
+expressions in a system to check whether failed includes may have caused
+subsequent include expressions and so on. Second, we measure how successful
+include expressions are resolved.
+
+Figure \ref{fig:include_coverage_results} illustrates the classification of
+include expressions for each system: We can either reach an include expression
+and successfully resolve it, or not. Or, we did not reach an include expression at all. 
+Figure \ref{fig:include_coverage_results} indicates that except for the
+small-scale systems as well as \sf{Anchor} and \sf{Nibbleblog} we did not really
+reach a great portion of include expressions, although for most systems the
+resolution success rate was over 80 percent. Note that one failed attempt
+to include a script file can result in a cascade of missing more include
+expressions as the include target script file is not included and executed.
 
@@
 detecting the functions’ access points: The nodes at the top depict callback
-candidates, the nodes at the bottom with labels X, Y, and Z depict function
-definitions containing not-reached output candidates. In Figure
-\ref{fig:callgraph} function X can only be accessed through callback
-candidates, as for Figure \ref{fig:callgraph} function Y can be partially
-accessed by callback candidates, whereas function Z is never accessible through
-callback candidates.
+candidates, the nodes at the bottom with labels \tt{X}, \tt{Y}, and \tt{Z}
+depict function definitions containing not-reached output candidates. In Figure
+\ref{fig:callgraph} function \tt{X} can only be accessed through callback
+candidates, as for Figure \ref{fig:callgraph} function \tt{Y} can be partially
+accessed by callback candidates, whereas function \tt{Z} is never accessible
+through callback candidates.
 
@@
 	\includegraphics[scale=0.55]{images-paper/callgraph_cases}
-    \caption{Callgraph analysis: The bottom nodes represent functions containing missed output candidates, the top nodes depict entry points to them.}
-    \label{fig:callgraph} output candidates we measured how many of those functions were only, or partially accessibly through indirect calling mechanisms (see Section \ref{sec:inaccessible}).
-As Figure \ref{fig:output_candidate_explanation} illustrates, around 80 percent (except for \textsf{TimeClock}) all missed output candidates of each system were only accessible through callback candidates, i.e., functions that are only accessible through indirect calling mechanisms.
-
+    \caption{Callgraph analysis: The bottom nodes represent functions
+    containing missed output candidates, the top nodes depict entry points to
+    them.}
+    \label{fig:callgraph} 
 \end{figure}
 
-Like for include expressions, we again sampled dead function cases and
-identified the lack of direct call sites to those functions. As described in
-Section \ref{sec:inaccessible}, functions despite having no direct call sites
-were called through indirect call features. We identified dynamically assembled
-expressions (function names) to be responsible for failed attempts of calling a
-function indirectly as the resolved function name eventually contained symbolic
-values. These expressions, as well as include expressions, contained
-information dependent on the system environment.
-
-For failed function calls to functions containing not-reached
-
-	% RESULT DIAGRAMS
-\begin{figure}[h!]
-	\begin{subfigure}[center]{0.48\textwidth}
-		\input{new_diagrams/not_reached}
-		\caption{\label{coverage}}
-	\end{subfigure}
-
-	\begin{subfigure}[center]{0.48\textwidth}
-		\input{new_diagrams/includes}
-		\caption{
-			Include coverage and accuracy.
-			\label{fig:include_coverage_results}
-		}
-	\end{subfigure}
-
-	% Callback candidate explanation
-	\begin{subfigure}[center]{0.48\textwidth}
-		\input{new_diagrams/callbacks}
-		\caption{
-			Distribution dead function output candidates that are par-
-			tially or completely explained due to callback candidates.
-			\label{fig:output_candidate_explanation}
-		}
-
-	\end{subfigure}
-	\caption{Results}
-\end{figure}
+%Like for include expressions, we again sampled dead function cases and
+% identified the lack of direct call sites to those functions. As described in
+% Section \ref{sec:inaccessible}, functions despite having no direct call sites
+% were called through indirect call features. We identified dynamically
+% assembled expressions (function names) to be responsible for failed attempts
+% of calling a function indirectly as the resolved function name eventually
+% contained symbolic values. These expressions, as well as include expressions,
+% contained information dependent on the system environment.
+
+For failed function calls to functions containing not-reached output
+candidates we measured how many of those functions were only, or
+partially accessibly through indirect calling mechanisms (see Section
+\ref{sec:experiment_dynamicfeaturess}). As Figure
+\ref{fig:output_candidate_explanation} illustrates, around 80 percent (except
+for \sf{TimeClock}) all missed output candidates of each system were only
+accessible through callback candidates, i.e., functions that are only
+accessible through indirect calling mechanisms.
 
 \section{Lessons Learned}
-\textcolor{blue}{\Blindtext}
+%\textcolor{blue}{\Blindtext}
 
@@
 \cite{wassermann2007sound,minamide2005static}
-\textcolor{blue}{\blindtext\blindtext}
+%\textcolor{blue}{\blindtext\blindtext}
 
@@
 
-\newpage
+%\newpage
 %References

### 2017-01-14 new_diagrams/callback_classification.tex
tikz is fun

% X, Y, Z = target nodes
% XS, YS, ZS = source nodes
% M = 'main' context node

\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,
                    semithick,thick,scale=0.7, every node/.style={transform
                    shape}] \tikzstyle{every state}=[fill=white,draw=black,text=black]

  \node[state] 	(XS)                    {};
  \node[state]  (YS) [right of = XS]    {};
  \node[state]  (M)  [right of = YS]    {\tt{main}};
  \node[state]  (ZS) [right of = M]     {};
  \node[state]  (X)  [below of = XS] 	{\tt{X()}};
  \node[state]  (Y)  [below of = YS]    {\tt{Y()}};
  \node[state]  (Z)  [below of = ZS]    {\tt{Z()}};

  \path (XS) edge node {} (X)
  		(YS) edge node {} (Y)
  		(M)  edge [bend right] node {} (Y)
  		(M)  edge node {} (ZS)
  		(ZS) edge node {} (Z);
\end{tikzpicture}

### 2017-01-16 paper.tex
tikz is fun
Section 5 draft

@@
 \usepackage{pgfplots, pgfplotstable}
+
+\usepackage{pgf}
+\usepackage{tikz}
+\usetikzlibrary{arrows,automata}
 
@@
 
-\section{Scalable Output Approximation}%
+\section{Scalable Output Approximation} \label{sec:3}
 The main objective of this experience report is to investigate whether we can
@@
 % Callback candidate description
-\begin{figure}[t!]
+\begin{figure}
 	%\centering
-	\includegraphics[scale=0.55]{images-paper/callgraph_cases}
+	%\includegraphics[scale=0.55]{images-paper/callgraph_cases}
+    \input{new_diagrams/callback_classification}
     \caption{Callgraph analysis: The bottom nodes represent functions
@@
 \section{Lessons Learned}
-%\textcolor{blue}{\Blindtext}
+In this section we summarize the insights gained from our observations during
+the implementation as well as the from the results obtained by our experimental
+evaluation.
+
+\paragraph{Features are problematic, but prevalently used.}
+From out observations with \sf{WordPress} in Section \ref{sec:3} we learn that
+dynamic language features, in particular include expressions and indirect function
+calls, in combination with symbolic values cannot be evaluated accurately since
+concrete values are required. For our observations we were able to explain most
+of the missed output candidates with indirect function calls. For include
+expressions we are indeed able to resolve most expressions successfully, yet we
+do not distinguish between static and dynamic includes in this case study as
+this would exceed the scope of our paper. However, we believe it is plausible
+to assume that defective include expressions can cause a cascade of missing
+files and subsequent includes.
+Throughout the experiment we have seen that dynamic features, are prevalent for
+modern PHP systems as they serve a genuine purpose in the programming language.
+Nevertheless their usage in situations when combined with assumed symbolic
+values restricts symbolic execution and any tool leveraging results obtained
+from that.
+
+\paragraph{Bypassing conceptual limitations.}
+Symbolic execution is used to approximate client page output since it allows to
+unfold the staged nature of dynamic web applications. Aside from conceptual
+limitations in the systems studied, limitations can be addressed. To use
+approaches incorporating symbolic execution with dynamic web applications in
+the long run, more concrete dummy information, similar to test oracles, needs
+to be provided to decrease the number of assumed symbolic inputs and avoid
+dynamic features being confronted with symbolic values.Also, if the usage of
+tools using static output approximation with symbolic execution is intended,
+avoiding dynamic features that are prone to be inaccurately executed for
+symbolic execution  may be an adjustment strategy to program in an analyzable
+way.
+
+\paragraph{Practical tools require practical evaluation.}
+Last, we want to address the limited informative value of analyses and tools
+presented in previous work \cite{minamide2005static,wassermann2007sound,Nguyen:2015:VIS:2819009.2819140,Nguyen:2015:CPS:2786805.2786872,Nguyen:2014:BCG:2635868.2635928,Nguyen:2011:AFH:2190078.2190142}
+ where only small PHP systems  have been studied. These papers introduce
+ techniques for static output approximation of PHP web applications [X, Y] or
+ propose tools based on respective approximations [Z, Wassermann]. As these
+ approaches address practical use cases, they require a more and further
+ comprehensive evaluation in order to  better understand possible limitations
+ and assess their practicality, in particular for large-scale systems.
+
+\paragraph{The road so far.}
+For this paper, the extension of the symbolic execution Symex [Hung] support
+for object-oriented programming was not the main goal, yet necessary to
+symbolically execute modern PHP systems. The observations regarding state state
+explosion for method calls documented in Section \ref{} are only anecdotal and,
+with the exception of \sf{WordPress}, we did not notice any significant impact
+on scalability. Nevertheless, more research is required to better and
+systematically understand possible drawbacks of object-oriented symbolic
+execution. This question though is beyond the scope of this paper, as we aimed
+to explore limitations of the practicality and scalability of static output
+approximation using symbolic execution. Future work might address design
+guidelines to enable programming analyzable applications, tool support to scale
+the number of assumed symbolic inputs or systematically provide concrete data
+to increase the practical benefit of the tools described above.
 
@@
 \cite{wassermann2007sound,minamide2005static}
-%\textcolor{blue}{\blindtext\blindtext}
+%\textcolor{blue}{\blindtext}
 

### 2017-01-17 paper.tex
rough draft related work

@@
 
-\author{Stefan Mühlbauer}
+%\author{Stefan Mühlbauer}
 \affiliation{%
-  \institution{Technische Universität Braunschweig}
+  %\institution{Technische Universität Braunschweig}
   %\streetaddress{P.O. Box 1212}
@@
 
-\author{Christian Kästner}
+%\author{Christian Kästner}
 %\orcid{1234-5678-9012}
 \affiliation{%
-  \institution{Carnegie Mellon University}
+  %\institution{Carnegie Mellon University}
   %\streetaddress{5000 Forbes Avenue}
@@
 
-\author{Tien N. Nguyen}
+%\author{Tien N. Nguyen}
 %\orcid{1234-5678-9012}
 \affiliation{%
-  \institution{University of Texas at Dallas}
+  %\institution{University of Texas at Dallas}
   %\streetaddress{P.O. Box 1212}
@@
 % The default list of authors is too long for headers}
-\renewcommand{\shortauthors}{\textcolor{blue}{put shortauthors here}}
+\renewcommand{\shortauthors}{
+	%\textcolor{blue}{put shortauthors here}
+}
 
@@
 	\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
+
 \end{itemize}
@@
 to be provided to decrease the number of assumed symbolic inputs and avoid
-dynamic features being confronted with symbolic values.Also, if the usage of
+dynamic features being confronted with symbolic values. Also, if the usage of
 tools using static output approximation with symbolic execution is intended,
@@
 way. 
+Future work to increase code coverage and approximation quality might involve
+automated test case generation as tools like \sf{Apollo}
+\cite{artzi_finding_2008,artzi_finding_2010} or previous
+work \cite{DynamicWassermann} to decrease the number of symbolic inputs and
+provide proper and concrete input variants.
 
@@
 \section{Related Work} \label{sec:related_work}
-\cite{Hills:2013:ESP:2483760.2483786}
-\cite{wassermann2007sound,minamide2005static}
+\paragraph{Static Output Approximation}
+To approximate program slices for web
+applications, Ricca et al. \cite{tonella_web_2005,tonella_2001,tonella_2002} approximate
+dynamically generated output. For output generating statements, such as \tt{echo} or
+\tt{print}, all strings are unquoted (code extrusion). If those statements
+contain variables, these are linked to string concatenations using a
+proposed flow analysis called \emph{string-cat propagation}. From these
+flows representing approximated output subsequent program slices are computed.
+
+Minamide \cite{minamide_static_2005} approximates client page output
+of web applications by describing possible output by a context-free
+grammar that is constructed statically from the PHP code for a given regular
+expression of user input. The constructed grammar enables analyses such as
+detecting cross-site scripting vulnerabilities by checking whether user input has been sanitized, and HTML validation by
+determining whether the constructed grammar is contained in a depth-bound
+HTML grammar. Based on Minamide's approximation approach several
+vulnerability analyses addressing cross-site scripting
+\cite{wassermann_static_2008} and SQL injection \cite{wassermann2007sound} have
+been proposed. Wang et al. \cite{wang_locating_2010} utilizes this string analysis to detect strings visible at the browser and enable
+internationalization of web applications.
+
+Another approach is proposed by Wang et al. \cite{wang_automating_2012}, where
+output for a web application is approximated using a hybrid approach: A
+dynamic webpage is executed with concrete input and the execution is
+recorded at runtime. Changes in the client-side output then can be mapped to 
+corresponding PHP code using static impact analysis.
+
+%Creating Test Input \cite{DynamicWassermann}
+
+Something about dynamic features and resolution
+  \cite{Hills:2013:ESP:2483760.2483786,hills2014static,hills2015variable}
+
+Tools using symbolic execution for static output approximation
+   \cite{artzi_finding_2008,artzi_finding_2010}
 %\textcolor{blue}{\blindtext}

### 2017-01-18 paper.tex
refactored 3 mainly, integrated tiens feedback

@@
 \usepackage{tikz}
+	\usepackage{pgf, pgfplots, pgfplotstable}
+	\usetikzlibrary{arrows,automata}
 	\usetikzlibrary{patterns}
@@
 \usepackage[zerostyle=d]{newtxtt}
-%\renewcommand*\familydefault{\ttfamily} %% Only if the base font of the document is to be typewriter style
 \usepackage[T1]{fontenc}
@@
 	basewidth={0.5em,0.4em}, % for other columns settings
-	commentstyle = \color{gray},
+	commentstyle = \color{black},
 	columns=fullflexible,
@@
 
-\usepackage{pgfplots, pgfplotstable}
-
-\usepackage{pgf}
-\usepackage{tikz}
-\usetikzlibrary{arrows,automata}
-
 \usepackage{xcolor}
@@
 % DOI
-\acmDOI{10.475/123_4}
+%\acmDOI{10.475/123_4}
 
 % ISBN
-\acmISBN{123-4567-24-567/08/06}
+%\acmISBN{123-4567-24-567/08/06}
 
@@
  California, USA}
-\acmYear{1997}
-\copyrightyear{2016}
-
-\acmPrice{15.00}
+%\acmYear{1997}
+%\copyrightyear{2016}
+
+%\acmPrice{15.00}
 
 \begin{document}
-\title{Does It Scale? Static Output Approximation of PHP Web Applications Using
-Symbolic Execution}
-\titlenote{Produces the permission block, and copyright information}
+\title{Does It Scale? Symbolic Execution of PHP Web Applications}
+%\titlenote{Produces the permission block, and copyright information}
 %\subtitle{Extended Abstract}
@@
 \begin{abstract}
+\textcolor{gray}{
 Dynamic web applications have become widely popular and are to
@@
 for vulnerability detection. Unfortunately, recent approximation approaches have
-only been evaluated for smaller systems.
-
-This paper presents an experience report about the scalability of output
+only been evaluated for smaller systems.}
+\textcolor{gray}{This paper presents an experience report about the scalability of output
 approximation using symbolic execution of state-of-the-art PHP web
@@
 tools and also to improve the development process of statically analyzable web
-applications.
+applications.}
 \end{abstract}
@@
 languages tailored specifically to the domain of web applications. PHP
-cite{phpNET} is a programming language focused on server-side application
+\cite{phpNET} is a programming language focused on server-side application
 development. As of 2012, it was used by 78.8 percent of the ten million most popular websites (according
@@
 
-Knowledge about different output output variants is leveraged by a number of
+Knowledge about different output variants is leveraged by a number of
 analyses for the domain of web applications. Previous work presented tools
@@
 as object-oriented programming. We evaluate our
-symbolic execution engine for large-scale and modern PHP systems.
+symbolic execution engine for several large-scale and modern PHP systems.
 
 During the introduction of new semantics, we addressed two  trade-offs between
-accuracy of our e xecution and performance: For method calls with ambiguous
+accuracy of our execution and performance: For method calls with ambiguous
 targets as well as concrete execution of loops the number of program states can
@@
 
-Based on out observations we identify conceptual limitations of symbolic
+Based on observations we identify conceptual limitations of symbolic
 execution for PHP.  Dynamic features such as indirect function calls and
@@
 these expressions may be assembled at runtime and can contain symbolic
-information, for many cases this restricts our engine from further execution.
+information, for many cases this restricts a symbolic execution engine from
+further execution.
 Although some effort to statically approximate include expression can be spent,
@@
 various inputs including databases, user inputs or configuration files. We
-empirically evaluate our observations for a corpus including large PHP systems that are still maintained.
+empirically evaluate our observations for a corpus including large, actively maintained PHP systems.
 
@@
 statically approximate the client-page output for PHP web applications using
-symbolic execution and (2) a report of out observations for state-of-the-art
+symbolic execution and (2) a report of observations for state-of-the-art
 web applications as well as an empirical evaluation of conceptual limitations
@@
 \section{State Of the Art}\label{sec:stateoftheart}
-Symbolic execution is an analysis techniques used by the tools
-whose practicality we aim to evaluate. This section recaps the idea of symbolic
+This section recaps the idea of symbolic
 execution, the tools which make use of it, and both conceptual and
@@
 \subsection{Symbolic Execution}
-Symbolic execution is a static analysis technique that was proposed by James C.
-King in 1976 \cite{King1976}. Symbolic execution allows to explore feasible
-paths in a program. Based on normal program execution, the execution semantics
+Symbolic execution is a static analysis technique that was proposed by King
+\cite{King1976} in 1976. Symbolic execution allows to explore feasible paths in a program. Based on normal program execution, the execution semantics
 is extended to handle symbolic values. An execution becomes symbolic by
@@
 \begin{lstlisting}
-<form>
+<form method="POST">
   <input type="text" name="name">
@@
 \end{lstlisting}
-		\caption{Snippet of HTML and PHP server-side web application code.
+		\caption{Snippet of server-side HTML and PHP web application code.
 		\label{fig:web_application}}
@@
 \begin{lstlisting}[mathescape]
-<form>
+<form method="POST">
   <input type="text" name="name">
@@
 		\caption{
-			Approximation client paege output of the server-side web application code in
-			Figure \ref{fig:web_application}. Variability as well as repetition annotations
-			are represented by preprocessor directives.
+			Approximated client page output of the server-side code in Figure
+			\ref{fig:web_application}. The directives in lines 6 and 8 highlight output
+			that can be repeated an arbitrary number of times, the directives in line 10
+			and 14 mark two alternative output parts.
 			\label{fig:output_example}
@@
 is illustrated in Figure \ref{fig:output_example}.
+
 First, for loops in the previous execution semantics
@@
 execution terminates. Hence, any output constructed during the single loop
-iteration is highlighted (\texttt{repeat \ldots endrepeat}) to possibly be repeated an arbitrary number of times.
-The output in line 7 is marked to be repeated an arbitrary number of times, as the number of loop iterations cannot be determined statically.
-Second, a similar procedure
-applies to recursive functions. Once called, a recursive function is  also only
-executed to recursion depth one per call, and any subsequent invocation of that
-function returns a symbolic value. This, again, is due to the difficulty to
-determine the exact recursion depth similar to the number of loop iterations.
+iteration is highlighted (\texttt{repeat \ldots endrepeat}) to possibly be
+repeated an arbitrary number of times. The output in line 7 is marked to be
+repeated an arbitrary number of times, as the number of loop iterations cannot
+be determined statically.
+
+Second, a similar procedure applies to recursive
+functions. Once called, a recursive function is  also only executed to recursion
+depth one per call, and any subsequent invocation of that function returns a
+symbolic value. This, again, is due to the difficulty to determine the exact
+recursion depth similar to the number of loop iterations.
+
 Third, any form of user interaction providing input data as well as any form of input referred
@@
 
-\subsection{Existing Tool Support for PHP}\label{sec:existing_tools}
+\subsection{Symbolic-Execution-based Tools For PHP}\label{sec:existing_tools}
 All tools proposed so far leverage a representation of all possible HTML client
@@
 	\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
-
+	\item \textcolor{gray}{Automated Web Application Testing Using Seacrch
+	Software Engineering} \cite{Alshahwan2011}
 \end{itemize}
@@
 
-\section{Scalable Output Approximation} \label{sec:3}
+\section{Scalable Output Approximation?} \label{sec:3}
 The main objective of this experience report is to investigate whether we can
-have a scalable symbolic execution engine as it is the underlying technique for
-static output approximation in previous work
+build a scalable symbolic execution engine as it is the underlying technique
+for static output approximation in previous work
 \cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140}.
-Our vision is a practical symbolic execution, i.e., it computes an accurate approximation for large and modern PHP systems with reasonable time and space consumption.
-
-Also, we are interested in the limitations of scalable symbolic execution for
-PHP web applications. Limitations may either be conceptional, where symbolic
-execution of PHP code, in contrast to concrete execution, is inaccurate: For
-example an include expression evaluated to a symbolic value simply provides to
-little information to include a script. Or, in spite of infeasible effort
-required, limitations could allow an accurate symbolic execution. For example,
-exhaustively exploring all feasible paths in an application may result in an
-infeasible number of program states to store; nevertheless, this so called
-state explosion problem may be addressed with additional engineering.
-
-\subsection{Towards Scalable Output Approximation?}
-As our goal is to have a scalable symbolic execution engine, we approach this
-question by re-implementing the symbolic execution semantics described in
-previous work \cite{Nguyen:2014:BCG:2635868.2635928} and extending it with
-support for additional language features. We aim to support symbolic execution
-for a real-world example system of reasonable size that incorporates the
-missing language features described in Section \ref{sec:limitations} and is
-still developed and maintained. \textsf{WordPress} is a popular open-source
-Content Management System (CMS) with around \textcolor{blue}{300.000} lines of
-code that provides a vivid plug-in environment and is maintained by a large
-community.
-For our implementation of a symbolic execution engine for PHP,
-\textsf{Oak}\footnote{The implementation can be found
-at \url{github.com/smba/oak}.}, we have chosen a test-driven approach with
-continuous integration using regression tests based on \textsf{WordPress} and \textsf{SchoolMate} (see Section \ref{sec:experiment_setup}). Oak incorporates the existing execution semantics
-of \cite{Nguyen:2014:BCG:2635868.2635928} and extends it with support for
-object-oriented orientation (class definition, class instantiation and method
-invocation) as well as most PHP array operations. A list of the functions
-provided by the PHP standard library can be found at \cite{PHP}.
-Throughout the implementation we did not achieve high code coverage for
-\sf{WordPress}. In the following we present our observations we made thorughout
-the implementation regarding issues impeding symbolic execution.
+Our vision is a practical symbolic execution, i.e., it computes an accurate
+approximation for large and modern PHP systems with reasonable time and space consumption.
+We approach this goal by re-implementing the symbolic execution semantics  of
+\sf{PhpSync} \cite{Nguyen:2014:BCG:2635868.2635928} and extending it with support for additional language
+including object-oriented programming and array operations. We aim to support symbolic
+execution for a real-world example system of reasonable size that incorporates
+the missing language features described in Section \ref{sec:limitations}.
+\textsf{WordPress} is a popular open-source Content Management System (CMS)
+with around 300.000 lines of code that provides a vivid
+plug-in environment and is used.  The full implementation of \sf{Oak}, the test
+suite as well as or test corpus (see Section \ref{sec:experiment_setup}) can be found at
+\url{http://www.github.com/smba/oak}.
+
+For our implementation of a symbolic execution engine for PHP, Oak, we have
+chosen a test-driven approach with continuous integration using regression
+tests based on \sf{WordPress} and \sf{SchoolMate} (see Section
+\ref{sec:experiment_setup}). For \sf{SchoolMate}, a rather straightforward and
+simple system, which we chose to compare our approximation results to those of
+\sf{PhpSync}, we could achieve similar output approximation results. For
+\sf{WordPress} though, we did not achieve high code coverage or an accurate
+approximation for \sf{WordPress} as we faced a number of obstacles hindering
+symbolic execution in different ways: The following Subsection
+\ref{sec:experience_report} documents our design choices we made, apparent trade-offs we recognized, and limitations
+we faced on our way towards a scalable symbolic execution engine; and what we
+have learned from bypassing limitations faced in spite of manual effort for
+future work. Subsection \ref{sec:experience_results} compares our observations
+and results to \emph{PhpSync} and presents the motivation for our empirical evaluation of our
+tool symbolic execution engine in Section \ref{sec:evaluation}.
 
 \subsection{Experience Report}\label{sec:experience_report}
-We identified trade-offs between accuracy and performance that may
-cause the approximation to be impractical, yet possible. Additionally, we
-detected limitations impeding symbolic execution conceptually: For certain
-language features, concrete information rather than symbolic values may be required to
-actually be executed properly as this defective execution can impact subsequent
-parts of the program which may become inaccessible and may be missed.
-
-\subsubsection{Symbolic vs Concrete Execution}
-Throughout the implementation, we basically followed the symbolic execution
-semantics described in previous work \cite{Nguyen:2014:BCG:2635868.2635928}.
-Nevertheless, for additional language features we experienced it to be beneficial to the
-approximation to execute them as precisely as possible.
-For instance, \sf{WordPress} uses a text filter mechanism, where different
-filter functions are applied sequentially to a string literal. Filters can
-be linked to a hook using  associative arrays so that when the hook is
-triggered, all filters that are linked to it are applied. As described in
-Section \ref{} we execute loops only once due to an possibly
-indeterminable number of iterations. PHP yet offers more loop constructs: the
-do-while loop and the foreach loop. While for the do-while loop as well as for
-the while loop the number of iterations is possibly indeterminable, the foreach
-loop iterates over all items of a given array whose number of items is known at
-runtime. For \sf{WordPress}, we are able to achieve a more accurate
-approximation and higher code coverage results by concretely executing foreach
-loops as more server-side code (filter functions in our case) are accessed and
-executed. In turn, exhaustively executing foreach loops and filter functions
-concretely is a trade-off between accuracy and performance, as a single hook
-call site can be called from various contexts since we are exploring multiple
-execution paths.
-
-An additional aspect on the trade-off between concrete and symbolic execution,
-or effort and accuracy respectively, is the extent to which program inputs are
-assumed to be symbolic. As shown in Section \ref{} introducing symbolic values
-can be problematic if they are propagated to language features requiring
-concrete information, such as include expressions or indirect function calls.
-We manually scaled the number of sources of symbolic inputs by providing static
-dummy inputs such as a concrete form input, file name or configuration entry.
-For specific cases, where concrete information is inevitably required, we can
-achieve higher code coverage and a more accurate approximation, as more code
-becomes accessible. Nevertheless, while detecting dynamic features can be
-automated, it requires manual inspection and understanding of the web
-application to provide appropriate dummy input data. Again, this is a trade-off
-between the effort to provide input data and accuracy of the approximation and
-code coverage respectively.
-
-\subsubsection{Multi-Target Method Calls}
-A non-conceptual problem we encountere is related to object-oriented programming. While exploring
-different execution paths when the control flow is split, a variable can have
-different values depending on the path condition under which a value has been
-assigned to it (see the code example in Figure \ref{fig:output_example}).
-Similarly, a variable can point to different objects depending on the path
-condition as the snippet in Figure \ref{fig:multi_target} illustrates. The variable \tt{\$person} can either
-point to object \tt{Alice} or \tt{Bob}. In addition, for any method invocated on
-that variable the actual target is ambiguous as there are two, depending on the
-path condition. In order to have a sound symbolic execution, both method calls
-need to be executed. Nevertheless, the number of contexts for which a method can
-be called can grow rapidly, and easily become infeasible to execute
-exhaustively.
-
-We implemented several heuristic approaches to tame this state space explosion
-for multi-target method calls, including executing a method only for a subset of
-target objects, or just a single object. This is a trade-off between accuracy
-and soundness of the symbolic execution and performance time required.
+
+\subsubsection{Symbolic/Concrete Loop Execution}
+Throughout the implementation we basically followed the symbolic execution
+semantics of \sf{PhpSync} \cite{Nguyen:2014:BCG:2635868.2635928}. As described
+in Section \ref{sec:workflow} for loops, only one iteration is executed since
+the number of iterations is indeterminable for most cases. While this is the
+case for while and do-while loops, the number of loop iterations for for or
+foreach loops is generally determinable and known before entering the loop
+body. For loops use a loop counter that usually is incremented or decremented
+after each loop iteration until a loop condition is not satisfied anymore;
+foreach loops simply iterate over a given array value.
+Our execution semantics follows this scheme and simulates concrete execution
+for those types of loops if possible. Otherwise, for instance, if a loop
+counter, loop counter limit for for loops, or variable for foreach loops is
+symbolic, we resort to executing only a single iteration. This design choice
+was made since it is more precise and, in contrast to only considering one
+iteration in all cases, slightly increased code coverage for \sf{WordPress}.
+
+\subsubsection{Method Target Space Explosion}
+For method invocations that for concrete execution have only one target object,
+exploring multiple execution paths can cause the space of target objects to
+grow rapidly. A minimal example is shown in the listing in Figure
+\ref{fig:multi_target} where two different objects can be assigned to variable \tt{\$person} in line 12 and
+14. For the method invocation in line 16 hence, the method is executed twice in
+total since we have two alternative target objects. To address this problem, we
+implement program state merging for symbolic execution. For the example in
+Figure \ref{fig:multi_target}, after the control flow is split in line 11 into
+two branches, we merge both branches together subsequently. For both branches two different
+values have been assigned to variable \tt{\$person} under different path
+conditions respectively. We implement state merging inspired by work of Sen et al. 
+\cite{Sen:2015:MMS:2786805.2786830} using value summaries. A value summary can be conceived as a value
+type. For state merging in our example, in the value summary, each value (or
+each object in our case) is associated to the path condition under that it has
+been assigned to a variable. This would allow to only pass the value summary to
+the method and only invocate it once, but for our execution to be sound, all
+objects would need to be of the same class type. If for instance, in our
+example there would be another class definition with a method named
+\tt{greet} and an instance of that class is assigned to variable \tt{\$person}
+in one of the two branches, the method invocation in line 16 actually needs to be
+executed for two different target objects of a different type.
+
+Therefore, for our execution engine to be sound, we were bound to execute
+invocated methods for multiple targets since we cannot assert that all values
+contained in summary are of the same type. Nevertheless, to scale the target
+object space, in addition, we implemented execution modes only allowing
+execution of a method for a subset of possible targets and assuming symbolic
+return values for the remaining targets. This execution mode was necessary for
+WordPress for several, but not all entry points that we studied, in order to
+terminate. All subsequent analyses in this paper do not use this execution
+mode, yet it is provided in out implementation.
 
 \subsubsection{Dynamic Language Features}
-\label{sec:experiment_dynamicfeaturess} 
-Conceptually, we identified two language features for which symbolic execution
-failed frequently. As stated in Section \ref{sec:workflow}, dynamic web
-applications are staged and client page output is computed at runtime. Symbolic
-execution has shown to be a useful technique to approximate dynamically
-generated output in theory, though it is challenged by language features that
-require concrete information rather than symbolic information at runtime. In particular, this
-applies to language features that provide access to reusable code of the web
-application: include expressions and function calls.
-
-First, include expressions in PHP usually require a string value representing
-the path to the script file or template to include. If this string value is
-statically provided by a string literal or can be constructed unambiguously, the
-symbolic execution engine was able to resolve the include expression and include
-the desired file. However, it is also common to assemble include expressions not
-only from static string literals, but from several sources including user input
-information, database query results or configuration files. Figure
-\ref{fig:dynamic_includes} shows a small example of a dynamic include containing
+\label{sec:experiment_dynamicfeatures} 
+Aside from the previous issues, we identified two dynamic language features
+that we were not able to execute accurately. Dynamic features in PHP are
+features that take as input rather a expression that is resolved at runtime
+than a static constant value. For \sf{WordPress}, we faced missing files and
+functions that are included using dynamic includes and dynamic function
+invocation respectively.
+
+First, file inclusion in PHP usually is managed by statements like \tt{include},
+\tt{include\_once} or \tt{require} and \tt{require\_once}. For transitive
+includes, statements ending with \tt{\_once} will not include the same file
+twice. The include statements make an attempt to include a file, but do not
+return a warning in contrast to the require statements. Each of these
+statements takes as input either a string constant (static include) or an
+expression that is evaluated at runtime (dynamic include). The string that the
+expression evaluates to represents an absolute or relative paths to the target
+file that the interpreter finally attempts to include.
+Figure \ref{fig:dynamic_includes} shows a small example of a dynamic include containing
 information retrieved from a database. Since any input for a symbolically
@@
 
-\subsubsection{Interpreter Customization}
-We have approached issues we encountered, including
-multiple-target method calls and foreach loops, in different ways as these
-represent a trade-off between accuracy of execution and performance.
-For method calls having ambiguous targets, we implemented different modes of
-execution, ranging from a complete mode where a method is executed for all
-possible targets to abstract modes with either a subset of targets or only a
-single target. In the latter cases where targets are discarded, we assumed
-symbolic values as method return values. Our symbolic interpreter can be
-customized to work in the desired mode to scale accuracy or performance.
-Nevertheless, for \sf{WordPress} these variations did not have significant
-impact on the output approximation as symbolic execution was rather impeded by
-dynamic features.
-Also, for foreach loops we are able to toggle between a single iteration mode
-and a complete iteration mode. Although we were able to increase code coverage
-by accessing more code, it did not significantly increase our output
-approximation.
-Note that these variants have only been tested for \sf{WordPress} to better
-understand possible trade-offs or simply achieve any output approximation in a
-reasonable period of time. We believe based on our experience dynamic features
-to be the main conceptual limitation impeding code coverage as well as output
-approximation quality. In the next Section we analyze in detail the impact of
-dynamic features with concrete foreach loop execution and complete
-multiple-target execution for method calls.
-
-\section{Evaluation}
+\subsubsection{Symbolic/Concrete Input}
+In the previous subsection, we argued that dynamic includes and dynamic
+function invocations hinder symbolic execution  due to imprecise resolution of
+expression if symbolic information is contained. For callable code parts as
+included scripts or invocated functions, we attempted to bypass this limitation
+by manually providing proper concrete input.
+
+Consider the code example in Figure \ref{fig:indirect_calls} where a function
+name could not be resolved precisely due to a missing string literal that is contained from a
+database query result. By manually providing either option \tt{serif} or
+\tt{sserif}, or both as alternatives, we are able to increase code coverage and
+the accuracy of our approximation with manual effort by making dynamic
+statements static. Of course we did not provide proper input exhaustively for
+every occurrence of dynamic features, yet this illustrates a basis for possible
+future work towards scalable symbolic execution. This manual approach addresses
+the same problem as dynamic test input generation for web applications, where
+proper concrete input is assumed rather than symbolic values to increase code
+coverage. We present work on this problem in Section \ref{sec:related_work}.
+
+\subsection{Experience Results} \label{sec:experience_results}
+The symbolic execution engine \sf{PhpSync} \cite{Nguyen:2014:BCG:2635868.2635928} so far does not face the
+problems we documented in Section \ref{sec:experience_report}. The evaluation of
+PhpSync, i.e., the PHP systems that the tool was applied to, did neither contain object-oriented
+features, nor dynamic features like dynamic includes or dynamic function
+invocation with symbolic values.
+
+As the observations so far did only refer to \sf{WordPress} as a large-scale
+system, we wanted to study the impeding impact of dynamic features to symbolic
+execution for a wider selection of PHP systems systematically apart from the
+systems that \sf{PhpSync} was applied to for two reasons. First, the usage of
+dynamic features, not only of includes and function invocations, varies from
+system to system \cite{hills2015variable}. Second, we were not able to obtain results for
+all entry points possible for \sf{WordPress} and hence could not compare our
+approximatio results to those of \sf{PhpSync} where client page output
+approximation is evaluated for all entry points. Therefore, we decided to
+systematically evaluate our symbolic execution engine for systems apart from
+\sf{WordPress}, which is documented in the following Section
+\ref{sec:evaluation}.
+
+\section{Evaluation} \label{sec:evaluation}
 As already stated in the previous Section \ref{sec:experience_report}, we
@@
 classifier from, but could not do so without decreasing precision. 
-\textcolor{blue}{In spite of missing half of the output candidates, we decided to use a simple, yet
-precise and distinctive classifier.}
+\textcolor{gray}{In spite of missing half of the output candidates, we decided
+to use a simple, yet precise and distinctive classifier.}
 
@@
 candidates are reached by an execution path and part of the output
-approximatiopn define two metrics to approach measuring accuracy.
+approximatiopn define two metrics to approximate measuring accuracy.
 
@@
 \label{WhyDidIncludesFail}
-From Section \ref{sec:experiment_dynamicfeaturess} we learn that include
+From Section \ref{sec:experiment_dynamicfeatures} we learn that include
 expressions can not be evaluated accurately for symbolic values, and from the
@@
 partially accessibly through indirect calling mechanisms (see Section
-\ref{sec:experiment_dynamicfeaturess}). As Figure
+\ref{sec:experiment_dynamicfeatures}). As Figure
 \ref{fig:output_candidate_explanation} illustrates, around 80 percent (except
@@
 
-\paragraph{Features are problematic, but prevalently used.}
+\paragraph{Dynamic features are problematic.}
 From out observations with \sf{WordPress} in Section \ref{sec:3} we learn that
@@
 Last, we want to address the limited informative value of analyses and tools
-presented in previous work \cite{minamide2005static,wassermann2007sound,Nguyen:2015:VIS:2819009.2819140,Nguyen:2015:CPS:2786805.2786872,Nguyen:2014:BCG:2635868.2635928,Nguyen:2011:AFH:2190078.2190142}
+presented in previous work
+\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140,minamide_static_2005,wassermann2007sound}
 where only small PHP systems have been studied. These papers introduce
- techniques for static output approximation of PHP web applications [X, Y] or
- propose tools based on respective approximations [Z, Wassermann]. As these
- approaches address practical use cases, they require a more and further
- comprehensive evaluation in order to  better understand possible limitations
- and assess their practicality, in particular for large-scale systems.
+techniques for static output approximation of PHP web applications
+\cite{minamide_static_2005,Nguyen:2014:BCG:2635868.2635928,wang_automating_2012} or propose tools
+based on respective approximations
+\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140,wassermann2007sound,wassermann_static_2008}.
+As these approaches address practical use cases, they require a more and further comprehensive evaluation in order to  better understand possible limitations and assess their practicality, in particular for large-scale systems.
 
 \paragraph{The road so far.}
-For this paper, the extension of the symbolic execution Symex [Hung] support
-for object-oriented programming was not the main goal, yet necessary to
-symbolically execute modern PHP systems. The observations regarding state state
-explosion for method calls documented in Section \ref{} are only anecdotal and,
-with the exception of \sf{WordPress}, we did not notice any significant impact
-on scalability. Nevertheless, more research is required to better and
-systematically understand possible drawbacks of object-oriented symbolic
-execution. This question though is beyond the scope of this paper, as we aimed
+For this paper, the extension of the symbolic execution \sf{PhpSync} \cite{Nguyen:2014:BCG:2635868.2635928}
+support for object-oriented programming was not the main goal, yet necessary to
+symbolically execute modern PHP systems. The observations regarding object
+target space explosion for method calls documented in Section
+\ref{sec:experiment_dynamicfeatures} are only anecdotal and, with the exception
+of \sf{WordPress}, we did not notice any significant impact on scalability.
+Nevertheless, more research is required to better and systematically understand
+possible drawbacks of object-oriented symbolic execution. 
+This question though is beyond the scope of this paper, as we aimed
 to explore limitations of the practicality and scalability of static output
@@
 expression of user input. The constructed grammar enables analyses such as
-detecting cross-site scripting vulnerabilities by checking whether user input has been sanitized, and HTML validation by
-determining whether the constructed grammar is contained in a depth-bound
-HTML grammar. Based on Minamide's approximation approach several
-vulnerability analyses addressing cross-site scripting
-\cite{wassermann_static_2008} and SQL injection \cite{wassermann2007sound} have
-been proposed. Wang et al. \cite{wang_locating_2010} utilizes this string analysis to detect strings visible at the browser and enable
-internationalization of web applications.
+detecting cross-site scripting vulnerabilities by checking whether user input
+has been sanitized, and HTML validation by determining whether the constructed
+grammar is contained in a depth-bound HTML grammar. Based on Minamide's
+approximation approach several vulnerability analyses addressing cross-site
+scripting \cite{wassermann_static_2008} and SQL injection
+\cite{wassermann2007sound} have been proposed. Wang et al.
+\cite{wang_locating_2010} utilizes this string analysis to detect strings
+visible at the browser and enable internationalization of web applications.
 
@@
 
-%Creating Test Input \cite{DynamicWassermann}
-
-Something about dynamic features and resolution
-  \cite{Hills:2013:ESP:2483760.2483786,hills2014static,hills2015variable}
-
-Tools using symbolic execution for static output approximation
-   \cite{artzi_finding_2008,artzi_finding_2010}
+\paragraph{Dynamic Features}
+According to a case study by Hills et al. \cite{Hills:2013:ESP:2483760.2483786}
+of the feature usage in PHP, based on a corpus of state-of-the-art-projects,
+dynamic includes are less frequently used than static includes, but usage
+frequency is varying from system to system. Further work by Hills et al.
+\cite{hills2014static,hills2014php} approaches static resolution of dynamic includes using
+both context-insensitive resolution on file level by simplifying PHP constants; and context-sensitive on program level for transitive includes. In spite of
+promising results, these approaches, similar to our results, face limitations
+for truly dynamic includes if resolution is not sound due to information not
+being available in the source code like database query results. Something about
+dynamic features and resolution.
+
+%Tools using symbolic execution for static output approximation \cite{artzi_finding_2008,artzi_finding_2010}
 %\textcolor{blue}{\blindtext}

### 2017-01-19 paper.tex
draft

@@
 easing development and maintenance by extending IDE support for web applications
-with code navigation
-\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
+with code navigation \cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
+Moreover, knowledge of output variants can increases code coverage for
+output-oriented testing \cite{Alshahwan2011}.
 
@@
 	\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
-	\item \textcolor{gray}{Automated Web Application Testing Using Seacrch
-	Software Engineering} \cite{Alshahwan2011}
+
 \end{itemize}
@@
 execution for a wider selection of PHP systems systematically apart from the
-systems that \sf{PhpSync} was applied to for two reasons. First, the usage of
-dynamic features, not only of includes and function invocations, varies from
-system to system \cite{hills2015variable}. Second, we were not able to obtain results for
-all entry points possible for \sf{WordPress} and hence could not compare our
-approximatio results to those of \sf{PhpSync} where client page output
-approximation is evaluated for all entry points. Therefore, we decided to
+systems that \sf{PhpSync} was applied to for two reasons.
+
+First, the usage of dynamic features, not only of includes and function invocations, varies from
+system to system \cite{hills2015variable}.
+
+Second, we were not able to obtain results for all entry points possible for
+\sf{WordPress} and hence could not compare our approximation results to those of
+\sf{PhpSync} where client page output approximation is evaluated for all entry points. Therefore, we decided to
 systematically evaluate our symbolic execution engine for systems apart from
@@
 First, the evaluation of the output approximation using the symbolic execution
-engine, \sf{Symex}, in previous work \cite{Nguyen:2014:BCG:2635868.2635928} has
-shown high code coverage for small systems. As our symbolic execution
-re-implements \sf{Symex} and extends with support for additional language
-features, we ask whether we can replicate high coverage results at least on a small-scale. Therefore, our PHP
-corpus contains four small-scale systems used in previous work \cite{Nguyen:2014:BCG:2635868.2635928}.
-
-Second, we ask whether system size and usage of modern language features are
-factors affecting the practicality of our symbolic execution engine. We
+engine, \sf{PhpSync}, in previous work \cite{Nguyen:2014:BCG:2635868.2635928}
+has shown high code coverage for small systems. As our symbolic execution
+re-implements \sf{PhpSync} and extends with support for additional language
+features, we ask: Can  we replicate high coverage results at least on a
+small-scale. Therefore, our PHP corpus contains four small-scale systems used
+in previous work \cite{Nguyen:2014:BCG:2635868.2635928}.
+
+Second, we ask :Do system size and usage of modern language features
+factors affect the practicality of our symbolic execution engine? We
 selected both a range of large-scale and small-scale modern PHP systems to
@@
 classifier from, but could not do so without decreasing precision. 
-\textcolor{gray}{In spite of missing half of the output candidates, we decided
-to use a simple, yet precise and distinctive classifier.}
+In spite of missing half of the output candidates, we decided
+to use a simple, yet precise and distinctive classifier. We rather work on a
+smaller sample subset of actual output candidates with less false positives
+than a larger set of string literals, where more are mistakenly classified as
+output candidates.
 
@@
 for truly dynamic includes if resolution is not sound due to information not
-being available in the source code like database query results. Something about
-dynamic features and resolution.
+being available in the source code like database query results.
+
+\paragraph{Input Generation for PHP}
+\textcolor{gray}{Apollo \cite{artzi_finding_2008,artzi_finding_2010}, automated
+test input generation using concolic testing \cite{DynamicWassermann}, test case generation
+using search based software engineering \cite{Alshahwan2011}}
 

### 2017-01-21 paper.tex
related work finished, conclusion draft

@@
 
-\definecolor{lightgrey}{HTML}{FFFFFF}
+\definecolor{lightgrey}{HTML}{FCFCFF}
 
@@
 %\titlenote{Produces the permission block, and copyright information}
-%\subtitle{Extended Abstract}
+\subtitle{Output-oriented Symbolic Execution for Static Output Approximation}
 %\subtitlenote{The full version of the author's guide is available as  \texttt{acmart.pdf} document}
@@
 
-One common property of all technologies for dynamic web applications is
-\emph{staged computation}: A dynamic web application as a whole
+A common characteristic of all dynamic web applications is their
+\emph{staged nature}: A dynamic web application as a whole
 consists of both static code, such as scripts, and dynamically generated code,
@@
 
-Second, we ask :Do system size and usage of modern language features
+Second, we ask: Do system size and usage of modern language features
 factors affect the practicality of our symbolic execution engine? We
@@
 selections we were bound to select a smaller number of system than we intended
-to since the parser used in our implementation, \sf{Quercus}\footnote{Quercus
-footnote}, did not support a number of language features. The full list of PHP
-systems for that we evaluate our symbolic execution engine along with
-descriptive code metrics can be found in Table \ref{corpus}.
+to since the parser used in our implementation,
+\sf{Quercus}\footnote{\sf{Quercus} is an open-source Java implementation of PHP
+5 by Caucho Technology. We reused the parser front-end for our implementation.
+\sf{Quercus} can be obtained from \url{http://quercus.caucho.com}.}, did not
+support a number of language features.
+The full list of PHP systems for that we evaluate our symbolic execution engine along with descriptive code metrics can be found in Table \ref{corpus}.
 
@@
 		\input{new_diagrams/not_reached}
-		\caption{\label{coverage}}
+		\caption{
+		Classification distribution of \emph{reached} output candidates
+		(Reached/Output, OnlyReached) and \emph{not-reached} output candidates
+		(FunctionContext, NotImported, Misc) during the analyis for each system.
+		\label{coverage}}
 	\end{subfigure}
+	\vspace{1mm}
 
@@
 		\caption{
-			Include coverage and accuracy.
+			Classification distribution of \emph{reached} output include expressions
+			(Reached/Output, OnlyReached) and \emph{not-reached} include expressions.
 			\label{fig:include_coverage_results}
@@
 	\end{subfigure}
+	\vspace{1mm}
 
@@
 		\caption{
-			Distribution dead function output candidates that are par-
-			tially or completely explained due to callback candidates.
+			Distribution of \emph{not-reached} output candidates (in function contexts)
+			that are only (Complete), partially (Partial), or not accessible via callback candidates. 
 			\label{fig:output_candidate_explanation}
@@
 	\end{subfigure}
-	\caption{Results}
+	\caption{Evaluation results for output candidates, include expressions and
+	callgraph analysis for not-reached output candidates}
 \end{figure}
 
-\subsubsection{What literals did we miss?}\label{WhatLiteralsDidWeMiss?}
+\subsubsection{Where did we miss output?}\label{WhatLiteralsDidWeMiss?}
 Our initial approach to understand missed output candidates is to find out
@@
 
-\subsubsection{Why did includes fail?}
+\subsubsection{(Dynamic) Includes}
 \label{WhyDidIncludesFail}
@@
 
-\subsubsection{Why did function calls fail?}
+\subsubsection{Dynamic Function Invocations}
 \label{WhyDidFunctionCallsFail}
@@
 approaches incorporating symbolic execution with dynamic web applications in
-the long run, more concrete dummy information, similar to test oracles, needs
+the long run, more concrete input, similar to test oracles, needs
 to be provided to decrease the number of assumed symbolic inputs and avoid
@@
 way.
-Future work to increase code coverage and approximation quality might involve
-automated test case generation as tools like \sf{Apollo}
-\cite{artzi_finding_2008,artzi_finding_2010} or previous
-work \cite{DynamicWassermann} to decrease the number of symbolic inputs and
-provide proper and concrete input variants.
 
@@
 \section{Related Work} \label{sec:related_work}
+
+\paragraph{Symbolic Execution}
+Symbolic execution \cite{King1976,Darringer1978} remained an conceptual approach
+for decades, but has been refined and extended over the last years. Several
+approaches incorporating constraint solving and leveraging concrete execution
+anlongside symbolic execution emerged and enable automated test case generation
+and testing; the scalability of these approaches is still challenged by path
+explosion and constraint solving \cite{CadarSen2013}.
+
+Dynamic symbolic execution \cite{CadarSen2013} incorporates symbolic execution
+alongside concrete execution and ahs shown promising results in automated test
+generation and testing \cite{artzi_finding_2008,artzi_finding_2010,DynamicWassermann} for PHP web
+applications and might help mitigate the conceptual limitations faced with
+symbolic values and dynamic features for future work on output-oriented
+symbolic execution.
+
+An overview about recent evolution of symbolic execution is provided by work of Cadar et al.
+\cite{CadarSen2013}, as well as a more tool-centered perspective \cite{Cadar2011}.
+
 \paragraph{Static Output Approximation}
@@
 
-\paragraph{Input Generation for PHP}
-\textcolor{gray}{Apollo \cite{artzi_finding_2008,artzi_finding_2010}, automated
-test input generation using concolic testing \cite{DynamicWassermann}, test case generation
-using search based software engineering \cite{Alshahwan2011}}
-
 %Tools using symbolic execution for static output approximation \cite{artzi_finding_2008,artzi_finding_2010}
@@
 \section{Conclusion}
+In this paper we have addressed symbolic execution of PHP web applications that
+is used for client page output approximation. This approximations are leveraged
+be several developer tools for PHP, yet have not been evaluated for larger
+systems. We re-implemented and extended an existing symbolic execution engine
+to analyze several real-world PHP applications with the intention to evaluate
+the scalability of symbolic execution for PHP. Our work illustrates the
+necessity of more comprehensive evaluation of developer tools as
+previous work did not explore these limitations. Dynamic features of PHP in
+combination with symbolic values though hinder precise symbolic execution and
+limit the scalability and practicality of tools on symbolically approximated
+client page output.
+
+In the future we would like to use dynamic symbolic execution for output
+approxiamtion and possibly bypass existing limitations, as dynamic symbolic
+execution has shown to be promising for automated testing and verification.
 

### 2017-01-21 new_diagrams/reverse.py

@@ -1,0 +1,8 @@
+f = open("text.txt", "rb")
+s = f.readlines()
+f.close()
+f = open("newtext.txt", "wb")
+s.reverse()
+for item in s:
+  print>>f, item
+f.close()

### 2017-01-22 paper.tex
updated table

@@
 Based on observations we identify conceptual limitations of symbolic
-execution for PHP.  Dynamic features such as indirect function calls and
+execution for PHP.  Dynamic features such as indirect function invocations and
 include expression require concrete information to be evaluated properly. Since
@@
 future work. Subsection \ref{sec:experience_results} compares our observations
-and results to \emph{PhpSync} and presents the motivation for our empirical evaluation of our
-tool symbolic execution engine in Section \ref{sec:evaluation}.
+and results to \sf{PhpSync} and presents the motivation for our empirical
+evaluation of our tool symbolic execution engine in Section \ref{sec:evaluation}.
 
@@
 
-Second, function can either be called directly by statically providing the name
-of the function (or method), or indirectly. Indirect function calls in PHP are
-enabled by using built-in functions like \tt{call\_user\_func},
+Second, function can either be invocated directly by statically providing the
+name of the function (or method), or indirectly. Indirect, or dynamic function
+invocation in PHP are enabled by using built-in functions like \tt{call\_user\_func},
 \tt{call\_user\_func\_array} or simply passing a string value with a direct
@@
 Again, we encountered issues similar to dynamic include expressions with
-indirect function call mechanisms: The example in Figure
 \ref{fig:indirect_calls} illustrates two functions \tt{header\_serif} and
@@
 	\caption{
-		Dynamic dispatch: Multiple targets for method calls
+		Dynamic Dispatch : Multiple target objects for method invocation \tt{greet()} 
 		\label{fig:multi_target}
@@
 	\caption{
-		Dynamic Include Resolution
+		Dynamic Include Resolution: The value of \tt{\$template} is retrieved from a
+		database query result.
 		\label{fig:dynamic_includes}
@@
 	\caption{
-		Function calls by indirect invocation mechanisms
+		Dynamic Function Invocation using \tt{call\_user\_func}
 		\label{fig:indirect_calls}
@@
 	\caption{
-		Code Examples of defective code features
+		Code examples for dynamic features
 		\label{fig:defective_code_features}
@@
 
+% count lines
+%find . -regextype "posix-extended" -iregex '.*\.(php|bit|inc|module)' | xargs wc -l
+
+% count files
+%find . -regextype "posix-extended" -iregex '.*\.(php|bit|inc|module)' | wc -l
+
 \begin{table*}[t]
 \centering 
-	\begin{tabular}{lrp{4cm}rrrr}
+	\begin{tabular}{lrp{5cm}rrrr}
 	\toprule
 	\textbf{System} & \textbf{Version} & \textbf{Classification} & \textbf{SLOC} &
-	\textbf{\#files} & \textbf{\#OCs} & \textbf{\#includes}\\
+	\textbf{\#files} & \textbf{\#Output candidates} & \textbf{\#Include
+	Expressions}\\
 	\midrule
-	AddressBook & 8.2.5.2 &  & 51,907 & 239  & 1009 & 186\\
-	SchoolMate & 1.5.4 &  & 8,118 & 65  & 853 & 88\\
-	TimeClock & 1.04 &  & 20,800 & 63  & 7920 & 306\\
-	WebChess & 1.0.0 &  & 5,219 & 28  & 470 & 56\\
+	AddressBook & 8.2.5.2 & Web-based address and phone book & 51,907 & 239  & 1009
+	& 186\\
+	SchoolMate & 1.5.4 & School administration & 8,118 & 65  & 853 & 88\\
+	TimeClock & 1.0.4 & Web-based time tracking & 20,800 & 63  & 7920 & 306\\
+	WebChess & 1.0.0 & Chess server software & 5,219 & 28  & 470 & 56\\
 	\midrule
-	Drupal & 7.5.0 & CMS & 52,464 & 125  & 3569 & 749\\
-	phpBB & 3.1.9 &  & 327,371 & 1,398  & 3606 & 206\\
-	phpMyAdmin & 4.6.3 &  & 303,582 & 871  & 7103 & 571\\
+	Drupal & 7.5.0 & Content Management System &  52,464 & 125  & 3569 & 749\\
+	phpBB & 3.1.9 & Forum & 327,371 & 1,398  & 3606 & 206\\
+	phpMyAdmin & 4.6.3 & Web-based MySQL administration & 303,582 & 871  & 7103 &
+	571\\
 	\midrule
-	Anchor & 0.12.1 &  & 15,054 & 201 & 987 & 32\\
-	Kirby & XXX & CMS & XXX & XXX  & 654 & 23\\
-	Automad & XXX & CMS & XXX & XXX  & 655 & 8\\
-	Monstra & XXX & CMS & XXX & XXX  & 1934 & 48\\
-	Nibbleblog & XXX & CMS & XXX & XXX  & 1013 & 28\\
+	Anchor & 0.12.1 & Content Management System & 15,054 & 201 & 987 & 32\\
+	Kirby & 2.4.1 & Content Management System & 51.778 & 435  & 654 & 23\\
+	Automad & 0.10 & Content Management System & 13.978 & 64  & 655 & 8\\
+	Monstra & 3.0.4 & Content Management System & 41.102 & 490  & 1934 & 48\\
+	Nibbleblog & 4.0.5 & Content Management System & 20.376 & 328  & 1013 & 28\\
 	\bottomrule
@@
 
-Given a symbolic value, indirect call mechanisms like callback-commands are
-likely to fail since target function name and the symbolic value do not match.
+Given a symbolic value, indirect function invocation mechanisms like
+callback-commands are likely to fail since target function name and the symbolic value do not match.
 This also applies for the evaluation of include expressions as concrete values
@@
 investigate how many functions fail due to inaccurately resolved indirect
-function calls. Moreover, we evaluate the accuracy of evaluation of include
-expressions in the next section.
+function invocations. Moreover, we evaluate the accuracy of evaluation of
+include expressions in the next section.
 
@@
 access points) having no direct call site as callback candidates since the only
-way to reach them is through indirect mechanisms. Measuring how many
-not-reached output candidates can be explained by functions that are callback
-candidates helps to understand whether, and if so, indirect call mechanisms and
-their usage may impact our analysis’ code coverage.
+way to reach them is through dynamic function invocation mechanisms. Measuring
+how many not-reached output candidates can be explained by functions that are callback
+candidates helps to understand whether, and if so, indirect function invocation
+mechanisms and their usage may impact our analysis’ code coverage.
 
@@
 candidates we measured how many of those functions were only, or
-partially accessibly through indirect calling mechanisms (see Section
-\ref{sec:experiment_dynamicfeatures}). As Figure
+partially accessibly through indirect function invocation mechanisms (see
+Section \ref{sec:experiment_dynamicfeatures}). As Figure
 \ref{fig:output_candidate_explanation} illustrates, around 80 percent (except
@@
 dynamic language features, in particular include expressions and indirect function
-calls, in combination with symbolic values cannot be evaluated accurately since
-concrete values are required. For our observations we were able to explain most
-of the missed output candidates with indirect function calls. For include
+invocations, in combination with symbolic values cannot be evaluated accurately
+since concrete values are required. For our observations we were able to explain most
+of the missed output candidates with indirect function invocations. For include
 expressions we are indeed able to resolve most expressions successfully, yet we
@@
 
-%Tools using symbolic execution for static output approximation \cite{artzi_finding_2008,artzi_finding_2010}
-%\textcolor{blue}{\blindtext}
-
 \section{Conclusion}

### 2017-01-22 paper.tex
abstract

@@
 
-\definecolor{lightgrey}{HTML}{FCFCFF}
+\definecolor{lightgrey}{HTML}{FFFFFF}
 
@@
 \begin{abstract}
-\textcolor{gray}{
-Dynamic web applications have become widely popular and are to
-a large proportion based on the scripting language PHP. Output approximation of web
-applications enables a range of additional tool support as well as possibilities
-for vulnerability detection. Unfortunately, recent approximation approaches have
-only been evaluated for smaller systems.}
-\textcolor{gray}{This paper presents an experience report about the scalability of output
-approximation using symbolic execution of state-of-the-art PHP web
-applications. For a symbolic execution engine extended with support for
-object-oriented programming and arrays, we identified language features and
-corresponding programming patterns that impede symbolic execution and limit the
-scalability of this approach. Our findings include: (1) Dynamic features such as
-functions and includes are prone to fail for certain programming patterns. (2)
-Expressions containing elements from I/O, databases or files can heavily impede
-symbolic execution. Our findings provide useful guidelines to design new
-tools and also to improve the development process of statically analyzable web
-applications.}
+Dynamic web applications have become widely popular and are to a large portion
+based on the scripting language PHP. Dynamic languages like PHP challenge
+static analysis, as they generate code at runtime. One approach to unfold their
+staged nature is symbolic execution by approximating client page output.
+Knowledge about possible output enables a range of developer tools and analyses
+for vulnerability detection.
+
+However, previous tools have only been evaluated for smaller systems and
+scalability for larger systems has not been shown. We criticize the missing
+evalaution of practicality of respective tools and ask whether symbolic
+execution for output approximation for PHP is scalable and what limitations can
+be faced. This paper presents our experiences with re-implementing and
+extending a previous symbolic execution engine for PHP that we evaluate for
+several large PHP systems.
+We find that dynamic language features are likely to be resolved imprecisely
+for symbolic values and impede symbolic execution. We explore
+strategies to mitigate limitations based on previous work on dynamic symbolic
+execution by manually providing concrete rather than symbolic values.
 \end{abstract}
@@
 approximation for \sf{WordPress} as we faced a number of obstacles hindering
-symbolic execution in different ways: The following Subsection
-\ref{sec:experience_report} documents our design choices we made, apparent trade-offs we recognized, and limitations
-we faced on our way towards a scalable symbolic execution engine; and what we
-have learned from bypassing limitations faced in spite of manual effort for
-future work. Subsection \ref{sec:experience_results} compares our observations
-and results to \sf{PhpSync} and presents the motivation for our empirical
-evaluation of our tool symbolic execution engine in Section \ref{sec:evaluation}.
+symbolic execution in different ways.
 
 \subsection{Experience Report}\label{sec:experience_report}
+We document our design choices we made, apparent trade-offs we
+recognized, and limitations we faced on our way towards a scalable symbolic
+execution engine; and what we have learned from bypassing limitations faced in
+spite of manual effort for future work.
 
@@
 
+% Comparison with PhpSync
 \subsection{Experience Results} \label{sec:experience_results}
@@
 	\end{tabular}
-	\caption{Corpus of twelve PHP systems. The file count includes files with a .php,
-	.inc, .bit or .module extension.}
+	\vspace{1mm}
+	\caption{Corpus of twelve PHP systems. The file count includes files with a
+	\tt{.php}, \tt{.inc}, \tt{.bit} or \tt{.module} extension.}
 	\label{corpus}

### 2017-01-23 paper.tex
5,6,7 refactored

@@
 %\titlenote{Produces the permission block, and copyright information}
-\subtitle{Output-oriented Symbolic Execution for Static Output Approximation}
+%\subtitle{Output-oriented Symbolic Execution for Static Output Approximation}
 %\subtitlenote{The full version of the author's guide is available as  \texttt{acmart.pdf} document}
 
-%\author{Stefan Mühlbauer}
+\author{Stefan Mühlbauer}
 \affiliation{%
-  %\institution{Technische Universität Braunschweig}
+  \institution{Technische Universität Braunschweig}
+  %\streetaddress{P.O. Box 1212}
+  %\city{Braunschweig} 
+  \state{Germany} 
+  %\postcode{43017-6221}
+}
+%\email{s.muehlbauer@tu-bs.de}
+
+\author{Christian Kästner}
+%\orcid{1234-5678-9012}
+\affiliation{%
+  \institution{Carnegie Mellon University}
+  %\streetaddress{5000 Forbes Avenue}
+  %\city{Pittsburgh} 
+  %\state{PA} 
+  %\postcode{15217}
+}
+%\email{s.muehlbauer@tu-bs.de}
+
+\author{Tien N. Nguyen}
+%\orcid{1234-5678-9012}
+\affiliation{%
+  \institution{University of Texas at Dallas}
   %\streetaddress{P.O. Box 1212}
@@
 
-%\author{Christian Kästner}
-%\orcid{1234-5678-9012}
-\affiliation{%
-  %\institution{Carnegie Mellon University}
-  %\streetaddress{5000 Forbes Avenue}
-  %\city{Pittsburgh} 
-  %\state{PA} 
-  %\postcode{15217}
-}
-%\email{s.muehlbauer@tu-bs.de}
-
-%\author{Tien N. Nguyen}
-%\orcid{1234-5678-9012}
-\affiliation{%
-  %\institution{University of Texas at Dallas}
-  %\streetaddress{P.O. Box 1212}
-  %\city{Braunschweig} 
-  %\state{Germany} 
-  %\postcode{43017-6221}
-}
-%\email{s.muehlbauer@tu-bs.de}
-
 % The default list of authors is too long for headers}
@@
 based on the scripting language PHP. Dynamic languages like PHP challenge
-static analysis, as they generate code at runtime. One approach to unfold their
+static analysis, as they generate code at run-time. One approach to unfold their
 staged nature is symbolic execution by approximating client page output.
 Knowledge about possible output enables a range of developer tools and analyses
-for vulnerability detection.
-
-However, previous tools have only been evaluated for smaller systems and
-scalability for larger systems has not been shown. We criticize the missing
-evalaution of practicality of respective tools and ask whether symbolic
-execution for output approximation for PHP is scalable and what limitations can
-be faced. This paper presents our experiences with re-implementing and
-extending a previous symbolic execution engine for PHP that we evaluate for
-several large PHP systems.
-We find that dynamic language features are likely to be resolved imprecisely
-for symbolic values and impede symbolic execution. We explore
-strategies to mitigate limitations based on previous work on dynamic symbolic
-execution by manually providing concrete rather than symbolic values.
+for vulnerability detection. However, previous tools have only been evaluated
+for smaller systems and scalability for larger systems has not been shown.
+
+In this paper, we criticize the limited expressiveness of the evaluation of
+practicality for respective tools. We ask whether symbolic execution for output
+approximation for PHP is scalable, and what limitations can be faced. By
+extending a symbolic execution engine based on previous work and evaluating it
+for some larger PHP systems we find dynamic features to be likely to hinder
+symbolic execution. We explore strategies to mitigate limitations inspired by
+work on dynamic symbolic execution and aim to push for more comprehensive
+evaluation for developer tools.
 \end{abstract}
@@
 such as client page output. The latter code, although assembled at
-runtime, may contain client-side parts of the web application, such as
+run-time, may contain client-side parts of the web application, such as
 JavaScript. So, to study dynamic web applications in its entirety, we need to
@@
 To investigate the question, whether we can have a practical and scalable
-symbolic executoin engine, we re-implemented the engine with the specifications
+symbolic execution engine, we re-implemented the engine with the specifications
 of the previous symbolic execution semantics \cite{Nguyen:2014:BCG:2635868.2635928} for PHP and additional features, such
@@
 include expression require concrete information to be evaluated properly. Since
-these expressions may be assembled at runtime and can contain symbolic
+these expressions may be assembled at run-time and can contain symbolic
 information, for many cases this restricts a symbolic execution engine from
@@
     \caption{
-    	Illustration of static ouptut approximation using symbolic
+    	Illustration of static output approximation using symbolic
     	execution
@@
 \texttt{make\_titles} defined in lines 7 – 13 prints the string \texttt{\$arg} passed as an argument between zero and an arbitrary number of
-times. Second, the function is indirectly called using the built-in mechnism
+times. Second, the function is indirectly called using the built-in mechanism
 \texttt{call\_user\_func} in line 15 with the string \texttt{Headline} and
@@
 For the greeting there are two possible client page output
-variants: If the input field \texttt{name} is empty at runtime, line 19 is
+variants: If the input field \texttt{name} is empty at run-time, line 19 is
 executed and \texttt{No name entered!} is printed, or, if a name was entered,
@@
 	provides editor services across stages for client-side code: Starting with an
-	ouput approximation a callgraph with conditional edges is constructed, which
+	output approximation a call-graph with conditional edges is constructed, which
 	allows navigation in client-site code (HTML, JS and CSS) although it is
@@
 case for while and do-while loops, the number of loop iterations for for or
-foreach loops is generally determinable and known before entering the loop
+for-each loops is generally determinable and known before entering the loop
 body. For loops use a loop counter that usually is incremented or decremented
 after each loop iteration until a loop condition is not satisfied anymore;
-foreach loops simply iterate over a given array value.
+for-each loops simply iterate over a given array value.
 Our execution semantics follows this scheme and simulates concrete execution
 for those types of loops if possible. Otherwise, for instance, if a loop
-counter, loop counter limit for for loops, or variable for foreach loops is
+counter, loop counter limit for for loops, or variable for for-each loops is
 symbolic, we resort to executing only a single iteration. This design choice
@@
 that we were not able to execute accurately. Dynamic features in PHP are
-features that take as input rather a expression that is resolved at runtime
+features that take as input rather a expression that is resolved at run-time
 than a static constant value. For \sf{WordPress}, we faced missing files and
@@
 statements takes as input either a string constant (static include) or an
-expression that is evaluated at runtime (dynamic include). The string that the
+expression that is evaluated at run-time (dynamic include). The string that the
 expression evaluates to represents an absolute or relative paths to the target
@@
 configuration or even just default values or placeholders. Hence, symbolic
-execuion is prone to fail for dynamic includes if symbolic information is contained.
+execution is prone to fail for dynamic includes if symbolic information is contained.
 
@@
 \tt{sserif}, a different function is called. Given that the function name is 
-concrete at runtime, the desired function can be called. Otherwise, if the
-function name is assembled at runtime and contains symbolic values, no function
+concrete at run-time, the desired function can be called. Otherwise, if the
+function name is assembled at run-time and contains symbolic values, no function
 can be called as the name is ambiguous or unknown. For this example, symbolic
@@
 present our methods to evaluate practicality of our symbolic output
-approximation. Moreover, we investigate to which extent the conceptual limitations encountered
-with \sf{WordPress} apply for a wider choice of PHP systems.
+approximation. Moreover, we investigate to which extent the limitations encountered
+with \sf{WordPress} apply for a wider choice of larger PHP systems.
 
 \subsection{Experiment Setup} \label{sec:experiment_setup}
-We have encountered several issues for \sf{WordPress} that conceptually limit
-the scalability of symbolic execution. Since our observations so far only considers
+We have encountered several problems for \sf{WordPress} that conceptually limit
+the scalability of symbolic execution. Since our observations so far only consider
 one example system an, we evaluate our symbolic execution engine for a wider
@@
 re-implements \sf{PhpSync} and extends with support for additional language
-features, we ask: Can  we replicate high coverage results at least on a
-small-scale. Therefore, our PHP corpus contains four small-scale systems used
-in previous work \cite{Nguyen:2014:BCG:2635868.2635928}.
-
-Second, we ask: Do system size and usage of modern language features
-factors affect the practicality of our symbolic execution engine? We
+features, we ask: \textit{Can  we replicate high coverage results at least on a
+small-scale?} Therefore, our PHP corpus contains four of the small-scale systems
+that \sf{PhpSync} has been evaluated for \cite{Nguyen:2014:BCG:2635868.2635928}.
+
+Second, we ask: \textit{Do system size and usage of modern language features
+factors affect the practicality of our symbolic execution engine?} We
 selected both a range of large-scale and small-scale modern PHP systems to
-better separate and understand the circumstances under which conceptual
-limitations might affect practicality. For the selection of large-scale systems
-we borrow three systems from a case study addressing the feature usage in PHP
-systems. Some more detailed description of this case study can be found in
-Section \ref{sec:related_work}. For the small-scale systems we selected five systems from
-a list of recent Content Management Systems (CMS) written in PHP \cite{codegeekz}. For both
-selections we were bound to select a smaller number of system than we intended
+better separate and understand the circumstances under which limitations might affect practicality. For the selection of large-scale systems
+we borrow three systems from a case study by Hills et al. \cite{hills2015variable} addressing the
+feature usage patterns and frequencies in PHP systems. Some more detailed description of this
+case study can be found in Section \ref{sec:related_work}. For our choice of modern small-scale systems we
+selected five systems from a list of recent Content Management Systems (CMS) written in PHP \cite{codegeekz}. For both
+selections we were bound to select a smaller number of systems than we intended
 to since the parser used in our implementation,
@@
 in and  scattered across server-side code) that may eventually become part of any
-client page output variant (in the following referred to as \emph{output
-candidate}) is contained in our symbolic output approximation.
-It is non-trivial to determine whether a string literal is an output candidate,
-so we heuristically classify string literals as output candidates if they
-contain the characters \tt{<}, \tt{>}, or both since we expect output to contain
-HTML tags.
+client page output variant is contained in our symbolic output approximation. In
+the following we refer to those expected output string literals as \emph{output candidates}.
+
+With respect to the staged nature of web applications it is
+non-trivialstatically to determine whether a string literal is an output
+candidate, so we heuristically classify string literals as output candidates if
+they contain the characters \tt{<}, \tt{>}, or both since we expect output to
+contain HTML tags.
 
@@
 classified incorrectly as false positives. In turn, the classifier is highly
-distinctive as 96 percent of the string literals are classified correctly. The
+distinctive as 94 percent of the output candidates are classified as such. The
 recall of 50 percent means that half of the string literals, which we manually 
-classified as output candidates, actually were responsive to the classifier. We
+classified as output candidates, actually are responsive to the classifier. We
 attempted to increase recall by looking for further distinctive properties to build a
-classifier from, but could not do so without decreasing precision. 
+classifier from, but were not able to do so without decreasing precision. 
 In spite of missing half of the output candidates, we decided
@@
 In Section \ref{heuristic} we introduced the definition of output candidates as
-expected output. For an approximation to be most accurate, it needs to contain
+expected output. For an approximation to be accurate, it needs to contain
 all output candidates of a system analyzed. Based on whether output
 candidates are reached by an execution path and part of the output
-approximatiopn define two metrics to approximate measuring accuracy.
+approximation we define two metrics to approximate measuring accuracy.
 
@@
 The coverage results for our case study are illustrated in Figure
-\ref{coverage}. We could replicate high coverage results for the four small-scale systems with a reach
-coverage and output coverage over 80 percent respectively. For the three
-modern/large-scale systems we measured poor coverage with reach coverage
+\ref{coverage}. For our first evaluation question in Section \ref{sec:experiment_setup} we
+could replicate high coverage results for the four small-scale systems with a reach
+coverage and output coverage over 80 percent respectively. Whereas for the
+three modern/large-scale systems we measured poor coverage with reach coverage
 ranging from 5 to 30 percent and output coverage ranging from 4 to 20 percent.
@@
 coverage metrics were poor for large-scale and small-scale/recent systems. To
-understand what output candidates we missed and why, we further investigated
-our approximation results and conducted three more measurements.
+understand what output candidates we missed and why, we further investigate
+our approximation results and conduct three subsequent measurements.
 
@@
 	\caption{Evaluation results for output candidates, include expressions and
-	callgraph analysis for not-reached output candidates}
+	call-graph analysis for not-reached output candidates}
 \end{figure}
@@
 Our initial approach to understand missed output candidates is to find out
-whether their surrounding program code was not accessed/accessible, and if so,
-why. We are interested in those output candidates that are located in HTML
+whether their surrounding program code cannot not accessed, and if
+so, why. We are interested in those output candidates that are located in HTML
 files (possibly with nested PHP), or located in a function definition. Plain
@@
 candidates that are part of a function definition are only missed if the
-corresponding function is never called at runtime.
+corresponding function is never invocated at run-time, or simply dead code.
 
@@
 for almost all systems tested (except for \textsf{AddressBook} and
-\textsf{TimeClock}) not-reached output candidates had a function context.
+\textsf{TimeClock}) not-reached output candidates have a function context.
 
@@
 
-For a function to be never reached there are several scenarios: A function is
-undefined at runtime if the corresponding script file is never included. In
-turn, if the function is defined at runtime, the function is dead
+For a function to never be reached there are several scenarios: A function is
+undefined at run-time if the corresponding script file is never included. In
+turn, if the function is defined at run-time, the function is dead
 code if there is no call site for this function, or the function is only called
-indirectly and the function name could not be resolved accurately. In PHP there
-are several ways to call a function beside direct call sites. 
+dynamically and the function name could not be resolved precisely. In PHP there
+are several ways to call a function beside static call sites, as described in
+Section \ref{sec:experiment_dynamicfeatures}.
 %The language offers indirect call mechanisms like callback-commands or an
@@
 Given a symbolic value, indirect function invocation mechanisms like
-callback-commands are likely to fail since target function name and the symbolic value do not match.
+callback-commands are likely to fail since target function name and the symbolic
+value the expression evaluates to do not match.
 This also applies for the evaluation of include expressions as concrete values
@@
 symbolic values is ambiguous. Since it is unlikely that all functions containing
-missed output candidates are dead code having no call site (direct or
-indirect) at all, in Section \ref{WhyDidFunctionCallsFail} we further
-investigate how many functions fail due to inaccurately resolved indirect
-function invocations. Moreover, we evaluate the accuracy of evaluation of
-include expressions in the next section.
+missed output candidates are dead code having no call site (static or dynamic)
+at all, in Section \ref{WhyDidFunctionCallsFail} we further investigate how
+many functions fail due to inaccurately resolved indirect function invocations.
+Moreover, we evaluate the accuracy of evaluation of include expressions in the next section.
 
@@
 From Section \ref{sec:experiment_dynamicfeatures} we learn that include
-expressions can not be evaluated accurately for symbolic values, and from the
-previous Section \ref{WhatLiteralsDidWeMiss?} that very little of missed output
-candidates can be explained by simply missing to include a file. To better
-understand the impact of include expression resolution throughout analyzing
-systems, we construct two additional metrics. First, we are interested in how
-many include expressions throughout the execution we actually reach in a system.
-Therefore we introduce the metric \emph{reach coverage for include expressions}
-as the ratio of reached include expressions and the total number of reach
-expressions in a system to check whether failed includes may have caused
-subsequent include expressions and so on. Second, we measure how successful
-include expressions are resolved.
+expressions containic symbolic valuescannot be resolution precisely, and from
+the previous Section \ref{WhatLiteralsDidWeMiss?} that very little of missed
+output candidates can be explained by simply missing to include a file. To
+better understand the impact of include expression resolution throughout
+analyzing systems, we construct two additional metrics.
+
+First, we are interested in how many include expressions throughout the
+execution we actually reach in a system. Therefore we measure the number of
+reached include expressions in a system to check whether failed includes may
+have caused subsequent include expressions and so on.
+
+Second, we measure how many reached include expressions are
+succesfully resolved throughout the analysis for each system.
 
@@
 Figure \ref{fig:include_coverage_results} indicates that except for the
-small-scale systems as well as \sf{Anchor} and \sf{Nibbleblog} we did not really
+small-scale systems as well as \sf{Anchor} and \sf{Nibbleblog} we did not
 reach a great portion of include expressions, although for most systems the
-resolution success rate was over 80 percent. Note that one failed attempt
-to include a script file can result in a cascade of missing more include
-expressions as the include target script file is not included and executed.
+resolution success rate (ratio of resolved and reached include expressions) was
+over 80 percent. Note that one failed attempt to include a script file can 
+result in a cascade of missing more include expressions as the include target
+script file is not included and executed. This may explain the low number of
+reached include expressions, but relatively high success rate.
 
@@
 Aside from failed or successful includes, this metric describes one scenario of
-why functions containing not-reached output candidates are never called: The
-trivial explanation would be that there are no direct calls of those functions,
-yet PHP offers a number of ways to call a function indirectly, for instance
-using built-in functions, or simply evaluating strings as PHP code using the
-\texttt{eval} function.
+why functions containing not-reached output candidates are never invocated: The
+trivial explanation would be that there are no static invocations of those
+functions, yet PHP offers a number of ways to invocate a function dynamically,
+for instance using built-in functions, or simply evaluating strings as PHP code
+at run-time using the \texttt{eval} function.
 
 Moreover we need to take into account that one missed function execution can
-result in missing even more function calls (direct or indirect) and executions.
+result in missing even more function invocations.
 Thus, for functions containing not-reached output candidates we measure whether
 they can be (1) only, (2) partially or (3) never accessed transitively from
-functions that have no direct call sites. We refer to those functions (or
-access points) having no direct call site as callback candidates since the only
-way to reach them is through dynamic function invocation mechanisms. Measuring
-how many not-reached output candidates can be explained by functions that are callback
+functions that have no direct call sites. We refer to those functions having no
+static call site as \emph{callback candidates} since the only way to reach them
+is through dynamic function invocation mechanisms. Measuring how many not-reached output candidates can be explained by functions that are callback
 candidates helps to understand whether, and if so, indirect function invocation
@@
 
-The callgraph snippets in Figure \ref{fig:callgraph} illustrate the idea of
+The call-graph snippets in Figure \ref{fig:callgraph} illustrate the idea of
 detecting the functions’ access points: The nodes at the top depict callback
@@
     \input{new_diagrams/callback_classification}
-    \caption{Callgraph analysis: The bottom nodes represent functions
+    \caption{Call-graph analysis: The bottom nodes represent functions
     containing missed output candidates, the top nodes depict entry points to
@@
 
-For failed function calls to functions containing not-reached output
+For failed function invocations to functions containing not-reached output
 candidates we measured how many of those functions were only, or
@@
 
+Hence, we believe the impact of imprecisely resolved function names for dynamic
+function invocations, aside from dead code, has a limiting impact on symbolic
+execution since callable code sections become inaccessible.
+
 \section{Lessons Learned}
@@
 From out observations with \sf{WordPress} in Section \ref{sec:3} we learn that
-dynamic language features, in particular include expressions and indirect function
-invocations, in combination with symbolic values cannot be evaluated accurately
-since concrete values are required. For our observations we were able to explain most
-of the missed output candidates with indirect function invocations. For include
-expressions we are indeed able to resolve most expressions successfully, yet we
-do not distinguish between static and dynamic includes in this case study as
-this would exceed the scope of our paper. However, we believe it is plausible
-to assume that defective include expressions can cause a cascade of missing
-files and subsequent includes.
-Throughout the experiment we have seen that dynamic features, are prevalent for
-modern PHP systems as they serve a genuine purpose in the programming language.
-Nevertheless their usage in situations when combined with assumed symbolic
+dynamic language features, in particular include expressions and dynamic
+function invocations containing symbolic values, often cannot be executed
+accurately since include targets or function names respectively are resolved
+imprecisely.
+For our observations we were able to explain most of the missed
+output candidates with indirect function invocations. For include expressions we
+are indeed able to resolve most reached expressions successfully, yet we do not
+distinguish between static and dynamic includes in this case study as this 
+would exceed the scope of our paper. However, we believe it is plausible to
+assume that imprecisely resolved include expressions can cause a cascade of
+missing files and subsequent includes.
+Throughout the evaluation we have seen that dynamic features, are commonly used
+in modern PHP systems as they serve a genuine purpose in the programming
+language.
+Nevertheless their usage in situations where they contain assumed symbolic
 values restricts symbolic execution and any tool leveraging results obtained
@@
 Symbolic execution is used to approximate client page output since it allows to
-unfold the staged nature of dynamic web applications. Aside from conceptual
-limitations in the systems studied, limitations can be addressed. To use
+unfold the staged nature of dynamic web applications.
+
+Aside from being limited by dynamic features, we exercised replacing dynamic
+features with respective static counterparts in order to bypass limitations by
+manually providing expressions that can be resolved precisely.To use
 approaches incorporating symbolic execution with dynamic web applications in
@@
 to be provided to decrease the number of assumed symbolic inputs and avoid
-dynamic features being confronted with symbolic values. Also, if the usage of
-tools using static output approximation with symbolic execution is intended,
-avoiding dynamic features that are prone to be inaccurately executed for
+dynamic features being confronted with symbolic values. Exising work approaching
+this goal \cite{artzi_finding_2008,artzi_finding_2010,DynamicWassermann}
+incorporates dynamic symbolic execution, i.e. concrete execution alongside symbolic execution.
+
+Also, if the usage of tools using static output approximation with symbolic
+execution is intended, avoiding dynamic features that are prone to be inaccurately executed for
 symbolic execution  may be an adjustment strategy to program in an analyzable
@@
 \paragraph{Practical tools require practical evaluation.}
-Last, we want to address the limited informative value of analyses and tools
+We wanted to address the limited informative value of analyses and tools
 presented in previous work
@@
 possible drawbacks of object-oriented symbolic execution. 
-This question though is beyond the scope of this paper, as we aimed
-to explore limitations of the practicality and scalability of static output
-approximation using symbolic execution. Future work might address design
-guidelines to enable programming analyzable applications, tool support to scale
+This question though is beyond the scope of this paper.
+Future work might address design guidelines to enable programming analyzable applications, tool support to scale
 the number of assumed symbolic inputs or systematically provide concrete data
@@
 approaches incorporating constraint solving and leveraging concrete execution
-anlongside symbolic execution emerged and enable automated test case generation
-and testing; the scalability of these approaches is still challenged by path
-explosion and constraint solving \cite{CadarSen2013}.
+alongside symbolic execution emerged and enable automated test case generation;
+the scalability of these approaches is still challenged by path explosion and constraint solving \cite{CadarSen2013}.
 
 Dynamic symbolic execution \cite{CadarSen2013} incorporates symbolic execution
-alongside concrete execution and ahs shown promising results in automated test
-generation and testing \cite{artzi_finding_2008,artzi_finding_2010,DynamicWassermann} for PHP web
+alongside concrete execution and has shown promising results in automated test
+generation\cite{artzi_finding_2008,artzi_finding_2010,DynamicWassermann} for PHP web
 applications and might help mitigate the conceptual limitations faced with
@@
 dynamic webpage is executed with concrete input and the execution is
-recorded at runtime. Changes in the client-side output then can be mapped to 
+recorded at run-time. Changes in the client-side output then can be mapped to 
 corresponding PHP code using static impact analysis.
@@
 of the feature usage in PHP, based on a corpus of state-of-the-art-projects,
-dynamic includes are less frequently used than static includes, but usage
-frequency is varying from system to system. Further work by Hills et al.
+dynamic includes are less frequently used than static includes, yet usage
+frequency varies from system to system. Further work by Hills et al.
 \cite{hills2014static,hills2014php} approaches static resolution of dynamic includes using
-both context-insensitive resolution on file level by simplifying PHP constants; and context-sensitive on program level for transitive includes. In spite of
+both context-insensitive resolution on file level by simplifying PHP constants;
+and context-sensitive on program level for transitive includes. In spite of 
 promising results, these approaches, similar to our results, face limitations
@@
 In this paper we have addressed symbolic execution of PHP web applications that
-is used for client page output approximation. This approximations are leveraged
-be several developer tools for PHP, yet have not been evaluated for larger
+is used for client page output approximation. These approximations are leveraged
+be several developer tools for PHP that yet have not been evaluated for larger
 systems. We re-implemented and extended an existing symbolic execution engine
@@
 necessity of more comprehensive evaluation of developer tools as
-previous work did not explore these limitations. Dynamic features of PHP in
-combination with symbolic values though hinder precise symbolic execution and
-limit the scalability and practicality of tools on symbolically approximated
-client page output.
-
-In the future we would like to use dynamic symbolic execution for output
-approxiamtion and possibly bypass existing limitations, as dynamic symbolic
-execution has shown to be promising for automated testing and verification.
+symbolic execution is a conceptually suitable, but not neccessary practical
+approach to approximate client page output. Dynamic features of PHP in
+combination with symbolic values though can hinder precise symbolic execution
+and limit the scalability and practicality of tools relying on symbolically
+approximated output.
+
+In the future we would like to apply dynamic symbolic execution for output
+approximation and possibly bypass existing limitations, as dynamic symbolic
+execution has shown promising results for automated testing and verification.
 

### 2017-01-23 paper.tex
1 refactored

@@
 
-\author{Stefan Mühlbauer}
+%\author{Stefan Mühlbauer}
 \affiliation{%
-  \institution{Technische Universität Braunschweig}
+  %\institution{Technische Universität Braunschweig}
   %\streetaddress{P.O. Box 1212}
   %\city{Braunschweig} 
-  \state{Germany} 
+  %\state{Germany} 
   %\postcode{43017-6221}
@@
 Dynamic web applications have become widely popular and are to a large portion
-based on the scripting language PHP. Dynamic languages like PHP challenge
-static analysis, as they generate code at run-time. One approach to unfold their
-staged nature is symbolic execution by approximating client page output.
-Knowledge about possible output enables a range of developer tools and analyses
-for vulnerability detection. However, previous tools have only been evaluated
-for smaller systems and scalability for larger systems has not been shown.
+implemented in the scripting language PHP. Dynamic languages like PHP though
+challenge static analysis, as they generate code at run-time. One approach to unfold their
+staged nature and to approximate client page output is symbolic execution.
+Knowledge about possible output variants enables a range of developer tools as
+well as analyses for vulnerability detection. However, previous tools have only
+been evaluated for smaller systems and scalability for larger systems has not
+been assessed.
 
@@
 widely popular. Various different implementation techniques have emerged,
-ranging from technologies for programming languages (JSP, ASP .NET) and web
+ranging from web technologies for programming languages (JSP, ASP .NET) and web
 application frameworks for  script languages (Ruby On Rails, Django) to
@@
 \cite{phpNET} is a programming language focused on server-side application
-development. As of 2012, it was used by 78.8 percent of the ten million most popular websites (according
-to Alexa popularity ranking) \cite{alexaPHP}, ranking 7th on the TIOBE
+development. As of 2012, PHP was used by 78.8 percent of the ten million most
+popular websites (according to Alexa popularity ranking) \cite{alexaPHP}, ranking 7th on the TIOBE
 programming community index \cite{tiobePHP}, and was the ranked as the 6th
@@
 run-time, may contain client-side parts of the web application, such as
-JavaScript. So, to study dynamic web applications in its entirety, we need to
-consider both static as well as dynamic aspects of systems.
+JavaScript. Therefore, to study dynamic web applications in its entirety, we
+need to consider both static as well as dynamic aspects of those systems.
 
@@
 and program proving. For a program, symbolic values are
-used instead of concrete inputs \cite{Darringer1978,King1976}. The underlying
+assumed instead of concrete inputs \cite{Darringer1978,King1976}. The underlying
 concept is to map program input to program output: Symbolic values are used and
 propagated throughout the symbolic execution and keep dependencies for program
-output traceable. This analysis technique helps unfold the \emph{staged} nature
-of dynamic web applications. Since every feasible path can be executed the
-symbolic output contains both invariant output as well as output variants that
-depend on program input.
+output traceable. This analysis technique helps unfold the staged nature
+of dynamic web applications. Since every feasible path can be explored
+and executed the symbolic output contains both invariant output as well as
+output variants that depend on program input.
 
 Knowledge about different output variants is leveraged by a number of
-analyses for the domain of web applications. Previous work presented tools
-for detecting and locating HTML validation errors
+tools and analyses for the domain of web applications. Previous work presented
+tools for detecting and locating HTML validation errors
 \cite{Nguyen:2011:AFH:2190078.2190142}, computing program slices across
@@
 with code navigation \cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
-Moreover, knowledge of output variants can increases code coverage for
+Moreover, knowledge of output variants can increase code coverage for
 output-oriented testing \cite{Alshahwan2011}.
@@
 execution has so far only been evaluated for small systems that are not
-maintained anymore. The tools presented by previous work
+actively maintained. The tools presented by previous work
 \cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2011:AFH:2190078.2190142}
 though are only practical, if for a given system the symbolic execution engine
-is scalable, i.e., will also approximate output accurately for larger and more
-recent systems with acceptable time and space consumption.
+is scalable, i.e., will also approximate client page output accurately for
+larger and more recent systems with acceptable time and space consumption.
 
 To investigate the question, whether we can have a practical and scalable
-symbolic execution engine, we re-implemented the engine with the specifications
-of the previous symbolic execution semantics \cite{Nguyen:2014:BCG:2635868.2635928} for PHP and additional features, such
-as object-oriented programming. We evaluate our
-symbolic execution engine for several large-scale and modern PHP systems.
-
-During the introduction of new semantics, we addressed two  trade-offs between
-accuracy of our execution and performance: For method calls with ambiguous
-targets as well as concrete execution of loops the number of program states can
-become infeasible. This may require additional engineering adjustments to tame
-the state space explosion at cost of accuracy.
-
-Based on observations we identify conceptual limitations of symbolic
-execution for PHP.  Dynamic features such as indirect function invocations and
-include expression require concrete information to be evaluated properly. Since
-these expressions may be assembled at run-time and can contain symbolic
-information, for many cases this restricts a symbolic execution engine from
-further execution.
-Although some effort to statically approximate include expression can be spent,
-it is often non-trivial since expressions can contain information coming from
-various inputs including databases, user inputs or configuration files. We
-empirically evaluate our observations for a corpus including large, actively maintained PHP systems.
+symbolic execution engine, we re-implemented the symbolic engine with the
+specifications of the previous symbolic execution semantics
+for \sf{PhpSync} \cite{Nguyen:2014:BCG:2635868.2635928} for PHP and
+extended it with additional features, such as object-oriented programming. We
+evaluate our symbolic execution engine for several large-scale and modern PHP systems.
+
+During the introduction of new semantics for object-oriented programming we
+faced a trade-off between accuracy and scalability: Whilst exploring multiple
+paths, for method invocations the number of target objects may grow rapidly and
+infeasible.
+
+Moreover, based on observations we identify limitations of symbolic execution
+for PHP. Dynamic features such as dynamic function invocations and dynamic
+include for an accurate execution require precisely resolved include targets
+and function names respectively. Since these expressions  can contain symbolic
+values at run-time, for many cases this restricts a symbolic execution engine
+from further execution as callable code sections become inaccessible.
+Although some effort can be spent to statically resolve dynamic include
+expressions \cite{hills2014static}, or to simply provide static counterparts manually, resolution
+is often not trivial since expressions can contain symbolic values assumed instead
+of concrete program input, including databases, user inputs or configuration
+files. We empirically evaluate our symbolic execution engine for a corpus
+including large and actively maintained PHP systems.
 
@@
 symbolic execution and (2) a report of observations for state-of-the-art
-web applications as well as an empirical evaluation of conceptual limitations
+web applications as well as an empirical evaluation of practical limitations
 of symbolic executions for PHP web applications.

### 2017-01-23 paper.tex
2 refactored

@@
 paths, for method invocations the number of target objects may grow rapidly and
-infeasible.
-
-Moreover, based on observations we identify limitations of symbolic execution
+infeasible. Moreover, based on observations we identify limitations of symbolic execution
 for PHP. Dynamic features such as dynamic function invocations and dynamic
@@
 \section{State Of the Art}\label{sec:stateoftheart}
-This section recaps the idea of symbolic
-execution, the tools which make use of it, and both conceptual and
-language-related limitations.
+This Section recaps the idea of symbolic
+execution, the tools using output approximation based on it and whose evaluation
+we criticize, and current conceptual and language-related limitations.
 
@@
 Symbolic execution is a static analysis technique that was proposed by King
-\cite{King1976} in 1976. Symbolic execution allows to explore feasible paths in a program. Based on normal program execution, the execution semantics
-is extended to handle symbolic values. An execution becomes symbolic by
-assuming symbolic values as program inputs rather than obtaining concrete
-values \cite{King1976,Darringer1978}. These symbolic values are then propagated
-and used throughout the execution.
-
-Execution starts with a tautology path
-condition. The control-flow can be split whenever the decision between branches
-is ambiguous. This is the case, when conditional expressions, for example for
-if-statements, evaluate to a symbolic value rather than an actual result.
-The corresponding path conditions (guards for instance) are conjoined with the
-previous path condition \cite{King1976}. Additionally, the conjoined path
-condition can be checked for satisfiability in order to exclude infeasible
-paths.
+\cite{King1976} in 1976. Symbolic execution allows to explore feasible paths in
+a program. Based on normal program execution, the execution semantics is
+extended to handle symbolic values. An execution becomes symbolic by assuming
+symbolic values as program inputs rather than obtaining concrete values
+\cite{King1976,Darringer1978}. These symbolic values are then propagated and
+used throughout the execution.
+
+Symbolic execution commences with a tautology path condition. The control-flow
+can be split whenever the decision between branches is ambiguous. This is the
+case, when conditional expressions, for example for if-statements, evaluate to
+a symbolic value rather than an actual result. The corresponding branch
+predicates are conjoined with the previous path condition \cite{King1976}.
+Additionally, the conjoined path condition can be checked for satisfiability in
+order to exclude infeasible paths.
 
@@
 (input form in lines 1 -- 4) as well as PHP code: First, the function
-\texttt{make\_titles} defined in lines 7 – 13 prints the string \texttt{\$arg} passed as an argument between zero and an arbitrary number of
-times. Second, the function is indirectly called using the built-in mechanism
-\texttt{call\_user\_func} in line 15 with the string \texttt{Headline} and
-\texttt{4} as arguments.  Finally, the input from the form in lines 1 – 4 is
-used: The variable \texttt{\$greeting} consists of three concatenated strings,
+\texttt{make\_titles} defined in lines 7 – 13 prints the string \texttt{\$arg} 
+passed as an argument between zero and an arbitrary number of times. Second, the
+function is dynamically invocated using the built-in mechanism
+\texttt{call\_user\_func} in line 15 with the string \texttt{Headline} and \texttt{4} as arguments.  
+Finally, the input from the form in lines 1 – 4 is used: The variable \texttt{\$greeting} consists of three concatenated strings,
 \texttt{Hello}, the value of input field \texttt{name}, and an exclamation mark.
@@
 variants: If the input field \texttt{name} is empty at run-time, line 19 is
-executed and \texttt{No name entered!} is printed, or, if a name was entered,
-line 21 is executed and the greeting constructed in line 17 is printed.
+executed and \texttt{No name entered!} is printed. Otherwise, if a non-empty
+name was entered, line 21 is executed and the greeting constructed in line 17 is printed.
 
@@
 previous work \cite{Nguyen:2014:BCG:2635868.2635928} illustrate the necessary
-modifications and extensions made to handle the level of ambiguity that comes with handling symbolic values. In addition to normal output, the
-output approximation contains preprocessor directives to express
+modifications and extensions made to handle the level of ambiguity that comes with handling symbolic values. 
+In addition to normal output, the output approximation illustrated in
+Figure \ref{fig:output_example} contains preprocessor directives to express
 repeatable parts as well as output variants. All approximated client page
-output corresponding to Figure \ref{fig:web_application} consisting of two different variants
-is illustrated in Figure \ref{fig:output_example}.
+output corresponding to Figure \ref{fig:web_application} consisting of two
+different variants is illustrated in Figure \ref{fig:output_example}.
 
@@
 values, it might not be feasible to determine whether, and if so, when a loop
-execution terminates. Hence, any output constructed during the single loop
+execution terminates. Hence, any output assembled during the single loop
 iteration is highlighted (\texttt{repeat \ldots endrepeat}) to possibly be
@@
 Second, a similar procedure applies to recursive
-functions. Once called, a recursive function is  also only executed to recursion
-depth one per call, and any subsequent invocation of that function returns a
-symbolic value. This, again, is due to the difficulty to determine the exact
+functions. Once invocated, a recursive function is also only executed to
+recursion depth one per call, and any subsequent invocation of that function returns a
+symbolic value. Again, this is due to the difficulty to determine the exact
 recursion depth similar to the number of loop iterations.
 
-Third, any form of user interaction providing input data as well as any form of input referred
-to from the deployment context of the web application, i.e. configuration files
-or databases, is assumed to be symbolic. The context variable
+Third, any form of user interaction providing concrete input data as well as any
+form of input referred to from the deployment context of the web application, i.e. configuration files
+or databases, is assumed to be symbolic. The PHP context variable
 \texttt{\$\_POST} represents input submitted to the server and, hence, is
-symbolic as every program input is per definition substituted by symbolic values. Finally, the
-output in lines 11 and 13 respectively are marked as alternative variants
+symbolic as every program input is per definition substituted by assumed
+symbolic values. Finally, the output in lines 11 and 13 respectively are marked as alternative variants
 (\texttt{if \ldots else \ldots endif}) as symbolic execution of lines 18 – 22
@@
 \subsection{Symbolic-Execution-based Tools For PHP}\label{sec:existing_tools}
-All tools proposed so far leverage a representation of all possible HTML client
-page output. Since any single client page output variant can be analyzed this
-enables tool support addressing the web application as a whole in spite of its
-staged nature, where parts of the web application are
+All tools proposed so far and presented in the following leverage symbolically
+approximated client page output. Since any single client page output variant can
+be analyzed this enables tool support addressing the web application as a whole
+in spite of its staged nature, where parts of the web application are
 generated dynamically.
-Based on the approximated output representation, subsequent analyses can be
-conducted:
+Based on the approximated output representation, subsequent developer tools and
+analyses have been proposed:
 \begin{itemize}
@@
 	conforms to syntactical specifications for HTML and other client-side
-	languages. After the tool detects validation errors, auto-fixes
-	can be provided. Otherwise, it traces validation errors back to
+	languages. After the tool detects validation errors, if possible, auto-fixes
+	are provided. Otherwise, it traces validation errors back to
 	source code responsible for the defect, which then can be refactored manually
@@
 	application for program slices, client page output needs to be taken into
-	account. WebSlice combines PHP data-flow information with an output
+	account. \sf{WebSlice} combines PHP data-flow information with an output
 	approximation to enable program slices across different languages
@@
 	\item \emph{Varis}: Editor services such as “jump to declaration” are
-	nontrivial for dynamic web applications due to their staged nature. Varis
+	not trivial for dynamic web applications due to their staged nature. \sf{Varis}
 	provides editor services across stages for client-side code: Starting with an
@@
 engine described in previous work \cite{Nguyen:2014:BCG:2635868.2635928} does
-not support advanced language features, including object-oriented programming. Moreover, the range of
-standard library functions supported are mostly string operations, yet PHP
+not yet support advanced language features, including object-oriented
+programming.
+Moreover, the range of standard library functions supported are mostly string operations, yet PHP
 provides a large collection of functions for array operations.
-All three analyses described in Section \ref{sec:existing_tools} though have
-only been applied to systems that are small and not maintained any more. Despite the
-functional benefits these tools provide to developers, their practicality for
-large and modern systems with respect to advanced language support has not been
-evaluated.
+All three tools described in Section \ref{sec:existing_tools} have
+only been applied to systems that are small and not actively maintained.
+Despite the functional benefits that these tools can provide to developers,
+their practicality for large and modern systems with respect to advanced language support has not been
+assessed.
 

### 2017-01-23 paper.tex
3 refactored

@@
 
+\definecolor{xlinkcolor}{HTML}{383838}
+\definecolor{xcitecolor}{HTML}{383838}
+\definecolor{xurlcolor}{HTML}{000055}
+
 \hypersetup{
     colorlinks,
-    linkcolor={red!65!black},
-    citecolor={blue!50!black},
-    urlcolor={blue!80!black}
+    linkcolor = xlinkcolor,
+    citecolor = xcitecolor,
+    urlcolor  = xurlcolor
 }
@@
 with around 300.000 lines of code that provides a vivid
-plug-in environment and is used.  The full implementation of \sf{Oak}, the test
-suite as well as or test corpus (see Section \ref{sec:experiment_setup}) can be found at
+plug-in environment and is actively maintained. The full implementation of
+\sf{Oak}, the test suite as well as or test corpus (see Section \ref{sec:experiment_setup}) can be found at
 \url{http://www.github.com/smba/oak}.
 
-For our implementation of a symbolic execution engine for PHP, Oak, we have
-chosen a test-driven approach with continuous integration using regression
-tests based on \sf{WordPress} and \sf{SchoolMate} (see Section
+For our implementation of a symbolic execution engine for PHP, \sf{Oak}, we
+chose a test-driven approach with continuous integration using regression tests
+based on \sf{WordPress} and \sf{SchoolMate} (see Section
 \ref{sec:experiment_setup}). For \sf{SchoolMate}, a rather straightforward and
@@
 approximation for \sf{WordPress} as we faced a number of obstacles hindering
-symbolic execution in different ways.
+symbolic execution.
 
 \subsection{Experience Report}\label{sec:experience_report}
-We document our design choices we made, apparent trade-offs we
+We document in this Section the design choices we made, trade-offs we
 recognized, and limitations we faced on our way towards a scalable symbolic
-execution engine; and what we have learned from bypassing limitations faced in
+execution engine, and what we have learned from bypassing limitations faced in
 spite of manual effort for future work.
@@
 \subsubsection{Symbolic/Concrete Loop Execution}
-Throughout the implementation we basically followed the symbolic execution
+Throughout the implementation we followed the symbolic execution
 semantics of \sf{PhpSync} \cite{Nguyen:2014:BCG:2635868.2635928}. As described
@@
 the number of iterations is indeterminable for most cases. While this is the
-case for while and do-while loops, the number of loop iterations for for or
-for-each loops is generally determinable and known before entering the loop
-body. For loops use a loop counter that usually is incremented or decremented
+case for while and do-while loops, the number of loop iterations for for-loops
+or for-each loops is generally determinable and known before entering the loop
+body. For-loops use a loop counter that usually is incremented or decremented
 after each loop iteration until a loop condition is not satisfied anymore;
 for-each loops simply iterate over a given array value.
+
 Our execution semantics follows this scheme and simulates concrete execution
 for those types of loops if possible. Otherwise, for instance, if a loop
-counter, loop counter limit for for loops, or variable for for-each loops is
+counter, loop counter limit for for-loops, or variable for for-each loops is
 symbolic, we resort to executing only a single iteration. This design choice
@@
 \ref{fig:multi_target} where two different objects can be assigned to variable \tt{\$person} in line 12 and
-14. For the method invocation in line 16 hence, the method is executed twice in
-total since we have two alternative target objects. To address this problem, we
+14 respectively. For the method invocation in line 16, hence, the method is
+executed twice in total since we have two alternative target objects. To address this problem, we
 implement program state merging for symbolic execution. For the example in
@@
 been assigned to a variable. This would allow to only pass the value summary to
-the method and only invocate it once, but for our execution to be sound, all
+the method and only invocate it once. For our execution though to be sound, all
 objects would need to be of the same class type. If for instance, in our
-example there would be another class definition with a method named
+example there would exist another class definition with a method named
 \tt{greet} and an instance of that class is assigned to variable \tt{\$person}
@@
 invocated methods for multiple targets since we cannot assert that all values
-contained in summary are of the same type. Nevertheless, to scale the target
-object space, in addition, we implemented execution modes only allowing
+contained in the value summary are of the same type. Nevertheless, to scale the
+target object space, in addition, we implemented execution modes only allowing
 execution of a method for a subset of possible targets and assuming symbolic
 return values for the remaining targets. This execution mode was necessary for
-WordPress for several, but not all entry points that we studied, in order to
-terminate. All subsequent analyses in this paper do not use this execution
-mode, yet it is provided in out implementation.
+\sf{WordPress} for several, but not all entry points that we studied, in order
+to terminate. All subsequent analyses in this paper do not use this execution
+mode, yet it is provided in our implementation.
 
@@
 twice. The include statements make an attempt to include a file, but do not
-return a warning in contrast to the require statements. Each of these
+return a warning in contrast to the \tt{require} statements. Each of these
 statements takes as input either a string constant (static include) or an
@@
 file that the interpreter finally attempts to include.
+
 Figure \ref{fig:dynamic_includes} shows a small example of a dynamic include containing
 information retrieved from a database. Since any input for a symbolically
-executed application is symbolic this information can be propagated to include
-expressions. Also, configuration files represent only one particular
+executed application is assumed to be symbolic this information can be
+propagated to include expressions. Also, configuration files represent only one particular
 configuration or even just default values or placeholders. Hence, symbolic
-execution is prone to fail for dynamic includes if symbolic information is contained.
-
-Second, function can either be invocated directly by statically providing the
-name of the function (or method), or indirectly. Indirect, or dynamic function
-invocation in PHP are enabled by using built-in functions like \tt{call\_user\_func},
-\tt{call\_user\_func\_array} or simply passing a string value with a direct
-function call to the \tt{eval} function. The first two functions take as
-arguments the name of the function two call, and all arguments to that
+execution is prone to fail for dynamic includes if symbolic values are
+contained.
+
+Second, a function can either be invocated directly by statically providing the
+name of the function (or method), or indirectly. Indirect (or dynamic) function
+invocations in PHP are enabled by using built-in functions like
+\tt{call\_user\_func}, \tt{call\_user\_func\_array} or simply passing a string
+value with a static function invocation to the \tt{eval} function. The first two
+functions take as arguments the name of the function two call, and all arguments to that
 functions either as additional arguments or a single associative array
@@
 literal.
-Again, we encountered issues similar to dynamic include expressions with
-\ref{fig:indirect_calls} illustrates two functions \tt{header\_serif} and
-\tt{header\_sserif} which both print a HTML headline. Depending on whether the
-value of the variable \tt{\$style} retrieved from a database is \tt{serif} or
-\tt{sserif}, a different function is called. Given that the function name is 
+Again, we encountered issues similar to those with dynamic include expressions.
+Figure \ref{fig:indirect_calls} illustrates two functions \tt{header\_serif}
+and \tt{header\_sserif} which both print a HTML headline. Depending on whether
+the value of the variable \tt{\$style} retrieved from a database is \tt{serif}
+or \tt{sserif}, a different function is called. Given that the function name is
 concrete at run-time, the desired function can be called. Otherwise, if the
-function name is assembled at run-time and contains symbolic values, no function
-can be called as the name is ambiguous or unknown. For this example, symbolic
-execution is not able to determine which function to call indirectly.
+function name is assembled at run-time and contains symbolic values, no
+function can be called as the name is ambiguous or unknown. For this example,
+symbolic execution is not able to determine which function to call dynamically.
 
@@
 In the previous subsection, we argued that dynamic includes and dynamic
-function invocations hinder symbolic execution  due to imprecise resolution of
-expression if symbolic information is contained. For callable code parts as
-included scripts or invocated functions, we attempted to bypass this limitation
+function invocations can hinder symbolic execution due to imprecise resolution
+of expressions if symbolic values are contained. For callable code sections as
+included scripts or functions, we attempted to bypass this limitation
 by manually providing proper concrete input.
@@
 Consider the code example in Figure \ref{fig:indirect_calls} where a function
-name could not be resolved precisely due to a missing string literal that is contained from a
-database query result. By manually providing either option \tt{serif} or
-\tt{sserif}, or both as alternatives, we are able to increase code coverage and
+name could not be resolved precisely due to a missing string literal that is
+obtained from a database query result. By manually providing either option
+\tt{serif} or \tt{sserif}, or both as alternatives, we are able to increase code coverage and
 the accuracy of our approximation with manual effort by making dynamic
@@
 problems we documented in Section \ref{sec:experience_report}. The evaluation of
-PhpSync, i.e., the PHP systems that the tool was applied to, did neither contain object-oriented
-features, nor dynamic features like dynamic includes or dynamic function
-invocation with symbolic values.
+\sf{PhpSyn}c, i.e., the PHP systems that the tool was applied to, did neither
+contain object-oriented features, nor dynamic features like dynamic includes or dynamic function
+invocation with symbolic values in a significant number.
 
 As the observations so far did only refer to \sf{WordPress} as a large-scale
-system, we wanted to study the impeding impact of dynamic features to symbolic
+system, we want to study the hindering impact of dynamic features to symbolic
 execution for a wider selection of PHP systems systematically apart from the
@@
 support a number of language features.
-The full list of PHP systems for that we evaluate our symbolic execution engine along with descriptive code metrics can be found in Table \ref{corpus}.
+The full list of PHP systems for that we evaluate our symbolic execution engine
+along with descriptive code metrics can be found in Table \ref{corpus}.
+
+All measurements in the following are cumulated for all possible entry points of a
+system, i.e., each script file. The symbolic execution operates with a timeout
+of ten minutes and terminates if this timeout is exceeded.
 

### 2017-01-23 paper.tex
References repair

@@
 \cite{phpNET} is a programming language focused on server-side application
-development. As of 2012, PHP was used by 78.8 percent of the ten million most
-popular websites (according to Alexa popularity ranking) \cite{alexaPHP}, ranking 7th on the TIOBE
-programming community index \cite{tiobePHP}, and was the ranked as the 6th
-most popular language on GitHub \cite{githubPHP}.
+development. As of 2017, PHP was used by 82.4 percent of the ten million most
+popular websites (according to Alexa popularity ranking) \cite{alexaPHP},
+ranking 10th on the TIOBE programming community index \cite{tiobePHP}, and, in
+2015,  was the ranked as the 4th most popular language on GitHub
+\cite{githubPHP}.
 
@@
 
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+     %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill
+    % etc.
       %(or a blank line to force the subfigure onto a new line)
@@
     \end{subfigure}
-    \vspace{2mm}
+    %\vspace{2mm}
     \caption{
@@
 	\end{subfigure}
-	\vspace{2mm}
+	\vspace{1mm}
 	\hfill
@@
 	\end{tabular}
-	\vspace{1mm}
+	%\vspace{1mm}
 	\caption{Corpus of twelve PHP systems. The file count includes files with a

### 2017-01-23 paper.tex
(a)spell checked

@@
-\documentclass[sigconf]{acmart}
+\documentclass[sigconf, preprint]{acmart}
 
@@
 With respect to the staged nature of web applications it is
-non-trivialstatically to determine whether a string literal is an output
-candidate, so we heuristically classify string literals as output candidates if
-they contain the characters \tt{<}, \tt{>}, or both since we expect output to
+not trivial to statically determine whether a string literal is an output
+candidate. Consequently, we heuristically classify string literals as output
+candidates if they contain the characters \tt{<}, \tt{>}, or both since we expect output to
 contain HTML tags.
@@
 		(Reached/Output, OnlyReached) and \emph{not-reached} output candidates
-		(FunctionContext, NotImported, Misc) during the analyis for each system.
+		(FunctionContext, NotImported, Misc) during the analysis for each system.
 		\label{coverage}}
@@
 From Section \ref{sec:experiment_dynamicfeatures} we learn that include
-expressions containic symbolic valuescannot be resolution precisely, and from
+expressions containing symbolic values cannot be resolution precisely, and from
 the previous Section \ref{WhatLiteralsDidWeMiss?} that very little of missed
@@
 Second, we measure how many reached include expressions are
-succesfully resolved throughout the analysis for each system.
+successfully resolved throughout the analysis for each system.
 
@@
 to be provided to decrease the number of assumed symbolic inputs and avoid
-dynamic features being confronted with symbolic values. Exising work approaching
+dynamic features being confronted with symbolic values. Existing work approaching
 this goal \cite{artzi_finding_2008,artzi_finding_2010,DynamicWassermann}
@@
 output for a web application is approximated using a hybrid approach: A
-dynamic webpage is executed with concrete input and the execution is
+dynamic web page is executed with concrete input and the execution is
 recorded at run-time. Changes in the client-side output then can be mapped to 
@@
 necessity of more comprehensive evaluation of developer tools as
-symbolic execution is a conceptually suitable, but not neccessary practical
+symbolic execution is a conceptually suitable, but not necessary practical
 approach to approximate client page output. Dynamic features of PHP in

### 2017-01-24 paper.tex
language refactoring

@@
-\documentclass[sigconf, preprint]{acmart}
+\documentclass[sigconf]{acmart}
+
+% removes copyright & conference info
+\settopmatter{printacmref=false} % Removes citation information below abstract
+\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with  conference information in first column 
+\pagestyle{plain} % removes running headers
 
@@
 %\setcopyright{acmlicensed}
-\setcopyright{rightsretained}
+%\setcopyright{rightsretained}
 %\setcopyright{usgov}
@@
 
-%\author{Stefan Mühlbauer}
+\author{Stefan Mühlbauer}
 \affiliation{%
-  %\institution{Technische Universität Braunschweig}
+  \institution{Technische Universität Braunschweig}
   %\streetaddress{P.O. Box 1212}
@@
 Knowledge about possible output variants enables a range of developer tools as
-well as analyses for vulnerability detection. However, previous tools have only
-been evaluated for smaller systems and scalability for larger systems has not
-been assessed.
+well as analyses for vulnerability detection. However, previous tools that
+leverage symbolically approximated output have only been evaluated for smaller
+systems and scalability for larger systems has not been assessed.
 
@@
 %
-\begin{CCSXML}
-<ccs2012>
- <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-  <concept_significance>500</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
-</ccs2012>  
-\end{CCSXML}
-
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
+%\begin{CCSXML}
+%<ccs2012>
+% <concept>
+%  <concept_id>10010520.10010553.10010562</concept_id>
+%  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+%  <concept_significance>500</concept_significance>
+% </concept>
+% <concept>
+%  <concept_id>10010520.10010575.10010755</concept_id>
+%  <concept_desc>Computer systems organization~Redundancy</concept_desc>
+%  <concept_significance>300</concept_significance>
+% </concept>
+% <concept>
+%  <concept_id>10010520.10010553.10010554</concept_id>
+%  <concept_desc>Computer systems organization~Robotics</concept_desc>
+%  <concept_significance>100</concept_significance>
+% </concept>
+% <concept>
+%  <concept_id>10003033.10003083.10003095</concept_id>
+%  <concept_desc>Networks~Network reliability</concept_desc>
+%  <concept_significance>100</concept_significance>
+% </concept>
+%</ccs2012>  
+%\end{CCSXML}
+
+%\ccsdesc[500]{Computer systems organization~Embedded systems}
+%\ccsdesc[300]{Computer systems organization~Redundancy}
+%\ccsdesc{Computer systems organization~Robotics}
+%\ccsdesc[100]{Networks~Network reliability}
 
@@
 
-\keywords{ACM proceedings, \LaTeX, text tagging}
+%\keywords{ACM proceedings, \LaTeX, text tagging}
 
@@
 consists of both static code, such as scripts, and dynamically generated code,
-such as client page output. The latter code, although assembled at
+such as client page output. The latter code, although generated at
 run-time, may contain client-side parts of the web application, such as
@@
 To investigate the question, whether we can have a practical and scalable
-symbolic execution engine, we re-implemented the symbolic engine with the
+symbolic execution engine, we re-implement the symbolic engine with the
 specifications of the previous symbolic execution semantics
 for \sf{PhpSync} \cite{Nguyen:2014:BCG:2635868.2635928} for PHP and
-extended it with additional features, such as object-oriented programming. We
+extend it with additional features, such as object-oriented programming. We
 evaluate our symbolic execution engine for several large-scale and modern PHP systems.
@@
 During the introduction of new semantics for object-oriented programming we
-faced a trade-off between accuracy and scalability: Whilst exploring multiple
-paths, for method invocations the number of target objects may grow rapidly and
-infeasible. Moreover, based on observations we identify limitations of symbolic execution
-for PHP. Dynamic features such as dynamic function invocations and dynamic
-include for an accurate execution require precisely resolved include targets
-and function names respectively. Since these expressions  can contain symbolic
-values at run-time, for many cases this restricts a symbolic execution engine
-from further execution as callable code sections become inaccessible.
+face a trade-off between accuracy and scalability: Whilst exploring multiple
+paths, for method invocations the number of target objects can grow rapidly and
+become infeasible. Moreover, based on observations we identify limitations of
+symbolic execution for PHP. For an accurate execution dynamic features, such as
+dynamic function invocations and dynamic includes, require precisely resolved
+include targets and function names respectively. Since these expressions  can
+contain symbolic values at run-time, for many cases this restricts a symbolic
+execution engine from further execution as callable code sections become
+inaccessible. 
 Although some effort can be spent to statically resolve dynamic include
@@
 execution, the tools using output approximation based on it and whose evaluation
-we criticize, and current conceptual and language-related limitations.
+we criticize as well as current conceptual and language-related limitations.
 
@@
 
-Therefore, for our execution engine to be sound, we were bound to execute
+Therefore, for our execution engine to be sound, we are bound to execute
 invocated methods for multiple targets since we cannot assert that all values
@@
 Aside from the previous issues, we identified two dynamic language features
-that we were not able to execute accurately. Dynamic features in PHP are
-features that take as input rather a expression that is resolved at run-time
+that, in many cases, we were not able to execute accurately. Dynamic features in
+PHP are features that take as input rather a expression that is resolved at run-time
 than a static constant value. For \sf{WordPress}, we faced missing files and
@@
 expression that is evaluated at run-time (dynamic include). The string that the
-expression evaluates to represents an absolute or relative paths to the target
+expression evaluates to represents an absolute or relative path to the target
 file that the interpreter finally attempts to include.
@@
 only using small-scale systems. Although evaluating a tool for a small systems
-can be a valuable proof-of-concept, it does not necessarily cover the question
+can be a valuable proof of concept, it does not necessarily cover the question
 of whether a tool is practical on a larger-scale. Therefore in the following we
@@
 better separate and understand the circumstances under which limitations might affect practicality. For the selection of large-scale systems
-we borrow three systems from a case study by Hills et al. \cite{hills2015variable} addressing the
-feature usage patterns and frequencies in PHP systems. Some more detailed description of this
+we borrowed three systems from a case study by Hills et al.
+\cite{hills2015variable} addressing the feature usage patterns and frequencies in PHP systems. Some more detailed description of this
 case study can be found in Section \ref{sec:related_work}. For our choice of modern small-scale systems we
@@
 candidates that are part of a function definition are only missed if the
-corresponding function is never invocated at run-time, or simply dead code.
+corresponding function is never invocated at run-time or simply just dead code.
 
@@
 for almost all systems tested (except for \textsf{AddressBook} and
-\textsf{TimeClock}) not-reached output candidates have a function context.
+\textsf{TimeClock}) not-reached output candidates had a function context.
 
@@
 incorporates dynamic symbolic execution, i.e. concrete execution alongside symbolic execution.
-
 Also, if the usage of tools using static output approximation with symbolic
@@
 the number of assumed symbolic inputs or systematically provide concrete data
-to increase the practical benefit of the tools described above.
+to increase the practicality of the tools described above.
 

### 2017-01-24 paper.tex
author(s) removed

@@
     keepspaces = true,
-    keywordstyle = \bfseries
+    keywordstyle = \bfseries,
+    %literate={\$}{{\textcolor{blue}{\$}}}1
 }

### 2017-02-03 paper.tex
snapshot 01-24-2017
Title -> "Does it Scale? Static Output Approximation For PHP Web Applications"
minor changes
creating related work file

@@
 \settopmatter{printacmref=false} % Removes citation information below abstract
-\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with  conference information in first column 
+%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with  conference information in first column 
 \pagestyle{plain} % removes running headers
@@
 
-\section{Related Work} \label{sec:related_work}
-
-\paragraph{Symbolic Execution}
-Symbolic execution \cite{King1976,Darringer1978} remained an conceptual approach
-for decades, but has been refined and extended over the last years. Several
-approaches incorporating constraint solving and leveraging concrete execution
-alongside symbolic execution emerged and enable automated test case generation;
-the scalability of these approaches is still challenged by path explosion and constraint solving \cite{CadarSen2013}.
-
-Dynamic symbolic execution \cite{CadarSen2013} incorporates symbolic execution
-alongside concrete execution and has shown promising results in automated test
-generation\cite{artzi_finding_2008,artzi_finding_2010,DynamicWassermann} for PHP web
-applications and might help mitigate the conceptual limitations faced with
-symbolic values and dynamic features for future work on output-oriented
-symbolic execution.
-
-An overview about recent evolution of symbolic execution is provided by work of Cadar et al.
-\cite{CadarSen2013}, as well as a more tool-centered perspective \cite{Cadar2011}.
-
-\paragraph{Static Output Approximation}
-To approximate program slices for web
-applications, Ricca et al. \cite{tonella_web_2005,tonella_2001,tonella_2002} approximate
-dynamically generated output. For output generating statements, such as \tt{echo} or
-\tt{print}, all strings are unquoted (code extrusion). If those statements
-contain variables, these are linked to string concatenations using a
-proposed flow analysis called \emph{string-cat propagation}. From these
-flows representing approximated output subsequent program slices are computed.
-
-Minamide \cite{minamide_static_2005} approximates client page output
-of web applications by describing possible output by a context-free
-grammar that is constructed statically from the PHP code for a given regular
-expression of user input. The constructed grammar enables analyses such as
-detecting cross-site scripting vulnerabilities by checking whether user input
-has been sanitized, and HTML validation by determining whether the constructed
-grammar is contained in a depth-bound HTML grammar. Based on Minamide's
-approximation approach several vulnerability analyses addressing cross-site
-scripting \cite{wassermann_static_2008} and SQL injection
-\cite{wassermann2007sound} have been proposed. Wang et al.
-\cite{wang_locating_2010} utilizes this string analysis to detect strings
-visible at the browser and enable internationalization of web applications.
-
-Another approach is proposed by Wang et al. \cite{wang_automating_2012}, where
-output for a web application is approximated using a hybrid approach: A
-dynamic web page is executed with concrete input and the execution is
-recorded at run-time. Changes in the client-side output then can be mapped to 
-corresponding PHP code using static impact analysis.
-
-\paragraph{Dynamic Features}
-According to a case study by Hills et al. \cite{Hills:2013:ESP:2483760.2483786}
-of the feature usage in PHP, based on a corpus of state-of-the-art-projects,
-dynamic includes are less frequently used than static includes, yet usage
-frequency varies from system to system. Further work by Hills et al.
-\cite{hills2014static,hills2014php} approaches static resolution of dynamic includes using
-both context-insensitive resolution on file level by simplifying PHP constants;
-and context-sensitive on program level for transitive includes. In spite of 
-promising results, these approaches, similar to our results, face limitations
-for truly dynamic includes if resolution is not sound due to information not
-being available in the source code like database query results.
+\input{related}
 

### 2017-02-03 related.tex
creating related work file

\section{Related Work} \label{sec:related_work}

\paragraph{Symbolic Execution}
Symbolic execution \cite{King1976,Darringer1978} remained an conceptual approach
for decades, but has been refined and extended over the last years. Several
approaches incorporating constraint solving and leveraging concrete execution
alongside symbolic execution emerged and enable automated test case generation;
the scalability of these approaches is still challenged by path explosion and constraint solving \cite{CadarSen2013}.

Dynamic symbolic execution \cite{CadarSen2013} incorporates symbolic execution
alongside concrete execution and has shown promising results in automated test
generation\cite{artzi_finding_2008,artzi_finding_2010,DynamicWassermann} for PHP web
applications and might help mitigate the conceptual limitations faced with
symbolic values and dynamic features for future work on output-oriented
symbolic execution.

An overview about recent evolution of symbolic execution is provided by work of Cadar et al.
\cite{CadarSen2013}, as well as a more tool-centered perspective \cite{Cadar2011}.

\paragraph{Static Output Approximation}
To approximate program slices for web
applications, Ricca et al. \cite{tonella_web_2005,tonella_2001,tonella_2002} approximate
dynamically generated output. For output generating statements, such as \tt{echo} or
\tt{print}, all strings are unquoted (code extrusion). If those statements
contain variables, these are linked to string concatenations using a
proposed flow analysis called \emph{string-cat propagation}. From these
flows representing approximated output subsequent program slices are computed.

Minamide \cite{minamide_static_2005} approximates client page output
of web applications by describing possible output by a context-free
grammar that is constructed statically from the PHP code for a given regular
expression of user input. The constructed grammar enables analyses such as
detecting cross-site scripting vulnerabilities by checking whether user input
has been sanitized, and HTML validation by determining whether the constructed
grammar is contained in a depth-bound HTML grammar. Based on Minamide's
approximation approach several vulnerability analyses addressing cross-site
scripting \cite{wassermann_static_2008} and SQL injection
\cite{wassermann2007sound} have been proposed. Wang et al.
\cite{wang_locating_2010} utilizes this string analysis to detect strings
visible at the browser and enable internationalization of web applications.

Another approach is proposed by Wang et al. \cite{wang_automating_2012}, where
output for a web application is approximated using a hybrid approach: A
dynamic web page is executed with concrete input and the execution is
recorded at run-time. Changes in the client-side output then can be mapped to 
corresponding PHP code using static impact analysis.

\paragraph{Dynamic Features}
According to a case study by Hills et al. \cite{Hills:2013:ESP:2483760.2483786}
of the feature usage in PHP, based on a corpus of state-of-the-art-projects,
dynamic includes are less frequently used than static includes, yet usage
frequency varies from system to system. Further work by Hills et al.
\cite{hills2014static,hills2014php} approaches static resolution of dynamic includes using
both context-insensitive resolution on file level by simplifying PHP constants;
and context-sensitive on program level for transitive includes. In spite of 
promising results, these approaches, similar to our results, face limitations
for truly dynamic includes if resolution is not sound due to information not
being available in the source code like database query results.

### 2017-02-03 related.tex
first draft

@@
 \paragraph{Symbolic Execution}
-Symbolic execution \cite{King1976,Darringer1978} remained an conceptual approach
-for decades, but has been refined and extended over the last years. Several
-approaches incorporating constraint solving and leveraging concrete execution
-alongside symbolic execution emerged and enable automated test case generation;
-the scalability of these approaches is still challenged by path explosion and constraint solving \cite{CadarSen2013}.
+An overview about recent evolution of symbolic execution is provided
+by Cadar {\em et al.}~\cite{CadarSen2013}, as well as a more
+tool-centered perspective~\cite{Cadar2011}.
+%
+Symbolic execution \cite{King1976,Darringer1978} remained a
+conceptual approach for decades, but has been refined and extended
+over the last years. Several approaches incorporating constraint
+solving and leveraging concrete execution alongside symbolic execution
+emerged and enable automated test case generation; the scalability of
+these approaches is still challenged by path explosion and constraint
+solving.
 
-Dynamic symbolic execution \cite{CadarSen2013} incorporates symbolic execution
-alongside concrete execution and has shown promising results in automated test
-generation\cite{artzi_finding_2008,artzi_finding_2010,DynamicWassermann} for PHP web
-applications and might help mitigate the conceptual limitations faced with
-symbolic values and dynamic features for future work on output-oriented
-symbolic execution.
+\paragraph{Dynamic symbolic execution} Cedar and Sen~\cite{CadarSen2013} incorporates symbolic execution
+alongside concrete execution and has shown promising results in
+automated test
+generation~\cite{artzi_finding_2008,artzi_finding_2010,DynamicWassermann}
+for PHP web applications and might help mitigate the conceptual
+limitations faced with symbolic values and dynamic features for future
+work on output-oriented symbolic execution.
 
-An overview about recent evolution of symbolic execution is provided by work of Cadar et al.
-\cite{CadarSen2013}, as well as a more tool-centered perspective \cite{Cadar2011}.
+\paragraph{Symbolic Execution and Dynamic Analysis for PHP-based Web Applications}
+PHPQuickFix and PHPRepair~\cite{ibm-icse12} detect and fix errors in
+server-side PHP applications leading to ill-formed generated HTML.
+PHPQuickFix examines {\em constant prints}, i.e., the PHP statements
+that print directly string literals and repairs HTML ill-formed
+errors, such as mismatched start and end tags and missing special
+characters within those literals. It analyzes each string literal
+separately and can only identify local issues. In contrast, PHPRepair
+follows a dynamic approach in which a given test suite is used to
+generate client-side code with different server-side executions, while
+tracing the origin of output strings.
+%% Final
+They identify ill-formed HTML code, infer and automatically apply
+fixes via using an external oracle and a string constraint solver.  In
+contrast to their dynamic strategy, we aim to study symbolic execution
+in this work.
+
+Wang {\em et al.}~\cite{wang-fse12}'s approach executes the
+(instrumented) server-side code with a specific input to generate the
+client code. Changes in that generated code can then be mapped to
+their origins in the PHP code using the recorded run-time mappings and
+static impact analysis.
+%We similarly track origin locations, but we symbolically execute the
+%PHP code.
+%
+%We previously used our symbolic execution engine \emph{PhpSync} for
+%similar purposes~\cite{drc-ase13,phpsync}. First, we 
+PHPSync~\cite{phpsync} used our static origin tracking to propagate
+changes in the client code (output of symbolic execution) back to the
+PHP code. Subsequently, the same authors designed DRC to analyze both
+PHP code and generated client-side code to detect cross-language and
+cross-stage dangling references~\cite{drc-ase13}. DRC matches path
+constraints of references/declarations without fully symbolic
+execution. It uses the path conditions from symbolic execution to
+determine in which configurations dangling references can occur.
+%
+%In contrast to this work, DRC extracts program entities 
+%(identifiers)
+%via heuristics:
+%%and {\em is not} able to parse and understand the structure and
+%%semantics of the embedded code; DRC
+%it just matches path constraints of
+%references/declarations without building the ASTs and DOM for embedded
+%code in different configurations.
+%%In the future, we plan to integrate DRC's analyses to derive
+%%additional call-graph edges that cross languages and stages.
+
+Apollo~\cite{artzi-icse10,artzi-tse10}, a more general fault
+localization tool for PHP code, rates the PHP echo/print
+statements higher in suspiciousness by running an instrumented
+interpreter on test cases. Clark {\em et al.}~\cite{clark-ase11}
+localize faults in an SQL command embedded within JSP code. Although
+not in our focus, we could conceptually use our tool to approximate
+embedded SQL queries and use a variability-aware SQL parser for
+analysis.
+
+%-------------------
+
+%-------------------------------------
 
 \paragraph{Static Output Approximation}
-To approximate program slices for web
-applications, Ricca et al. \cite{tonella_web_2005,tonella_2001,tonella_2002} approximate
-dynamically generated output. For output generating statements, such as \tt{echo} or
-\tt{print}, all strings are unquoted (code extrusion). If those statements
-contain variables, these are linked to string concatenations using a
-proposed flow analysis called \emph{string-cat propagation}. From these
-flows representing approximated output subsequent program slices are computed.
+To approximate program slices for web applications, Ricca et
+al. \cite{tonella_web_2005,tonella_2001,tonella_2002} approximate
+dynamically generated output. For output generating statements, such
+as \tt{echo} or \tt{print}, all strings are unquoted (code
+extrusion). If those statements contain variables, these are linked to
+string concatenations using a proposed flow analysis called
+\emph{string-cat propagation}. From these flows representing
+approximated output subsequent program slices are computed.
 
@@
 
+%Minamide proposed a string analyzer~\cite{minamide05} that takes a PHP
+%program and a regular expression describing the input, and validates
+%approximate HTML output via context-free grammar analysis.
+%regular expression describing all of its possible inputs, and then
+%statically approximates and validates the output via a context-free
+%grammar. In comparison, his goal is to validate approximated HTML
+%outputs from a PHP program {\em without fixing support}. Moreover,
+%{\tool} performs symbolic execution
+%to produce D-models without
+%requiring an input specification as in Minamide's.
+%%Using a similar string analyzer,
+%Wang {\em et al.}~\cite{wang-fse10} compute the approximated output of
+%PHP code and identify the constant strings visible from a browser
+%for translation. Both do not aim to analyze multiple variants of
+%embedded code in JS or CSS.
+% getting Telvik Bultan work
+
+Several string taint-analysis techniques were built for PHP web
+programs and software-security
+problems~\cite{kieyzun09,su-icse08,xieaiken06,tevfik11}.  The key idea
+behind string tainting is to check the code expression by expression,
+variable by variable to have a list of all the ones potentially
+influenced by outside input.  If these variables are used in running a
+dangerous command, we will have potentially dangerous tainted
+variables.
+%
+Gould {\em et al.}~\cite{gould04} use string analysis to guarantee
+well-typed SQL queries generated by a Java program. The type system
+in~\cite{tabuchi03} is based on regular expressions with string
+concatenation and pattern matching.
+
+%A CFG-based type system for string analysis is presented
+%in~\cite{thiemann05}.
+
 \paragraph{Dynamic Features}

### 2017-02-03 related.tex
revised version

@@
 
-\paragraph{Symbolic Execution}
-An overview about recent evolution of symbolic execution is provided
-by Cadar {\em et al.}~\cite{CadarSen2013}, as well as a more
+\paragraph{Symbolic execution}
+A literature review on symbolic execution is provided by Cadar in a
+recent overview {\em et al.}~\cite{CadarSen2013}, as well as in a more
 tool-centered perspective~\cite{Cadar2011}.
@@
 %similar purposes~\cite{drc-ase13,phpsync}. First, we 
-PHPSync~\cite{phpsync} used our static origin tracking to propagate
+PHPSync~\cite{phpsync} used static origin tracking to propagate
 changes in the client code (output of symbolic execution) back to the
@@
 determine in which configurations dangling references can occur.
-%
-%In contrast to this work, DRC extracts program entities 
-%(identifiers)
-%via heuristics:
-%%and {\em is not} able to parse and understand the structure and
-%%semantics of the embedded code; DRC
-%it just matches path constraints of
-%references/declarations without building the ASTs and DOM for embedded
-%code in different configurations.
-%%In the future, we plan to integrate DRC's analyses to derive
-%%additional call-graph edges that cross languages and stages.
 
@@
 interpreter on test cases. Clark {\em et al.}~\cite{clark-ase11}
-localize faults in an SQL command embedded within JSP code. Although
-not in our focus, we could conceptually use our tool to approximate
-embedded SQL queries and use a variability-aware SQL parser for
-analysis.
-
-%-------------------
-
-%-------------------------------------
+localize faults in an SQL command embedded within JSP code.
 

### 2017-02-04 paper.tex
first draft
revised related work

@@
 \begin{document}
-\title{Does It Scale? Symbolic Execution of PHP Web Applications}
+%\title{Does It Scale? Symbolic Execution of PHP Web Applications}
+\title{Output Approximation for PHP at Scale:\\ Experience and Analysis}
 %\titlenote{Produces the permission block, and copyright information}
@@
 approach to approximate client page output. Dynamic features of PHP in
-combination with symbolic values though can hinder precise symbolic execution
+combination with symbolic values can hinder precise symbolic execution
 and limit the scalability and practicality of tools relying on symbolically
@@
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{bibliography}
+\bibliography{bibliography,dblp2_short,tien}
 

### 2017-02-04 related.tex
revised related work

@@
 determine in which configurations dangling references can occur.
-
+%
 Apollo~\cite{artzi-icse10,artzi-tse10}, a more general fault
@@
 
-Another approach is proposed by Wang et al. \cite{wang_automating_2012}, where
+In another approach by Wang et al. \cite{wang_automating_2012}, 
 output for a web application is approximated using a hybrid approach: A
@@
 corresponding PHP code using static impact analysis.
+
+Several string taint-analysis techniques were built for PHP web
+programs and software-security
+problems~\cite{kieyzun09,su-icse08,xieaiken06,tevfik11}.  The key idea
+behind string tainting is to check the code expression by expression,
+variable by variable to have a list of all the ones potentially
+influenced by outside input.  If these variables are used in running a
+dangerous command, we will have potentially dangerous tainted
+variables.
+%
+Gould {\em et al.}~\cite{gould04} use string analysis to guarantee
+well-typed SQL queries generated by a Java program. The type system
+in~\cite{tabuchi03} is based on regular expressions with string
+concatenation and pattern matching.
+
+\paragraph{Dynamic Output Approximation}
+Zou {\em et al.}~\cite{zou-issta14} introduce a V-DOM coverage for web
+applications. They convert a PHP program into C code and performs
+static analysis for control flows and data flows to build a V-DOM
+tree. V-DOM coverage is defined as the ratio of the number of covered
+DOM objects over the total number of DOM
+objects. DomCovery~\cite{mesbah-issta14} is another DOM-based coverage
+criteria for web applications. The coverage is defined at two levels:
+(1) the percentage of DOM states and transitions covered in the total
+state space, and (2) the percentage of elements covered in each
+particular DOM state.
 
@@
 
-Several string taint-analysis techniques were built for PHP web
-programs and software-security
-problems~\cite{kieyzun09,su-icse08,xieaiken06,tevfik11}.  The key idea
-behind string tainting is to check the code expression by expression,
-variable by variable to have a list of all the ones potentially
-influenced by outside input.  If these variables are used in running a
-dangerous command, we will have potentially dangerous tainted
-variables.
-%
-Gould {\em et al.}~\cite{gould04} use string analysis to guarantee
-well-typed SQL queries generated by a Java program. The type system
-in~\cite{tabuchi03} is based on regular expressions with string
-concatenation and pattern matching.
-
 %A CFG-based type system for string analysis is presented
@@
 
-\paragraph{Dynamic Features}
+\paragraph{Dynamic Features in Web applications}
 According to a case study by Hills et al. \cite{Hills:2013:ESP:2483760.2483786}

### 2017-10-11 paper.tex
minor improvements

@@
 tools for detecting and locating HTML validation errors
-\cite{Nguyen:2011:AFH:2190078.2190142}, computing program slices across
+\cite{phpsync}, computing program slices across
 server-side and client-page code \cite{Nguyen:2015:CPS:2786805.2786872}, or
@@
 actively maintained. The tools presented by previous work
-\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2011:AFH:2190078.2190142}
+\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,phpsync}
 though are only practical, if for a given system the symbolic execution engine
@@
 	source code responsible for the defect, which then can be refactored manually
-	\cite{Nguyen:2011:AFH:2190078.2190142}.
+	\cite{phpsync}.
 
@@
 for static output approximation in previous work
-\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140}.
+\cite{phpsync,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140}.
 Our vision is a practical symbolic execution, i.e., it computes an accurate
@@
 execution engine and tools described in previous work
-\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140}
+\cite{phpsync,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140}
 are the missing support for widely-used language features, and the evaluation
@@
 \cite{hills2015variable} addressing the feature usage patterns and frequencies in PHP systems. Some more detailed description of this
-case study can be found in Section \ref{sec:related_work}. For our choice of modern small-scale systems we
+case study can be found in Section \ref{sec:related_work}.
+Many of the older small-scale systems in our corpus have been studied before in similar contexts.
+For instance, Nguyen et al.~\cite{phpsync,Nguyen:2014:BCG:2635868.2635928} have selected among others these systems from \sf{sourceforge.net} while considering different sizes and domains as well as little use of object-oriented constructs.
+For our choice of modern small-scale systems we
 selected five systems from a list of recent Content Management Systems (CMS) written in PHP \cite{codegeekz}. For both
@@
 \end{figure}
+\vspace{-3mm}
 
@@
 \paragraph{Dynamic features are problematic.}
-From out observations with \sf{WordPress} in Section \ref{sec:3} we learn that
+From our observations with \sf{WordPress} in Section \ref{sec:3} we learn that
 dynamic language features, in particular include expressions and dynamic
@@
 dynamic features being confronted with symbolic values. Existing work approaching
-this goal \cite{artzi_finding_2008,artzi_finding_2010,DynamicWassermann}
+this goal \cite{artzi_finding_2008,artzi-tse10,DynamicWassermann}
 incorporates dynamic symbolic execution, i.e. concrete execution alongside symbolic execution.
@@
 presented in previous work
-\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140,minamide_static_2005,wassermann2007sound}
+\cite{phpsync,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140,minamide_static_2005,wassermann2007sound}
 where only small PHP systems have been studied. These papers introduce
 techniques for static output approximation of PHP web applications
-\cite{minamide_static_2005,Nguyen:2014:BCG:2635868.2635928,wang_automating_2012} or propose tools
+\cite{minamide_static_2005,Nguyen:2014:BCG:2635868.2635928,wang-fse12} or propose tools
 based on respective approximations
-\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140,wassermann2007sound,wassermann_static_2008}.
+\cite{phpsync,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140,wassermann2007sound,wassermann_static_2008}.
 As these approaches address practical use cases, they require a more and further comprehensive evaluation in order to  better understand possible limitations and assess their practicality, in particular for large-scale systems.

### 2017-10-11 related.tex
minor improvements

@@
 \paragraph{Symbolic execution}
-A literature review on symbolic execution is provided by Cadar in a
-recent overview {\em et al.}~\cite{CadarSen2013}, as well as in a more
+A literature review on symbolic execution is provided by Cadar and Sen~\cite{CadarSen2013} in a
+recent overview, as well as in a more
 tool-centered perspective~\cite{Cadar2011}.
@@
 
-\paragraph{Dynamic symbolic execution} Cedar and Sen~\cite{CadarSen2013} incorporates symbolic execution
+\paragraph{Dynamic symbolic execution} Cadar and Sen~\cite{CadarSen2013} incorporates symbolic execution
 alongside concrete execution and has shown promising results in
 automated test
-generation~\cite{artzi_finding_2008,artzi_finding_2010,DynamicWassermann}
+generation~\cite{artzi_finding_2008,artzi-tse10,DynamicWassermann}
 for PHP web applications and might help mitigate the conceptual
@@
 
-In another approach by Wang et al. \cite{wang_automating_2012}, 
+In another approach by Wang et al. \cite{wang-fse12}, 
 output for a web application is approximated using a hybrid approach: A
@@
 programs and software-security
-problems~\cite{kieyzun09,su-icse08,xieaiken06,tevfik11}.  The key idea
+problems~\cite{kieyzun09,wassermann_static_2008,xieaiken06,tevfik11}.  The key idea
 behind string tainting is to check the code expression by expression,
